### 2024-04-15 paper.tex
Initial commit

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Challenges and Solutions of Developing and
Implementing a Revolutionary Novel Desktop-as-a-Service}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.

\begin{credits}
\subsubsection{\ackname} A bold run-in heading in small font size at the end of the paper is
used for general acknowledgments, for example: This study was funded
by X (grant number Y).

\subsubsection{\discintname}
It is now necessary to declare any competing interests or to specifically
state that the authors have no competing interests. Please place the
statement with a bold run-in heading in small font size beneath the
(optional) acknowledgments\footnote{If EquinOCS, our proceedings submission
system, is used, then the disclaimer can be provided directly in the system.},
for example: The authors have no competing interests to declare that are
relevant to the content of this article. Or: Author A has received research
grants from Company W. Author B has received a speaker honorarium from
Company X and owns stock in Company Y. Author C is a member of committee Z.
\end{credits}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}, last accessed 2023/10/25
\end{thebibliography}
\end{document}

### 2024-04-15 paper.tex
More content

@@
 %
-\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
-Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
-Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
+\author{Christian Baun\orcidID{0000-1111-2222-3333} \and
+Johannes Bouché\orcidID{1111-2222-3333-4444}}
 %
@@
 %
-\institute{Princeton University, Princeton NJ 08544, USA \and
-Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
+\institute{Faculty of Computer Science and Engineering, Frankfurt University of Applied Sciences,\\
+Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
 \email{lncs@springer.com}\\
@@
 \begin{abstract}
-The abstract should briefly summarize the contents of the paper in
-150--250 words.
+The paper will cover very novel and non-published knowledge we gained
+from developing and implementing a DaaS that is funded by the German
+Federal Ministry for Economic Affairs and Climate Action. The topic
+fits very well into the topics of the conference.
 
@@
 %
-\section{First Section}
-\subsection{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
+\section{Introduction}
+% \subsection{A Subsection Sample}
+% Please note that the first paragraph of a section or subsection is
+% not indented. The first paragraph that follows a table, figure,
+% equation etc. does not need an indent, either.
+% 
+% Subsequent paragraphs, however, are indented.
 
-Subsequent paragraphs, however, are indented.
+\subsection{Motivation}
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
+\section{Related Work}
 
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
+\section{Architecture}
+
+TODO: Christian
+
+\section{Challenges and Solutions}
+
+TODO: Christian
+
+\subsection{Compatibility}
+
+TODO: Christian
+
+Anwendungen (Windows / Linux / MacOS)
+
+Sound (via sound-deamon oder via Guacamole)
+
+Drucker (via cups oder via Guacamole)
+
+Externe Geräte (USB-Stick)
+
+\subsection{Performance}
+
+Welche Performance-Paramter / Kriterien sind hier relevant für die Anwendung / Nutzung
+Fokus: Bottlenecks, Möglichkeiten der Performance-Steigerung, Grenzen/Limits, die man akzeptieren muss
+
+Latenz hängt auch immer von Entfernung und Übertragungsmedium ab.
+
+Skalierbarkeit: Stand und Möglichkeiten
+
+TODO Christian: Latenz bei Cloud-Gaming Erfahrungen
+
+\subsection{Stability}
+
+Was ist hier relevant?
+
+Verfügbarkeit des Service und der Benutzerdaten
+
+Redundante Komponenten (Hardware und Software)
+
+Welche Komponenten kann man sinnvoll redundant vorhalten?
+
+Was sind die Vor- und Nachteile?
+
+Auch Kosten?
+
+Wo sehen wir aktuell in unserer Architektur Risiken für Dienstausfall und Datenverlust?
+
+Features von Proxmox
+
+Docker haben/nicht haben / Vor- und Nachteile
+
+Hier auch auf die Skalierbarkeit beziehen
+
+\subsection{Usability}
+
+\newpage
 

### 2024-04-15 paper.tex
More content

@@
 %
-\author{Christian Baun\orcidID{0000-1111-2222-3333} \and
-Johannes Bouché\orcidID{1111-2222-3333-4444}}
+\author{Christian Baun\orcidID{0009-0004-9955-3752} \and
+Johannes Bouché\orcidID{TODO}}
 %
-\authorrunning{F. Author et al.}
+\authorrunning{C. Baun, J. Bouché}
 % First names are abbreviated in the running head.
@@
 Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
-\email{lncs@springer.com}\\
-\url{http://www.springer.com/gp/computer-science/lncs} \and
-ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
-\email{\{abc,lncs\}@uni-heidelberg.de}}
+\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
 %
@@
 Federal Ministry for Economic Affairs and Climate Action. The topic
-fits very well into the topics of the conference.
+fits very well into the topics of the conference. TODO: Christian
 
-\keywords{First keyword  \and Second keyword \and Another keyword.}
+\keywords{DaaS \and Compatibility \and Performance \and Stability \and Usability}
 \end{abstract}
@@
 \section{Introduction}
-% \subsection{A Subsection Sample}
-% Please note that the first paragraph of a section or subsection is
-% not indented. The first paragraph that follows a table, figure,
-% equation etc. does not need an indent, either.
-% 
-% Subsequent paragraphs, however, are indented.
 
-\subsection{Motivation}
+Inkl. Motivation
+
+TODO: Christian (1)
+TODO: Johannes (2)
 
@@
 
+TODO: Christian (1)
+TODO: Johannes (2)
+
 \section{Architecture}
 
-TODO: Christian
+TODO: Christian (1)
+TODO: Johannes (2)
 
@@
 
-TODO: Christian
+TODO: Christian (1)
+TODO: Johannes (2)
 
@@
 
+TODO: Johannes (1)
+TODO: Christian (2)
+TODO Christian: Latenz bei Cloud-Gaming Erfahrungen
+
 Welche Performance-Paramter / Kriterien sind hier relevant für die Anwendung / Nutzung
@@
 
-TODO Christian: Latenz bei Cloud-Gaming Erfahrungen
+Ceph-Dateisystem
+
+Datenbank SQlite aktuell. Perspektivisch MariaDB
+
+Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host.
 
 \subsection{Stability}
+
+TODO: Johannes (1)
+TODO: Christian (2)
 
@@
 
+Ceph-Dateisystem
+
+Datenbank SQlite aktuell. Perspektivisch MariaDB
+
+Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host. Proxmox kann sich darum kümmern?!
+
 Docker haben/nicht haben / Vor- und Nachteile
@@
 
+TODO Johannes: Beschreibung zur Position der Komponenten, Möglichkeiten der Parallelisierung und Einschränkungen
+
 \subsection{Usability}
 
-\newpage
+TODO: Christian (1)
+TODO: Johannes (2)
 
-\begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
-\begin{tabular}{|l|l|l|}
-\hline
-Heading level &  Example & Font size and style\\
-\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
-\hline
-\end{tabular}
-\end{table}
+Ideen und Konzepte zur Trennung von User und Admin-Ansicht
 
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
+Reduktion der Komplexität der Bedienung durch Fokus auf notwendigste Interaktionsschritte
 
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig1}
-\end{figure}
+Desktop vs. einzelne Anwendung im Browser-Tab
 
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
-book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
-and a homepage~\cite{ref_url1}. Multiple citations are grouped
-\cite{ref_article1,ref_lncs1,ref_book1},
-\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
+Anpassung des Browser-Inhalts auf die Größe der Anwendung
 
-\begin{credits}
-\subsubsection{\ackname} A bold run-in heading in small font size at the end of the paper is
-used for general acknowledgments, for example: This study was funded
-by X (grant number Y).
+Herausforderungen, Entwicklung, Implementierung, Lessons-learned
 
-\subsubsection{\discintname}
-It is now necessary to declare any competing interests or to specifically
-state that the authors have no competing interests. Please place the
-statement with a bold run-in heading in small font size beneath the
-(optional) acknowledgments\footnote{If EquinOCS, our proceedings submission
-system, is used, then the disclaimer can be provided directly in the system.},
-for example: The authors have no competing interests to declare that are
-relevant to the content of this article. Or: Author A has received research
-grants from Company W. Author B has received a speaker honorarium from
-Company X and owns stock in Company Y. Author C is a member of committee Z.
-\end{credits}
-%
+TODO Johannes: Erlebnisse der letzten 3-4 Wochen aufschreiben
+
+\section{Conclusion}
+
+TODO: Zusammen am Schluss
+
+\section{Outlook}
+
+TODO: Zusammen am Schluss
+
+\section*{Acknowledgements}
+
+This work was funded by the Federal Ministry for Economic Affairs and Climate Action
+('\textsl{Bundesministerium f\"ur Wirtschaft und Klimaschutz}')
+in the framework of the central innovation programme
+for small and medium-sized enterprises
+('\textsl{Zentrales Innovationsprogramm Mittelstand}').
+
+We thank our project partners from Nuromedia GmbH for their support.
+We especially thank Björn Goetschke, Mike Ludemann, Dario Savella, Holger Sprengel, and Rahul Tomar.
+
+\section*{References}
+
 % ---- Bibliography ----
@@
 %
-\begin{thebibliography}{8}
-\bibitem{ref_article1}
-Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)
 
-\bibitem{ref_lncs1}
-Author, F., Author, S.: Title of a proceedings paper. In: Editor,
-F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
-Springer, Heidelberg (2016). \doi{10.10007/1234567890}
+% \bibliography{\jobname_no_dashes_at_repeated_names,references}{}
 
-\bibitem{ref_book1}
-Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
-Location (1999)
-
-\bibitem{ref_proc1}
-Author, A.-B.: Contribution title. In: 9th International Proceedings
-on Proceedings, pp. 1--2. Publisher, Location (2010)
-
-\bibitem{ref_url1}
-LNCS Homepage, \url{http://www.springer.com/lncs}, last accessed 2023/10/25
-\end{thebibliography}
 \end{document}

### 2024-04-19 paper.tex
Add chapters performance + stability (drafts)

@@
 \section{Related Work}
-
+% TODO:  Introduce: Infratstructure-as-a-Service  and others (IaaS, PaaS, SaaS, DaaS)
 TODO: Christian (1)
@@
 \subsection{Performance}
+% Initial brainstorm:
+% + Welche Performance-Paramter / Kriterien sind hier relevant für die Anwendung / Nutzung
+% + Fokus: Bottlenecks, Möglichkeiten der Performance-Steigerung, Grenzen/Limits, die man akzeptieren muss
+% + Latenz hängt auch immer von Entfernung und Übertragungsmedium ab.
+% + Skalierbarkeit: Stand und Möglichkeite
+% + Ceph-Dateisystem
+% + Datenbank SQlite aktuell. Perspektivisch MariaDB
+% + Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host.
+%
+% Structure:
+% performance of:
+% - our serice itself
+% - backend services
+% - Viewer performance
+% Limiting Factors
+% - Hosting
+% - Network
+% - Internal data and communication
+% Chosen techniques/solutions/protocols:
+% - Filesystem (Ceph)
+% - Database (sqlite, mariadb)
+% - Guacamole (Proxy approach)
+% Similar_Examples:
+% - Cloud Gaming efforts 
+%    - technical solutions, 
+%    - latencies, 
+%    - hardware support and integration)
+%
+% TODO: Johannes (1)
+% TODO: Christian (2)
+% TODO: Christian: Latenz bei Cloud-Gaming Erfahrungen
+% TODO: 
+%       - Include overall takeaway messages somewhere before:
+%           - IaaS, PaaS and SaaS combined is defined as DaaS
+%           - Define terms: Backend-service, Host-System, Instance
+%           - We want a stable and usable system first, performance comes second
+%           - Low effort to use and integrate is the most crucial aspect!!!
+%           - No "Gamestation", but all modern applications shall be suported
+%               - "Regular" office apps (Spreadsheets, Editor, Email, Filemanager)
+%               - 3D modelling (Blender, Mayo, etc.)
+%               - Audio/Video Editing? (kdenlive, Adobe Premiere, OBS?)
+%               - Services (Webserver, Emailserver, Fileserver, etc.)
+%           - If several solutions are plausible, implement all and let the user decide
+%           - Give definitions for all specified daas-usecase
+%       - More references
+%       - Illustrate each subsection
+%       - Typos, grammar, etc.
 
-TODO: Johannes (1)
-TODO: Christian (2)
-TODO Christian: Latenz bei Cloud-Gaming Erfahrungen
+In order to describe the mission-critical performance aspects of our system,
+we initially give a short overview on potential challenges,
+inherent bottlenecks as well as other problem areas.
+We do that with respect to the underlying usecase model of DaaS-Systems
+and then follow with a more thorough description
+of our own specific design decissions and particular solutions.
 
-Welche Performance-Paramter / Kriterien sind hier relevant für die Anwendung / Nutzung
-Fokus: Bottlenecks, Möglichkeiten der Performance-Steigerung, Grenzen/Limits, die man akzeptieren muss
+From our perspective, performance in general can be seen as the most crucial system property
+which influences a majority of all other other relevant properties,
+such as user experience, system responsivness and overall efficiency.
+As a result, we had to address multifaceted
+and potentially cross-concerning challenges
+on several distinct layers within our own architecture.
 
-Latenz hängt auch immer von Entfernung und Übertragungsmedium ab.
+From the core infratructure to backend services
+and user-facing elements, each component inidvidually
+has to achieve certain minimum requirements in its own domain
+and in order to contribute to the overall performance goals of the whole system.
+Hence, we outline key limiting factors that impede performance,
+along with our experiences and tailored solutions
+aimed at mitigating these challenges.
+From our point of view the three most critial performance aspects
+for Daas-Systems are:
+\begin{itemize}
+	\item Infrastructure and Platform
+	\item Network Latency
+	\item Internal Processing and Communication
+\end{itemize}
+As there is no generalized measurment system available,
+we ordered them by their magnitude of impact
+and based them to the best of our knowledge
+on the obtained experiences during our research project.
+Since each of them have their own characteristics,
+specific behavior and requirements, implying a potential impact
+on the overall system performance and resource utilization,
+we describe them individually and in more detail
+in the following seperate subsections.
 
-Skalierbarkeit: Stand und Möglichkeiten
+\subsubsection{Infrastructure and Platform}
+In general can be said, that the chosen infrastructure, platform and utilized software
+may have the most influental impact on cpu and resource consumption
+which has a dramatic impact on the overall system performance.
+As a DaaS combines all aspects of IaaS, PaaS, SaaS into one system
+each related performance characteristic has to be taken into account.
 
-Ceph-Dateisystem
+For Infrastructure-as-a-Service can be said
+that our solution generally offers possibilities
+to use either virtual machines or docker containers.
+As can be seen in several scientific publications
+discussing relevant performance aspects of such systems,
+the obvious advanatge in most categories
+has a lightweight container solution compared to virtual machine
+if a identical program was tested in both environments
+~\cite{felter2015updated}
+~\cite{potdar2020performance}
+~\cite{seo2014performance}.
 
-Datenbank SQlite aktuell. Perspektivisch MariaDB
+On one hand, that implies for our solution,
+that a container-based approach is most likely preferable,
+if that choice is available.
+On the other hand, that  choice is not always available,
+as not every app can be run in a container-based environment.
+In order to maximize coverage of possible apps being able to run on our system,
+we offer container-based and virtual machine based application hosting
+as a key concept of our solution.
 
-Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host.
+As a second observation in the area of PaaS can be added
+that in most studies there is no dramatic difference
+if derivates of operating systems are compared to each other.
+Nevertheless, there are derivates
+which are more driven towards being lightweight and more performant
+~\cite{boras2020performance}  %nix vs. nix win vs. win
+~\cite{balen2020performance}. %win in vm
+In cases where a Windows virtual machine can be replaced with a natively operating wine platform,
+additional performance potentials can be leveraged
+~\cite{huang2012performance}.  % win inside vm
+In contrast to that if operating system of different types are compared to each other,
+there are more obvious results
+especially when taking into account implementations of low-level operating system functionality,
+such as interrupt handling, memory allocation, driver-specific implementations
+or other more subtle requirements such as boot time
+or the time needed for maintenance or system updates
+~\cite{sergeev2022docker} % docker, win vs nix
+~\cite{sulaiman2021comparison}. % office apps win vs nix
+
+As a result of our research we aim to spport
+any Windows or Linux based operating system in virtual machines
+but can only support Linux and Wine in containers
+due to the fact that only terminal based Windows containers are supported by Microsoft.
+
+\subsubsection{Network Latency and Bandwidth}
+Another important aspect when providing web-based access to DaaS apps
+is that also network latency and bandwidth play a very important role
+which contribute to a large degree to the responsivenes of the system
+directly experienced by the user.
+As the user utilizes its application through a web browser
+we aim to support standardized frame-buffer protocols
+to enable user interaction and access to external devices such as printers or usb sticks.
+
+Within such protocols and in cases where video content is transmitted,
+the provided network bandwidth might be the main bottleneck
+especially when it comes to high-resolution video streams.
+In cases where audio, mouse movement or keyboard strokes are transmitted,
+the network latency plays a more important role.
+Therefore, the experienced responsivness and usability
+crucially depends on the hosted application.
+
+In general, such application specific requirements can to a certain degree be controlled
+by protocol-inherent tuning parameters and utilized compression algorithms.
+Therefore, our solution aims to maximize coverage of potential usecases
+by offering several solutions and a set of appropriate tuning options
+which intentionally forces the user to decide
+which solution actually fits best for him.
+
+For that purpose we utilize the desktop gateway implementation Guacamole,
+which offers support for most utilized frame-buffer protocols VNC and RDP
+as well as all relevant tuning options.
+Guacamole not only implements out-of-the-box support for the protocols itself
+but also adds the ability to integrate audio support for VNC or other external devices
+as well as a javascript based webclient.
+
+Nevertheless, in order to integrate guacamole into the system,
+there are still things to consider which are not handled by guacamole
+but would be assumed from a full fledged daas solution.
+
+First and foremost must be considered how authentication is handled.
+On one hand, from a daas system a user would expect the system to work
+without forcing him to authenticate himself
+to more than one authentication system and also only once per session.
+On the other hand crucial information such as user passwords
+might in general better be kept secret to the system,
+such that users can not manipulate the provided system in an unintended way.
+A potential problem arises if taken into account
+that by default RDP requires a valid user session to the system.
+
+For that particular reason guacamole was integrated
+in a proxied approach by using proxied websockets.
+This enables to keep credentials of the generated system secret to the host system
+while still providing authorized but passwordless access to the hosted application.
+
+Further the proxy mechanism enables to support dynamic desktop resizing
+which is a default guacamole feature for RDP,
+but which is not available in VNC through guacamole.
+By intercepting websocket opcodes and listening for size requests,
+the requested resolution can then be enforced
+by directly interacting with the related container or virtual machine.
+
+Both changes to the default guacamole usecase are mandatory
+in order to comply with the requirements for the underlying daas design
+but both also impact the overall system performance and responsiveness in a negative way.
+From a network perspective the integration as proxy adds an additonal network hop
+and leads to a noticable delay in the network communication.
+The protocol level evaluation of opcodes enhances that delay additonally
+and the perceived responsiveness might suffer
+if resize requests take more time then expected.
+Overall, the introduced overhead is still acceptable
+and enables platform-independent and seamless access for all defined daas usecases.
+
+\subsubsection{Internal processing and Communication}
+Another important aspect directly related with performance and responsiveness
+is the approach in which data is internally processed, stored and communicated
+to all involved components and instances.
+This includes the metadata being collected and processed by essential components
+but also implies raw physical data being directly exchanged with object instances.
+
+In order to provide a satisfying user experience in a daas system
+certain components in a host system are required.
+This includes access to all backend services
+being used to create containers and virtual machines,
+a local or remote database, a local or remote authentication service,
+local and distributed filesystems as well as
+virtual or physical network devices.
+Further a webserver is integrated to provide all relevant api endpoints
+for the frontend communication
+as well as all websocket endpoints for the viewer communication.
+
+Generally spoken all of these components might become a potential bottleneck
+when being utilized with high throughputs
+or large amounts of concurrent applications demanding low latencies.
+This is especially the case if being used in a sequential manner
+or being served from the same hardware.
+The total time needed for a request being fully processed
+can increase drastically if delays add up to each other.
+
+On a architectural level all relevant components are therefore designed
+for being distributed to different hosting systems
+if requirements of a certain user scenario suggests it.
+Nevertheless, it might still be advisable
+to keep certain components directly on the host system
+in order to ensure low latency component access.
+
+For all components with realtime and low latency usecases it is generally advisable
+to keep them as close to the core hosting system as possble.
+This includes all data and communication
+relevant to audio, video, mouse and keyboard transmission,
+but also the relevant backend-service to host containers and virtual machines.
+That might be in the best case to host them natively on the host itself
+which enables virtual network and hardware access.
+In the worst case that might be at least connected to the same physical network
+in order to reduce network latencies as much as possbile.
+
+For all other components of overarching nature
+such as database access, authentication provider and filesystems
+it is still advisable to keep a close distance to the host
+but it also highly depending on the total amounts of users
+and the specifics of their usecases.
+In small scales it is still advisable to keep such components on the same system
+but in larger scales it preferable to externalize such functionality to dedicated systems.
 
 \subsection{Stability}
+% TODO: Johannes (1)
+% TODO: Christian (2)
+% TODO: Johannes: Beschreibung zur Position der Komponenten, Möglichkeiten der Parallelisierung und Einschränkungen
+%
+% Was ist hier relevant?
+% Verfügbarkeit des Service und der Benutzerdaten
+% Redundante Komponenten (Hardware und Software)
+% Welche Komponenten kann man sinnvoll redundant vorhalten?
+% Was sind die Vor- und Nachteile?
+% Auch Kosten?
+% Wo sehen wir aktuell in unserer Architektur Risiken für Dienstausfall und Datenverlust?
+% Features von Proxmox
+% Ceph-Dateisystem
+% Datenbank SQlite aktuell. Perspektivisch MariaDB
+% Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host. Proxmox kann sich darum kümmern?!
+% Docker haben/nicht haben / Vor- und Nachteile
+% Hier auch auf die Skalierbarkeit beziehen
+% Structure:
+%	- Stability
+%		- Redundancy
+%		- Scalability
+%	- Backend-services
+%		- Database
+%		- Authentication
+%		- Distributed Filesystem
+%		- Container/VM (Proxmox, Kubernetes?) 
+%		- Instance error => Recreate
+%	- Considerations:
+%		- Divide and conqueror
+%		- Requirement: Distributable components
+%	- Risks:
+%		- Filesystem error
+%		- Database error
+%		- Authentication error
+%		- Infrastructure error (Container/VM)
+%	- Benefits/Drawbacks:
+%		- Distribution of components is distribution of load => Performance
+%		- Eliminates bottlenecks
+%		- Errors affect only errornous component
 
-TODO: Johannes (1)
-TODO: Christian (2)
+Stability in Desktop as a Service (DaaS) applications is of major importance
+to ensure that our solution can meet the demands of continuous
+and time-critical operations across varied user scenarios.
+From our point of view any daas system must ensure the stability
+of all system components during runtime
+in order to maintain the ability to provide service continuity,
+handle growing user loads and to transparently recover from failures,
+while still providing a seamless end-user experience.
+Such considerations are an integral part of our architectural concept
+as they directly impact the operational integrity
+of the system and its hosted instances,
+making them suitable for enterprise-level deployment and trust.
 
-Was ist hier relevant?
+This definition of stability is underpinned by several architectural
+and operational strategies including redundancy of components, scalability of nodes,
+and a robust backend service framework
+providing means of communication between components and nodes.
+These strategies as well as their drawbacks and benefits
+will be further discussed on a component level
+and within the following subsections.
 
-Verfügbarkeit des Service und der Benutzerdaten
+\subsubsection{Scalability of Components}
+Scalability in general, but also specificaly in daas environments
+can be approached in two fundamental ways:
+horizontal and vertical scaling.
+Horizontal scaling, or scaling out, involves adding more machines or instances
+to a pool to manage increased load.
+In contrast to that, vertical scaling or scaling up,
+refers to adding more power (CPU, RAM)
+to an existing machine~\cite{vaquero2011dynamically}.
 
-Redundante Komponenten (Hardware und Software)
+While horizontal scaling is generally more flexible
+and often preferred in cloud environments
+due to its ease of integration with existing architectures,
+vertical scaling remains important for applications
+requiring strong single-thread performance or low latencies.
 
-Welche Komponenten kann man sinnvoll redundant vorhalten?
+In order to comply to both we defined means of configuration
+such that the system itself as well as all of its backend services
+are designed to be split up into logical units
+which can be distributed to dedicated systems.
+Esssentially, all backend services receive their own dedicated configuration
+and can be accessed via tuples of hostname and port.
+Any component using such a backend service can obtain the relevant tuple
+and then use it to fetch requests from that service.
+By default and in a standlone configuration
+all backend-services reside on the same native host
+and relevant tuples can be derived from the host configuration.
+In a distributed context a selection of these services
+can be distributed to a dedicated and remote system
+in order to comply to horizontal scaling principles.
+The relevant tuple can then be derived from any node in the system
+by utilizing a common database containing the relevant metainformation.
 
-Was sind die Vor- und Nachteile?
+Vertical Scaling of daas nodes is currently not planned
+but a adminstrators are free to use custom 3rd party toolchains, e.g for GPU virtualization.
+Allthough vertical scaling on a host basis
+is beyond scope of our research project,
+it is more plausbile to scale the instances
+being hosted within that system and for both dimensions.
 
-Auch Kosten?
+For vertical scaling of instances appropriate parameters and api endpoints are present
+in order to manipulate the configuration of their resource assignments,
+such that arbitrary configurations are possible.
 
-Wo sehen wir aktuell in unserer Architektur Risiken für Dienstausfall und Datenverlust?
+For horizontal scaling of instances
+the builtin mechanisms of each backend service are automatically configured and utilized
+such that instances can be arbitrily distributed across a pool of connected nodes.
+In the case of virtual machines the responsible backend service proxmox implements
+such mechanisms as part of its 'Proxmox VE High Availability (HA)' cluster.
+In the case of containers any backend service fullfilling
+similar concepts can be integrated, as for example with kubernetes.
+Within such clusters any daas instance is fully agnostic to its hosting system
+and must only be reachable via a known hostname or ip.
 
-Features von Proxmox
+\subsubsection{Redundancy of Components}
 
-Ceph-Dateisystem
+Redundancy is critical for most distributed systems
+but especially for daas architectures,
+and serves as a safeguard against service interruptions and data loss
+by duplicating essential components, functions and data.
+This design principle ensures that systems remain highly available and reliable,
+even during component failures.
+In the following a list of components eligable for redundancy is given
+which briefly describes the most relevant strategies contained in our architecture.
 
-Datenbank SQlite aktuell. Perspektivisch MariaDB
+\begin{itemize}
+	\item \textbf{Frontend/Backend: } In our final version
+	      redundancy of the frontend and backend components will offer means of configuration
+	      such that multiple instances can be distributed to dedicated systems
+	      and can be target of standardized load balancing strategies.
+	      By doing so the frontend or the backend may just act as a gateway
+	      between the user and the backend service containing the relevant instances.
+	      In that manner redundancy is achieved and frontend or backend
+	      can forward requests and responses
+	      by identifying the correct communication endpoint
+	      based on url's or specific parts of them,
+	      as for example by extracting hostnames
+	      or specific identifiers being unique within the daas domain.
+	\item \textbf{Database: } The database is integrated as a backend service
+	      and as such completely unaware of surrounding components
+	      and contexts in which it is used.
+	      The database can therefore be distributed to a dedicated system if needed.
+	      Generally, any modern database (e.g. MariaDB) implements rich features
+	      to maintain specific data consistency strategies
+	      such as data replication, clustering,
+	      dedicated failover mechnisms or also RAID control.
+	      Therefore redundancies of our database system are currently not planned,
+	      but can easily be applied by users with such needs
+	      and by using database specific features and a modified database adapter.
+	\item \textbf{Authentication: } Authentication within our architecture
+	      is also implemented as a backend service
+	      and can therefore be distributed to any dedicated system,
+	      can be a target of standardized load balancing mechanisms and strategies
+	      and is as well fully agnostic of surrounding components
+	      and calling contexts.
+	      The authentication system maintains
+	      its own credential registry in a seperate database,
+	      a customizable set of user and group privileges,
+	      and implements the well defined OAuth 2.0 standard
+	      to allow remote authentication.
+	      In our final version, a customizable amount of redundant systems might be configured.
+	      % TODO: Discuss: Is that true? Do we want to make Darios authenticator redundant?
+	\item \textbf{Distributed Filesystem: } Within our architecture
+	      we use the distributed filesystem ceph which is a distributed storage system
+	      providing object, block and file storage. In ceph redundancy is achieved
+	      through its self-healing and self-managing capabilities utilizing an algorithm
+	      called CRUSH (Controlled Replication Under Scalable Hashing).
+	      Data availability and fault-tolerance is obtained
+	      by automatically managing replication
+	      of data pieces across different devices.
+	      Our designed architecture uses these capabilities
+	      to reliably provide
+	      file storage to host systems and all hosted daas instances
+	      and provides block storage to the infrastructural backend services.
+	      By hosting daas instances directly from the distributed block storage,
+	      arbitrary redundancy scenarios might be covered
+	      by using specific configuration parameters.
+	\item \textbf{Instances: } Redundancy of instances
+	      is not a necessary condition for daas systems in general
+	      but it might make sense to apply such strategies under certain conditions.
+	      By using highly available clusters within our backend services
+	      redundancy can simply be achieved by using the corresponding api endpoints
+	      previously described and by parameterizing them appropriately.
+\end{itemize}
 
-Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host. Proxmox kann sich darum kümmern?!
+\subsubsection{Internal Communication}
+After discussing the principles being used to achieve scalability and redundancy
+it adheres to describe how communication flows are maintained
+in such a distributed enironment.
 
-Docker haben/nicht haben / Vor- und Nachteile
+Our proposed solution is at the core designed
+as a message oriented architecture
+which additionally aims to fullfill daas properties
+for scalability and redundancy as previously described.
 
-Hier auch auf die Skalierbarkeit beziehen
+That implies fundamentally, that any action being triggered,
+may it either be in the frontend on the client side,
+in the backend on the host side
+or within any of the hosted instances
+must always fullfill the condition that the required information to execute that action
+must either be directly derivable from a request or a corresponding response
+or otherwise, must be immidiately available to the component executing that action.
 
-TODO Johannes: Beschreibung zur Position der Komponenten, Möglichkeiten der Parallelisierung und Einschränkungen
+We achieve that by storing all runtime information of instances in the distributed block storage,
+by storing all relevant configuration files in our distributed file storage,
+by utilizing a common database to persist all object metadata
+and by using our decentralized OAuth component to handle authentication.
+
+By doing so, any change being trigerred by a node within the daas network
+is immidiately available to any other node in the system
+which relies on that information change.
+Apart from that any additional information needed by a component
+is fully contained within the related request or response.
+
+Therefore, arbitrary distributed configurations are possible
+as long as request and response messages reach their desired target eventually
+and as long availability of all relevant backend services is guaranteed.
+
+If that is the case, each distributed daas node can typically
+just contain all time-critical components as a bare minimum
+in order to maintain an acceptable performance throughout the whole user-facing network stack.
+This would typically include the backend service hosting the relevant instances
+as well as the websocket proxy directly connected to the guacamole daemon
+and all other components may reside somewhere in the daas network.
 
@@
 %
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
+\bibliographystyle{splncs04}
+\bibliography{biblio}
 %

### 2024-04-21 paper.tex
Added introduction (section 1)

@@
 
-Inkl. Motivation
+This paper discusses challenges during the development and implementation of a novel Desktop-as-a-Service (DaaS) solution that enables the deployment and usage of unmodified Linux and Windows applications in the same way as web applications and can run inside public resources but can also be deployed in a private context.
+Since all interaction with the DaaS and the applications that reside inside is done via a user's web browser, the users can use any client, no matter what hardware or host operating system it is compromised of.
+A browser is the only software component required to use the novel DaaS service DESIGN.
 
-TODO: Christian (1)
-TODO: Johannes (2)
+Undoubtedly, the development of a feature-rich DaaS system is a complex task, riddled with numerous obstacles and challenges.
+This paper meticulously addresses these challenges, categorizing them under compatibility, performance, stability, and usability.
+It not only brings these obstacles to light but also presents the solutions we devised to overcome them, along with the valuable lessons we learned in the process.
+
+This document is structured as follows.
+Section~\ref{sec:relatedworkAbschnitt} discusses related work on DaaS solutions from an academic perspective, whereas section~\ref{sec:RecommendDaaSarchitecture} describes the state-of-the-art DaaS architecture we designed and implemented.
+Section~\ref{sec:AnalysisPossibleComponents} presents the most relevant challenges we faced, analyses our options for handling these, and describes the solution we found.
+Finally, section~\ref{sec:Conclusions} discusses conclusions, and section~\ref{sec:NextSteps} includes directions for future work.
+
+% TODO: Christian (1)
+% TODO: Johannes (2)
 
 \section{Related Work}
+\label{sec:relatedworkAbschnitt}
 % TODO:  Introduce: Infratstructure-as-a-Service  and others (IaaS, PaaS, SaaS, DaaS)
@@
 \section{Architecture}
+\label{sec:RecommendDaaSarchitecture}
 
@@
 
+(see Figure~\ref{figure_architecture}).
+
+\begin{figure}
+\includegraphics[width=\textwidth]{images/DaaS_DESIGN_Architecture_v11_english.pdf}
+\caption{Architecture of the novel DaaS DESIGN.} \label{figure_architecture}
+\end{figure}
+
 \section{Challenges and Solutions}
+\label{sec:AnalysisPossibleComponents}
 
@@
 \section{Conclusion}
+\label{sec:Conclusions}
 
@@
 \section{Outlook}
+\label{sec:NextSteps}
 
@@
 
-\section*{References}
-
 % ---- Bibliography ----

### 2024-04-21 paper.tex
Added related work (section 2) and a big part of architecture (section 3)

@@
 %
-\institute{Faculty of Computer Science and Engineering, Frankfurt University of Applied Sciences,\\
+\institute{Faculty of Computer Science and Engineering,\\
+Frankfurt University of Applied Sciences,\\
 Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
@@
 % TODO:  Introduce: Infratstructure-as-a-Service  and others (IaaS, PaaS, SaaS, DaaS)
-TODO: Christian (1)
-TODO: Johannes (2)
+% TODO: Christian (1)
+% TODO: Johannes (2)
+
+DaaS has been a well-known service category since the emergence of cloud computing. Still, it got much less attention in research and literature
+The challenges and obstacles when developing DaaS solutions, rather than infrastructure (IaaS) and platform services (PaaS), have seldom been discussed in the literature.
+
+Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using OpenStack and analyzed the characteristics and performance aspects of using noVNC, SPICE, and Apache Guacamole as solutions for providing access to the desktop via a browser.
+One focus of the paper is the redirection of the sound interface of the virtual desktop.
+The paper's authors conclude that Guacamole, in combination with the protocol RDP, is the best solution.
+The authors evaluated the lag time for remote audio playback in an Internet scenario.
+by measuring the time between the automated initiation of the live streaming using the VLC media player software and the arrival of the music data by analyzing network traffic via monitoring software Wireshark.
+When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms.
+When using Guacamole combined with the protocol VNC, the average lag was around 1750\,ms.
+The paper also includes a short evaluation of the lag for video frames represented by background color changes of a Java application.
+Again, the configuration Guacamole and RDP offered the best performance video update with an average lag of around 300\,ms.
+The configuration of Guacamole and VNC caused an average lag of around 500\,ms.
+
+Magana et al.~\cite{magana2019remote} compared in 2019 the most popular remote desktop protocols and software implementations.
+The authors analyzed five protocols: PCoIP used in the Amazon WorkSpaces, Microsoft RDP, TeamViewer, VNC, and Citrix Independent Computing Architecture (ICA).
+In the paper, the network transfer rate and its relation to the quality experienced by the DaaS user are evaluated by assuming three scenarios: using an office software suite, web browsing, and video streaming.
+One emphasis of the paper is comparing the transfer rate of the different protocols for three scenarios.
+In the Office Software Suite scenario, the transfer rate is measured and compared when performing several different tasks like opening and editing a document (typing), loading an image, and saving a document.
+The web browsing scenario covers requesting three different web pages, and for the video streaming scenario, the same YouTube video file was viewed at different resolutions.
+The authors conclude that RDP requires fewer downstream and upstream network resources for all three scenarios.
+It is important to understand that the protocols RDP and VNC are implemented in open-source projects and proprietary solutions for many different operating systems.
+Depending on the different quality levels of the many existing implementations, it is difficult to generalize about the performance of the protocols, as individual implementations may have better or worse performance characteristics than expected.
+
+Several works in literature have analyzed the latency of Cloud gaming services that have, in some aspects, similar requirements to DaaS offerings.
+For both cloud service categories, users expect a latency value that is not considered annoying when interacting with the service and its applications.
+Choy et al.~\cite{ChoyWongSimonRosenberg2012}, Claypool et al.~\cite{claypool2010latency}, and Jarschel et al.~\cite{jarschel2011evaluation} consider a maximum tolerable latency, based on subjective tests, being around 100\,ms.
+This relatively low value sets the demand for a fast network interaction between client and service and limits their maximum distance.
 
@@
 \label{sec:RecommendDaaSarchitecture}
+% 
+% TODO: Christian (1)
+% TODO: Johannes (2)
 
-TODO: Christian (1)
-TODO: Johannes (2)
+Figure~\ref{figure_architecture} shows the architecture of our novel DaaS solution DESIGN that can run unmodified Linux and Windows applications in Linux containers (Linux and Windows applications compatible with Wine) and virtual machines (Windows applications incompatible with Wine).
 
-(see Figure~\ref{figure_architecture}).
+The open-source server virtualization platform Proxmox VE runs bare-metal and offers virtual machines with the Kernel-based Virtual Machine (KVM) and container-based virtualization via LXC Linux container.
+Since the API of LXC is, in comparison to the Docker API, rather poorly documented and lacks several features, we considered Docker a better container virtualization solution.
+A deployment of docker directly in the host-operating system or inside a virtual machine is both possible.
+
+Exporting the graphical user interface of the Linux and Windows applications,
+is possible using the a protocols VNC or RDP.
+Free server implementations exist for both operating system families.
+Since the focus of our DaaS solution is to export only the application's graphical user interface and not full desktops -- a feature not all implementations include -- is the number of available projects and solutions limited.
+For Linux containers, the free software projects x11vnc and Xrdp can be used in the DESIGN DaaS. For Windows VMs, the free software project TightVNC and the Windows-internal RDP server can be used.
+
+One of the main goals of DESIGN is to allow all interactions to be carried out solely by using a browser, a broker or proxy software that mediates between the VNC or RDP implementation and the browser is required.
+The most advanced solution available is the free software project Apache Guacamole because it supports VNC and RDP both and includes further relevant features regarding sound and printing.
+
+The core component of DESIGN is our self-developed Web UNI and Control Application, that .... TODO!!!
 

### 2024-04-22 paper.tex
Small fix
Some small fixes

@@
 %
-\author{Christian Baun\orcidID{0009-0004-9955-3752} \and
-	Johannes Bouché\orcidID{TODO}}
+\author{Christian Baun \and 
+Johannes Bouché}
+% \author{Christian Baun\orcidID{0009-0004-9955-3752} \and
+% 	Johannes Bouché\orcidID{TODO}}
 %
@@
 
-The core component of DESIGN is our self-developed Web UNI and Control Application, that .... TODO!!!
+The core component of DESIGN is our self-developed Web UI and the Control Application, that carries out the communication between all components of the service.
+
+When starting an instance (virtual machine or container), they include a SSH service automatically.
+Our self-developed instance component waits until the SSH service is ready to satisfy requests in case applications are to be started or modifications are to be carried out, the instance component sends these requests to the virtual machine or container.
+Modification requests include resolution changes and fetching host status information.
+
+In general, die self-developed instance component allows to carry out essential administration and usage tasks the operating systems does not offer via a generalized API.
+
+TODO: Die letzten 3 Abschnitte noch überarbeiten und präzisieren
 

### 2024-04-22 paper.tex
Some small additions in related work

@@
 This relatively low value sets the demand for a fast network interaction between client and service and limits their maximum distance.
+As mentioned in Clincy et al.~\cite{clincy2013subjective}, The former cloud gaming service provider OnLive specified in the year 2010 that the distance between service provider and consumer should not be greater than 1000 miles (approx.
+1600\,km) because otherwise the round-trip communications delay time between will be too long for video games.
 

### 2024-04-22 paper.tex
Some small fixes

@@
 
-The core component of DESIGN is our self-developed Web UI and the Control Application, that carries out the communication between all components of the service.
+The core components of DESIGN are our self-developed Web UI and the Control Application, which carry out the communication between all service components.
 
-When starting an instance (virtual machine or container), they include a SSH service automatically.
-Our self-developed instance component waits until the SSH service is ready to satisfy requests in case applications are to be started or modifications are to be carried out, the instance component sends these requests to the virtual machine or container.
+When starting an instance (virtual machine or container), it automatically includes an SSH service.
+Our self-developed instance component waits until the SSH service is ready to satisfy requests.
+In case applications are to be started or modifications are to be carried out, the instance component sends these requests to the virtual machine or container.
 Modification requests include resolution changes and fetching host status information.
 
-In general, die self-developed instance component allows to carry out essential administration and usage tasks the operating systems does not offer via a generalized API.
+In general, the self-developed instance component allows users to carry out essential administration and usage tasks that the operating system does not offer via a generalized API.
 
-TODO: Die letzten 3 Abschnitte noch überarbeiten und präzisieren
+% TODO: Die letzten 3 Abschnitte noch überarbeiten und präzisieren
 

### 2024-04-22 paper.tex
Some small fixes

@@
 
-TODO: Christian (1)
-TODO: Johannes (2)
+% TODO: Christian (1)
+% TODO: Johannes (2)
+
+While developing the novel DaaS platform, we faced multiple obstacles and challenges.
+Each one these issues belong to one the four categories compatibility, performance, stability and usability.
+This section presents the issues we were confronted with and, the possible approaches we analyzed and the solutions we chose.
 

### 2024-04-22 paper.tex
Some content about compatibility

--- 2024-04-22 paper.tex
@@
 
-TODO: Christian (1)
-TODO: Johannes (2)
+% TODO: Christian (1)
+% TODO: Johannes (2)
+
+While developing the novel DaaS platform, we faced multiple obstacles and challenges.
+Each of these issues belongs to one of the four categories: compatibility, performance, stability, and usability.
+This section presents the issues we were confronted with, the possible approaches we analyzed, and the solutions we chose.
 
 \subsection{Compatibility}
+% 
+% TODO: Christian (1)
+% TODO: Johannes (2)
 
-TODO: Christian (1)
-TODO: Johannes (2)
+The first relevant characteristic of a DaaS that a user is confronted with is the list of compatible applications.
+In the best case, all applications of all popular operating systems can be integrated and used with such a novel platform.
+Several DaaS projects and solutions from the last two decades, as well as similar concepts like web desktops or webtops, lacked native support for popular everyday applications.
+
+Examples worth mentioning in this context is eyeOS (2005)~\cite{liu2012research}~\cite{vidyabanu2011implementation}, which simulated a desktop environment using HTML, PHP, JavaScript, and MySQL.
+Despite being open source, it has never become widely used since it is impossible to import and use native Windows or Linux applications with this software.
+All required applications have to be developed as web applications.
+
+Kommerzielle Anwendungen...
 

### 2024-04-23 paper.tex
More content for section 4.1 (compatibility)

@@
 In the best case, all applications of all popular operating systems can be integrated and used with such a novel platform.
-Several DaaS projects and solutions from the last two decades, as well as similar concepts like web desktops or webtops, lacked native support for popular everyday applications.
+Several DaaS projects and solutions from the last two decades and similar concepts like web desktops or webtops lacked native support for popular everyday applications.
 
@@
 All required applications have to be developed as web applications.
+This limitation especially caused most potential users to lose interest in web desktops because many popular native applications can hardly be replaced or re-developed, and many feature-rich and well-established Software-as-a-Service offerings like Microsoft Office 365 and Google Docs already exist.
 
-Kommerzielle Anwendungen...
-
-Anwendungen (Windows / Linux / MacOS)
+The development of DaaS solutions in the past shows that a new DaaS will only be accepted by a broad user group if it enables the integration of native applications in an unchanged manner.
+It is also desirable that a DaaS supports applications for the Linux and Windows operating systems to enable flexible working and make the operation of a virtualization solution on the desktops unnecessary.
+Ideally, it should also be possible to run native MacOS applications.
+However, this goal is difficult to achieve under legal conditions due to the licensing restrictions of the MacOS operating system.
 

### 2024-04-23 paper.tex
More content for section 4.1 (compatibility)

@@
 
-Examples worth mentioning in this context is eyeOS (2005)~\cite{liu2012research}~\cite{vidyabanu2011implementation}, which simulated a desktop environment using HTML, PHP, JavaScript, and MySQL.
-Despite being open source, it has never become widely used since it is impossible to import and use native Windows or Linux applications with this software.
+Examples worth mentioning in this context are eyeOS and SilveOS.
+The project eyeOS~\cite{liu2012research}~\cite{vidyabanu2011implementation} which simulated a desktop environment using HTML, PHP, JavaScript, and MySQL.
+The project SilveOS~\cite{garmpis2016design} mimics the look and feel of Windows 10 in the browser using the discontinued application framework Microsoft Silverlight.
+Despite being open source, both solutions have never become widely used since it is impossible to import and use native Windows or Linux applications with such solutions.
 All required applications have to be developed as web applications.
@@
 It is also desirable that a DaaS supports applications for the Linux and Windows operating systems to enable flexible working and make the operation of a virtualization solution on the desktops unnecessary.
-Ideally, it should also be possible to run native MacOS applications.
-However, this goal is difficult to achieve under legal conditions due to the licensing restrictions of the MacOS operating system.
+Ideally, it should also be possible to run native Mac OS applications.
+However, this goal is difficult to achieve under legal conditions due to the Mac OS operating system's licensing restrictions.
 
-Sound (via sound-deamon oder via Guacamole)
+Further essential features for most users are printing, sound, and compatibility with external storage media.
 
-Drucker (via cups oder via Guacamole)
+% !!!!!!
+% Vermutlich ist das einfach gar nicht relevant hier... wenn doch, müssen wir es sinnvoll einbauen.
+% !!!!!!
+% On Linux operating systems, sound and microphone devices are accessed via sound servers like PulseAudio and PipeWire, which interact with the sound interface present in the kernel, e.g., Advanced Linux Sound Architecture (ALSA) or Open Sound System (OSS).
+...
 
-Externe Geräte (USB-Stick)
+Guacamole supports redirecting audio, printing, and disk access via the protocol RDP.
+The printing feature does not allow access to local or remote printers; instead, it allows users to print directly to PDF files, which are received by the user's web browser.
+As a prerequisite, the GhostScript interpreter must be deployed on the Guacamole server.
+File transfer over RDP supports Guacamole by emulating a virtual disk drive that persistently stores data on the Guacamole server.~\cite{ApacheGuacamole2024}
+
+Compared to RDP, the protocol VNC offers fewer features in Guacamole since it does not support sound on its own.
+However, Guacamole supports sound using PulseAudio as a workaround solution.
+VNC also does not offer file transfer and access to external storage media, but as a workaround solution, Guacamole supports file transfer file transfer over SSH using the protocol SFTP.
 

### 2024-04-24 paper.tex
More content for section Related Work

@@
 Several works in literature have analyzed the latency of Cloud gaming services that have, in some aspects, similar requirements to DaaS offerings.
+According to Lampe et al.~\cite{lampe2014assessing}, the latency of locally executed games is compromised of the following components:
+
+\begin{itemize}
+\item Input lag caused by the controller device (e.g., mouse or keyboard).
+\item CPU time for processing the user input and the program itself.
+\item GPU time of the graphics processor for rendering the next frame.
+\item Time required for delivering the frame into the frame buffer.
+\item LCD response time (time required to display the frame on the screen).
+\end{itemize}
+
+In cloud gaming scenarios, the following components increase latency further:
+
+\begin{itemize}
+\item Transfer time (upstream) for sending the user input to the service provider.
+\item Time required for capturing the frames and encoding it as a video stream.
+\item Transfer time (downstream) for sending the video stream to the client.
+\item Time required for decoding the video stream back into a frame.
+\end{itemize}
+
 For both cloud service categories, users expect a latency value that is not considered annoying when interacting with the service and its applications.
@@
 1600\,km) because otherwise the round-trip communications delay time between will be too long for video games.
+
+On first thought, the latency requirements of a modern computer game are hardly the same as those of a desktop application.
+However, just as with computer games, the lagging operation of desktop programs has a negative impact on the user experience, and with video conferencing or other graphics-intensive applications, the utilization of resources and latency requirements are very similar to computer games.
 

### 2024-04-26 paper.tex
More content (e.g. in subsection 4.4)

@@
 \includegraphics[width=\textwidth]{images/DaaS_DESIGN_Architecture_v11_english.pdf}
-\caption{Architecture of the novel DaaS DESIGN.} \label{figure_architecture}
+\caption{Architecture of the novel DaaS DESIGN~\cite{OJCC_2023v8i1n01_Baun}.} \label{figure_architecture}
 \end{figure}
@@
 Examples worth mentioning in this context are eyeOS and SilveOS.
-The project eyeOS~\cite{liu2012research}~\cite{vidyabanu2011implementation} which simulated a desktop environment using HTML, PHP, JavaScript, and MySQL.
+The project eyeOS~\cite{liu2012research,vidyabanu2011implementation} which simulated a desktop environment using HTML, PHP, JavaScript, and MySQL.
 The project SilveOS~\cite{garmpis2016design} mimics the look and feel of Windows 10 in the browser using the discontinued application framework Microsoft Silverlight.
@@
 if a identical program was tested in both environments
-~\cite{felter2015updated}
-~\cite{potdar2020performance}
-~\cite{seo2014performance}.
+~\cite{felter2015updated,potdar2020performance,seo2014performance}.
 
@@
 which are more driven towards being lightweight and more performant
-~\cite{boras2020performance}  %nix vs. nix win vs. win
-~\cite{balen2020performance}. %win in vm
+~\cite{balen2020performance,boras2020performance}.  %nix vs. nix win vs. win
+% ~\cite{balen2020performance}. %win in vm
 In cases where a Windows virtual machine can be replaced with a natively operating wine platform,
@@
 or the time needed for maintenance or system updates
-~\cite{sergeev2022docker} % docker, win vs nix
-~\cite{sulaiman2021comparison}. % office apps win vs nix
+~\cite{sergeev2022docker,sulaiman2021comparison}. % docker, win vs nix
+% ~\cite{sulaiman2021comparison}. % office apps win vs nix
 
@@
 
-TODO: Christian (1)
-TODO: Johannes (2)
+% TODO: Christian (1)
+% TODO: Johannes (2)
 
-Ideen und Konzepte zur Trennung von User und Admin-Ansicht
+An essential feature of a DaaS platform is its strong user-friendliness, which eliminates the complexity of the different operating systems used inside virtual machines and containers, makes the underlying hardware transparent, and simplifies the necessary interactions as much as possible.
 
-Reduktion der Komplexität der Bedienung durch Fokus auf notwendigste Interaktionsschritte
+Since users already have an operating system on their clients, and the motivation for using a DaaS is the simple use of various applications, it makes no sense for most users to have additional desktops available in the browser.
+The focus should be on the applications.
+For this reason, when developing a new DaaS solution, the main goal should be to reduce the graphical output to the limits of the respective application.
+The concept uses a separate browser tag for each application whose content exclusively represents the application.
 
-Desktop vs. einzelne Anwendung im Browser-Tab
+Developing the automatic adaptation of the output to the limits of the respective application window is challenging.
+Only a few VNC and RDP servers make exporting the output of only individual windows or processes possible~\cite{OJCC_2023v8i1n01_Baun}.
+For Windows operating systems, the proprietary built-in RDP server and the free software TightVNC offer the export of full desktops and single windows.
+For Linux operating system deployments, the free software projects xrdp and x11vnc enable all desired features (see Table~\ref{tab:DaaS_Services_Overview}).
 
-Anpassung des Browser-Inhalts auf die Größe der Anwendung
+\begin{table}
+	\centering
+	\caption{Services used by our DaaS for Remote Access to individual Linux or Windows Applications}
+	\begin{tabular}{|l|c|c|c|c|}
+		\hline
+		\textbf{Service}    &  \textbf{Protocol}  & \textbf{Export of}           & \textbf{Operating} & \textbf{Software}  \\
+		\textbf{implementation}  &                & \textbf{single Window only}  & \textbf{System}    & \textbf{License}   \\
+		\hline
+		TightVNC            &  VNC                & supported                    & Windows            & GPL2               \\
+		\hline
+		Windows RDP server  &  RDP                & supported                    & Windows            & proprietary        \\
+		\hline
+		x11vnc              &  VNC                & supported                    & Linux              & GPL2               \\
+		\hline
+		Xrdp                &  RDP                & supported                    & Linux              & Apache 2.0         \\
+		\hline
+	\end{tabular}
+	\label{tab:DaaS_Services_Overview}
+\end{table}
 
-Herausforderungen, Entwicklung, Implementierung, Lessons-learned
+The automatic adaptation of browser tabs content size, is done by.... TOTO: Johannes...
+
+TODO: Ideen und Konzepte zur Trennung von User und Admin-Ansicht
+
+TODO: Herausforderungen, Entwicklung, Implementierung, Lessons-learned
 

### 2024-04-26 paper.tex
More content (e.g. in subsection 4.4)

@@
 Developing the automatic adaptation of the output to the limits of the respective application window is challenging.
-Only a few VNC and RDP servers make exporting the output of only individual windows or processes possible~\cite{OJCC_2023v8i1n01_Baun}.
+Only a few RDP and VNC server implementations offer to export the graphical output of only individual windows or processes.
 For Windows operating systems, the proprietary built-in RDP server and the free software TightVNC offer the export of full desktops and single windows.
-For Linux operating system deployments, the free software projects xrdp and x11vnc enable all desired features (see Table~\ref{tab:DaaS_Services_Overview}).
+For Linux operating system deployments, the free software projects xrdp and x11vnc enable all desired features.
+Table ~\ref{tab:DaaS_Services_Overview} includes an overview of the mentioned service implementations for RDP and VNC protocols and their relevant characteristics.
+Further implementations like TigerVNC and protocols like NoMachine NX exist but do not seem useful in a novem DaaS design and implementation because they do not meet all feature requirements.~\cite{OJCC_2023v8i1n01_Baun}
 
@@
 		\hline
-		\textbf{Service}    &  \textbf{Protocol}  & \textbf{Export of}           & \textbf{Operating} & \textbf{Software}  \\
-		\textbf{implementation}  &                & \textbf{single Window only}  & \textbf{System}    & \textbf{License}   \\
+		\textbf{Service}    &  \textbf{Protocol}  & \textbf{Single Window}       & \textbf{Operating} & \textbf{Software}  \\
+		\textbf{implementation}  &                & \textbf{Mode}                & \textbf{System}    & \textbf{License}   \\
 		\hline

### 2024-04-26 paper.tex
More content (e.g. Abstract)

@@
 \begin{abstract}
-	The paper will cover very novel and non-published knowledge we gained
-	from developing and implementing a DaaS that is funded by the German
-	Federal Ministry for Economic Affairs and Climate Action. The topic
-	fits very well into the topics of the conference. TODO: Christian
+Desktop-as-a-Service (DaaS) is a cloud service category with great potential for improving the everyday work of running and using applications.
+However, existing DaaS solutions have several drawbacks, which have limited the popularity of this service category in the past.
+Our novel DaaS solution not only overcomes the limitations of existing solutions but also offers significant benefits.
+It has the potential to revolutionize the way we use applications, supporting non-modified Linux and Windows applications with just a browser on the client side.
+This paper describes some of the knowledge we gained during the development because of the challenges we faced and the solutions we found.
 
@@
 Since all interaction with the DaaS and the applications that reside inside is done via a user's web browser, the users can use any client, no matter what hardware or host operating system it is compromised of.
-A browser is the only software component required to use the novel DaaS service DESIGN.
+A browser is the only software component required to use the novel DaaS service.
 
@@
 
-Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using OpenStack and analyzed the characteristics and performance aspects of using noVNC, SPICE, and Apache Guacamole as solutions for providing access to the desktop via a browser.
+Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using OpenStack~\cite{sefraoui2012openstack} and analyzed the characteristics and performance aspects of using noVNC, SPICE, and Apache Guacamole as solutions for providing access to the desktop via a browser.
 One focus of the paper is the redirection of the sound interface of the virtual desktop.
-The paper's authors conclude that Guacamole, in combination with the protocol RDP, is the best solution.
+The paper's authors conclude that Guacamole, in combination with the Remote Desktop Protocol (RDP), is the best solution.
 The authors evaluated the lag time for remote audio playback in an Internet scenario.
-by measuring the time between the automated initiation of the live streaming using the VLC media player software and the arrival of the music data by analyzing network traffic via monitoring software Wireshark.
+by measuring the time between the automated initiation of the live streaming using the media player software VLC and the arrival of the music data by analyzing network traffic via monitoring software Wireshark.
 When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms.
-When using Guacamole combined with the protocol VNC, the average lag was around 1750\,ms.
+When using Guacamole combined with the protocol Virtual Network Computing (VNC), the average lag was around 1750\,ms.
 The paper also includes a short evaluation of the lag for video frames represented by background color changes of a Java application.
@@
 Magana et al.~\cite{magana2019remote} compared in 2019 the most popular remote desktop protocols and software implementations.
-The authors analyzed five protocols: PCoIP used in the Amazon WorkSpaces, Microsoft RDP, TeamViewer, VNC, and Citrix Independent Computing Architecture (ICA).
+The authors analyzed five protocols: PCoIP used in the Amazon WorkSpaces, RDP, TeamViewer, VNC~\cite{rfc6143}, and Citrix Independent Computing Architecture (ICA).
 In the paper, the network transfer rate and its relation to the quality experienced by the DaaS user are evaluated by assuming three scenarios: using an office software suite, web browsing, and video streaming.
@@
 Table ~\ref{tab:DaaS_Services_Overview} includes an overview of the mentioned service implementations for RDP and VNC protocols and their relevant characteristics.
-Further implementations like TigerVNC and protocols like NoMachine NX exist but do not seem useful in a novem DaaS design and implementation because they do not meet all feature requirements.~\cite{OJCC_2023v8i1n01_Baun}
+Further implementations like TigerVNC and protocols like NX NoMachine exist but do not seem useful for designing and implementing a novel DaaS because they do not meet all feature requirements.~\cite{OJCC_2023v8i1n01_Baun}
 
@@
 
-The automatic adaptation of browser tabs content size, is done by.... TOTO: Johannes...
+The automatic adaptation of browser tabs content size, is done by.... TOTO: Johannes´...
 

### 2024-04-26 paper.tex
Some improvements

@@
 However, existing DaaS solutions have several drawbacks, which have limited the popularity of this service category in the past.
-Our novel DaaS solution not only overcomes the limitations of existing solutions but also offers significant benefits.
+Our novel DaaS solution called DESIGN not only overcomes the limitations of existing solutions but also offers significant benefits.
 It has the potential to revolutionize the way we use applications, supporting non-modified Linux and Windows applications with just a browser on the client side.
@@
 
-This paper discusses challenges during the development and implementation of a novel Desktop-as-a-Service (DaaS) solution that enables the deployment and usage of unmodified Linux and Windows applications in the same way as web applications and can run inside public resources but can also be deployed in a private context.
-Since all interaction with the DaaS and the applications that reside inside is done via a user's web browser, the users can use any client, no matter what hardware or host operating system it is compromised of.
+This paper discusses challenges during the development and implementation of a novel Desktop-as-a-Service (DaaS) solution called DESIGN that enables the deployment and usage of unmodified Linux and Windows applications in the same way as web applications and can run inside public resources but can also be deployed in a private context.
+Since all interaction with DESIGN and the applications that reside inside is done via a user's web browser, the users can use any client, no matter what hardware or host operating system it is compromised of.
 A browser is the only software component required to use the novel DaaS service.
@@
 This document is structured as follows.
-Section~\ref{sec:relatedworkAbschnitt} discusses related work on DaaS solutions from an academic perspective, whereas section~\ref{sec:RecommendDaaSarchitecture} describes the state-of-the-art DaaS architecture we designed and implemented.
+Section~\ref{sec:relatedworkAbschnitt} discusses related work on DaaS solutions from an academic perspective, whereas section~\ref{sec:RecommendDaaSarchitecture} describes the state-of-the-art DaaS architecture DESIGN we designed and implemented.
 Section~\ref{sec:AnalysisPossibleComponents} presents the most relevant challenges we faced, analyses our options for handling these, and describes the solution we found.
@@
 
-Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using OpenStack~\cite{sefraoui2012openstack} and analyzed the characteristics and performance aspects of using noVNC, SPICE, and Apache Guacamole as solutions for providing access to the desktop via a browser.
+Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using the IaaS solution OpenStack~\cite{sefraoui2012openstack} and analyzed the characteristics and performance aspects of using noVNC, which is a client for the protocol Virtual Network Computing (VNC) that is implemented in the Hypertext Markup Language (HTML) and JavaScript, the protocol SPICE (Simple Protocol for Independent Computing Environments), and the remote desktop gateway Apache Guacamole as solutions for providing access to the desktop via a browser.
 One focus of the paper is the redirection of the sound interface of the virtual desktop.
@@
 When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms.
-When using Guacamole combined with the protocol Virtual Network Computing (VNC), the average lag was around 1750\,ms.
+When using Guacamole combined with the protocol VNC, the average lag was around 1750\,ms.
 The paper also includes a short evaluation of the lag for video frames represented by background color changes of a Java application.
@@
 Magana et al.~\cite{magana2019remote} compared in 2019 the most popular remote desktop protocols and software implementations.
-The authors analyzed five protocols: PCoIP used in the Amazon WorkSpaces, RDP, TeamViewer, VNC~\cite{rfc6143}, and Citrix Independent Computing Architecture (ICA).
+The authors analyzed five protocols: PCoIP used in the Amazon WorkSpaces, RDP, TeamViewer, VNC, and Citrix Independent Computing Architecture (ICA).
 In the paper, the network transfer rate and its relation to the quality experienced by the DaaS user are evaluated by assuming three scenarios: using an office software suite, web browsing, and video streaming.
@@
 
-For both cloud service categories, users expect a latency value that is not considered annoying when interacting with the service and its applications.
+For every deployment scenario or cloud service category possible, users expect a latency value that is not annoying when interacting with the service and its applications.
 Choy et al.~\cite{ChoyWongSimonRosenberg2012}, Claypool et al.~\cite{claypool2010latency}, and Jarschel et al.~\cite{jarschel2011evaluation} consider a maximum tolerable latency, based on subjective tests, being around 100\,ms.
 This relatively low value sets the demand for a fast network interaction between client and service and limits their maximum distance.
-As mentioned in Clincy et al.~\cite{clincy2013subjective}, The former cloud gaming service provider OnLive specified in the year 2010 that the distance between service provider and consumer should not be greater than 1000 miles (approx.
-1600\,km) because otherwise the round-trip communications delay time between will be too long for video games.
+As mentioned in Clincy et al.~\cite{clincy2013subjective}, The former cloud gaming service provider OnLive specified in the year 2010 that the distance between the service provider and the consumer should not be greater than 1000 miles (approx.
+1600\,km) because otherwise, the round-trip communications delay time between will be too long for video games.
 
@@
 The open-source server virtualization platform Proxmox VE runs bare-metal and offers virtual machines with the Kernel-based Virtual Machine (KVM) and container-based virtualization via LXC Linux container.
-Since the API of LXC is, in comparison to the Docker API, rather poorly documented and lacks several features, we considered Docker a better container virtualization solution.
+Since the LXC API is, in comparison to the Docker API, rather poorly documented and lacks several features, we considered Docker a better container virtualization solution.
 A deployment of docker directly in the host-operating system or inside a virtual machine is both possible.
 
-Exporting the graphical user interface of the Linux and Windows applications,
-is possible using the a protocols VNC or RDP.
+Exporting the graphical user interface of the Linux and Windows applications, is possible using the a protocols VNC or RDP.
 Free server implementations exist for both operating system families.
 Since the focus of our DaaS solution is to export only the application's graphical user interface and not full desktops -- a feature not all implementations include -- is the number of available projects and solutions limited.
-For Linux containers, the free software projects x11vnc and Xrdp can be used in the DESIGN DaaS. For Windows VMs, the free software project TightVNC and the Windows-internal RDP server can be used.
+For Linux containers, the free software projects x11vnc and Xrdp can be used in the DESIGN DaaS.
+For Windows VMs, the free software project TightVNC and the Windows-internal RDP server can be used.
 

### 2024-04-26 paper.tex
Some improvements

@@
 
-While developing the novel DaaS platform, we faced multiple obstacles and challenges.
+While developing the novel DaaS platform DESIGN, we faced multiple obstacles and challenges.
 Each of these issues belongs to one of the four categories: compatibility, performance, stability, and usability.
@@
 Examples worth mentioning in this context are eyeOS and SilveOS.
-The project eyeOS~\cite{liu2012research,vidyabanu2011implementation} which simulated a desktop environment using HTML, PHP, JavaScript, and MySQL.
-The project SilveOS~\cite{garmpis2016design} mimics the look and feel of Windows 10 in the browser using the discontinued application framework Microsoft Silverlight.
+The project eyeOS~\cite{liu2012research,vidyabanu2011implementation} which simulated a desktop environment using HTML, the scripting language PHP, JavaScript, and the MySQL database management system.
+The project SilveOS~\cite{garmpis2016design} mimics the look and feel of a Windows desktop in the browser using the discontinued application framework Microsoft Silverlight.
 Despite being open source, both solutions have never become widely used since it is impossible to import and use native Windows or Linux applications with such solutions.
 All required applications have to be developed as web applications.
-This limitation especially caused most potential users to lose interest in web desktops because many popular native applications can hardly be replaced or re-developed, and many feature-rich and well-established Software-as-a-Service offerings like Microsoft Office 365 and Google Docs already exist.
+This limitation especially caused most potential users to lose interest in web desktops because many popular native applications can hardly be replaced or re-developed, and many feature-rich and well-established Software-as-a-Service (SaaS) offerings like Microsoft Office 365 and Google Docs already exist.
 
@@
 
-% !!!!!!
-% Vermutlich ist das einfach gar nicht relevant hier... wenn doch, müssen wir es sinnvoll einbauen.
-% !!!!!!
-% On Linux operating systems, sound and microphone devices are accessed via sound servers like PulseAudio and PipeWire, which interact with the sound interface present in the kernel, e.g., Advanced Linux Sound Architecture (ALSA) or Open Sound System (OSS).
-...
-
 Guacamole supports redirecting audio, printing, and disk access via the protocol RDP.
@@
 Compared to RDP, the protocol VNC offers fewer features in Guacamole since it does not support sound on its own.
-However, Guacamole supports sound using PulseAudio as a workaround solution.
-VNC also does not offer file transfer and access to external storage media, but as a workaround solution, Guacamole supports file transfer file transfer over SSH using the protocol SFTP.
+However, Guacamole supports sound using PulseAudio\footnote{On Linux operating systems, sound and microphone devices are accessed via sound servers like PulseAudio and PipeWire, which interact with the sound interface present in the kernel, e.g., Advanced Linux Sound Architecture (ALSA) or Open Sound System (OSS)} as a workaround solution.
+VNC also does not offer file transfer and access to external storage media, but as a workaround solution, Guacamole supports file transfer file transfer via the SSH File Transfer Protocol (SFTP) which is an extension of the Secure Shell protocol (SSH).
 
@@
 	\item Infrastructure and Platform
-	\item Network Latency
+	\item Network Latency and Bandwidth
 	\item Internal Processing and Communication

### 2024-04-26 paper.tex
Many language improvemens

@@
 
-Stability in Desktop as a Service (DaaS) applications is of major importance
+Stability in Desktop as a Service (DaaS) applications is of significant importance
 to ensure that our solution can meet the demands of continuous
 and time-critical operations across varied user scenarios.
-From our point of view any daas system must ensure the stability
+From our point of view, any DaaS system must ensure the stability
 of all system components during runtime
-in order to maintain the ability to provide service continuity,
-handle growing user loads and to transparently recover from failures,
+to maintain the ability to provide service continuity,
+handle growing user loads and transparently recover from failures,
 while still providing a seamless end-user experience.
@@
 This definition of stability is underpinned by several architectural
-and operational strategies including redundancy of components, scalability of nodes,
+and operational strategies, including redundancy of components, scalability of nodes,
 and a robust backend service framework
 providing means of communication between components and nodes.
-These strategies as well as their drawbacks and benefits
+These strategies, as well as their drawbacks and benefits
 will be further discussed on a component level
@@
 \subsubsection{Scalability of Components}
-Scalability in general, but also specificaly in daas environments
+
+Scalability in general, but also specifically in DaaS environments
 can be approached in two fundamental ways:
-horizontal and vertical scaling.
+Horizontal and vertical scaling.
 Horizontal scaling, or scaling out, involves adding more machines or instances
-to a pool to manage increased load.
+to a pool to manage the increased load.
 In contrast to that, vertical scaling or scaling up,
@@
 
-In order to comply to both we defined means of configuration
-such that the system itself as well as all of its backend services
+To comply with both, we defined means of configuration
+such that the system itself, as well as all of its backend services
 are designed to be split up into logical units
 which can be distributed to dedicated systems.
-Esssentially, all backend services receive their own dedicated configuration
+Essentially, all backend services receive their dedicated configuration
 and can be accessed via tuples of hostname and port.
@@
 and then use it to fetch requests from that service.
-By default and in a standlone configuration
-all backend-services reside on the same native host
-and relevant tuples can be derived from the host configuration.
-In a distributed context a selection of these services
+By default and in a standalone configuration,
+all backend services reside on the same native host, and relevant tuples can be derived from the host configuration.
+In a distributed context, a selection of these services
 can be distributed to a dedicated and remote system
-in order to comply to horizontal scaling principles.
+to comply with horizontal scaling principles.
 The relevant tuple can then be derived from any node in the system
-by utilizing a common database containing the relevant metainformation.
+by utilizing a common database containing the relevant meta information.
 
-Vertical Scaling of daas nodes is currently not planned
-but a adminstrators are free to use custom 3rd party toolchains, e.g for GPU virtualization.
-Allthough vertical scaling on a host basis
-is beyond scope of our research project,
-it is more plausbile to scale the instances
+Vertical scaling of DaaS nodes is currently not planned
+but administrators are free to use custom 3rd party toolchains, e.g., for GPU virtualization.
+Although vertical scaling on a host basis
+is beyond the scope of our research project,
+it is more plausible to scale the instances
 being hosted within that system and for both dimensions.
 
-For vertical scaling of instances appropriate parameters and api endpoints are present
-in order to manipulate the configuration of their resource assignments,
+For vertical scaling of instances, appropriate parameters and API endpoints are present
+to manipulate the configuration of their resource assignments,
 such that arbitrary configurations are possible.
@@
 the builtin mechanisms of each backend service are automatically configured and utilized
-such that instances can be arbitrily distributed across a pool of connected nodes.
-In the case of virtual machines the responsible backend service proxmox implements
+such that instances can be arbitrarily distributed across a pool of connected nodes.
+In the case of virtual machines, the responsible backend service proxmox implements
 such mechanisms as part of its 'Proxmox VE High Availability (HA)' cluster.
-In the case of containers any backend service fullfilling
-similar concepts can be integrated, as for example with kubernetes.
-Within such clusters any daas instance is fully agnostic to its hosting system
-and must only be reachable via a known hostname or ip.
+In the case of containers, any backend service fulfilling
+similar concepts can be integrated, for example, with Kubernetes.
+Within such clusters, any DaaS instance is entirely agnostic to its hosting system
+and must only be reachable via a known hostname or IP.
 
@@
 Redundancy is critical for most distributed systems
-but especially for daas architectures,
+but especially for DaaS architectures,
 and serves as a safeguard against service interruptions and data loss
-by duplicating essential components, functions and data.
+by duplicating essential components, functions, and data.
 This design principle ensures that systems remain highly available and reliable,
 even during component failures.
-In the following a list of components eligable for redundancy is given
+In the following, a list of components eligible for redundancy is given
 which briefly describes the most relevant strategies contained in our architecture.
@@
 \begin{itemize}
-	\item \textbf{Frontend/Backend: } In our final version
+\item \textbf{Frontend/Backend:} In our final version
 redundancy of the frontend and backend components will offer means of configuration
 such that multiple instances can be distributed to dedicated systems
-	      and can be target of standardized load balancing strategies.
-	      By doing so the frontend or the backend may just act as a gateway
+and can be the target of standardized load-balancing strategies.
+By doing so, the frontend or the backend may act as a gateway
 between the user and the backend service containing the relevant instances.
-	      In that manner redundancy is achieved and frontend or backend
+In that manner, redundancy is achieved, and frontend or backend
 can forward requests and responses
 by identifying the correct communication endpoint
-	      based on url's or specific parts of them,
-	      as for example by extracting hostnames
-	      or specific identifiers being unique within the daas domain.
-	\item \textbf{Database: } The database is integrated as a backend service
-	      and as such completely unaware of surrounding components
+based on URL's or specific parts of them,
+for example, by extracting hostnames
+or specific identifiers being unique within the DaaS domain.
+\item \textbf{Database:} The database is integrated as a backend service
+and as such, completely unaware of surrounding components
 and contexts in which it is used.
-	      The database can therefore be distributed to a dedicated system if needed.
-	      Generally, any modern database (e.g. MariaDB) implements rich features
+The database can, therefore, be distributed to a dedicated system if needed.
+Generally, any modern database solution (e.g., MariaDB) implements rich features
 to maintain specific data consistency strategies
 such as data replication, clustering,
-	      dedicated failover mechnisms or also RAID control.
-	      Therefore redundancies of our database system are currently not planned,
+dedicated failover mechanisms or also RAID control.
+Therefore, redundancies in our database system are currently not planned,
 but can easily be applied by users with such needs
-	      and by using database specific features and a modified database adapter.
+and by using database-specific features and a modified database adapter.
 \item \textbf{Authentication: } Authentication within our architecture
 is also implemented as a backend service
-	      and can therefore be distributed to any dedicated system,
-	      can be a target of standardized load balancing mechanisms and strategies
+and can, therefore, be distributed to any dedicated system,
+can be a target of standardized load-balancing mechanisms and strategies
 and is as well fully agnostic of surrounding components
@@
 The authentication system maintains
-	      its own credential registry in a seperate database,
+its credential registry in a separate database,
 a customizable set of user and group privileges,
-	      and implements the well defined OAuth 2.0 standard
+and implements the well-defined OAuth 2.0 standard
 to allow remote authentication.
-	      In our final version, a customizable amount of redundant systems might be configured.
+A customizable amount of redundant systems might be configured in our final version.
 	      % TODO: Discuss: Is that true? Do we want to make Darios authenticator redundant?
-	\item \textbf{Distributed Filesystem: } Within our architecture
-	      we use the distributed filesystem ceph which is a distributed storage system
-	      providing object, block and file storage. In ceph redundancy is achieved
+\item \textbf{Distributed Filesystem:} Within our architecture
+we use the distributed filesystem Ceph, which is a distributed storage system
+providing object, block, and file storage. In Ceph, redundancy is achieved
 through its self-healing and self-managing capabilities utilizing an algorithm
 called CRUSH (Controlled Replication Under Scalable Hashing).
-	      Data availability and fault-tolerance is obtained
+Data availability and fault tolerance are obtained
 by automatically managing replication
@@
 to reliably provide
-	      file storage to host systems and all hosted daas instances
+file storage to host systems and all hosted DaaS instances
 and provides block storage to the infrastructural backend services.
-	      By hosting daas instances directly from the distributed block storage,
+By hosting DaaS instances directly from the distributed block storage,
 arbitrary redundancy scenarios might be covered
 by using specific configuration parameters.
-	\item \textbf{Instances: } Redundancy of instances
-	      is not a necessary condition for daas systems in general
-	      but it might make sense to apply such strategies under certain conditions.
+\item \textbf{Instances:} Redundancy of instances
+It is not a necessary condition for DaaS systems in general, but applying such strategies under certain conditions might make sense.
 By using highly available clusters within our backend services
-	      redundancy can simply be achieved by using the corresponding api endpoints
+redundancy can be achieved by using the corresponding API endpoints
 previously described and by parameterizing them appropriately.
@@
 it adheres to describe how communication flows are maintained
-in such a distributed enironment.
+in such a distributed environment.
 
-Our proposed solution is at the core designed
-as a message oriented architecture
-which additionally aims to fullfill daas properties
+Our proposed solution is, at the core, designed
+as a message-oriented architecture
+which additionally aims to fulfill DaaS properties
 for scalability and redundancy as previously described.
 
-That implies fundamentally, that any action being triggered,
+That implies fundamentally that any action being triggered,
 may it either be in the frontend on the client side,
@@
 or within any of the hosted instances
-must always fullfill the condition that the required information to execute that action
+must always fulfill the condition that the required information to execute that action
 must either be directly derivable from a request or a corresponding response
-or otherwise, must be immidiately available to the component executing that action.
+or otherwise, must be immediately available to the component executing that action.
 
@@
 
-By doing so, any change being trigerred by a node within the daas network
-is immidiately available to any other node in the system
+By doing so, any change being triggered by a node within the DaaS network
+is immediately available to any other node in the system
 which relies on that information change.
-Apart from that any additional information needed by a component
+Apart from that, any additional information needed by a component
 is fully contained within the related request or response.
 
-Therefore, arbitrary distributed configurations are possible
-as long as request and response messages reach their desired target eventually
-and as long availability of all relevant backend services is guaranteed.
+Therefore, arbitrarily distributed configurations are possible
+as long as request and response messages reach their desired target, eventually
+and as long as the availability of all relevant backend services is guaranteed.
 
-If that is the case, each distributed daas node can typically
-just contain all time-critical components as a bare minimum
-in order to maintain an acceptable performance throughout the whole user-facing network stack.
+If that is the case, each distributed DaaS node can typically
+just contain all time-critical components at a bare minimum
+to maintain an acceptable performance throughout the whole user-facing network stack.
 This would typically include the backend service hosting the relevant instances
-as well as the websocket proxy directly connected to the guacamole daemon
-and all other components may reside somewhere in the daas network.
+as well as the WebSocket proxy directly connected to the guacamole daemon
+and all other components may reside somewhere in the DaaS network.
 

### 2024-04-28 paper.tex
Many language improvemens

@@
 
-Exporting the graphical user interface of the Linux and Windows applications, is possible using the a protocols VNC or RDP.
+Exporting the graphical user interface of Linux and Windows applications, is possible using the a protocols VNC or RDP.
 Free server implementations exist for both operating system families.
-Since the focus of our DaaS solution is to export only the application's graphical user interface and not full desktops -- a feature not all implementations include -- is the number of available projects and solutions limited.
+Since the focus of our DaaS solution is to export only the application's graphical user interface and not full desktops -- a feature not all server implementations include -- is the number of available projects and solutions limited.
 For Linux containers, the free software projects x11vnc and Xrdp can be used in the DESIGN DaaS.
-For Windows VMs, the free software project TightVNC and the Windows-internal RDP server can be used.
+For Windows VMs, the free software TightVNC and the Windows-internal RDP server can be used.
 
-One of the main goals of DESIGN is to allow all interactions to be carried out solely by using a browser, a broker or proxy software that mediates between the VNC or RDP implementation and the browser is required.
-The most advanced solution available is the free software project Apache Guacamole because it supports VNC and RDP both and includes further relevant features regarding sound and printing.
+One of the main goals of DESIGN is to allow all interactions to be carried out solely by using a browser.
+Thus, a broker or proxy software that mediates between the VNC or RDP implementation and the browser is required.
+The most advanced solution available is the free software project Apache Guacamole which supports VNC and RDP both and includes further relevant features regarding sound and printing.
 
@@
 
-When starting an instance (virtual machine or container), it automatically includes an SSH service.
+When starting a virtual machine or container instance, it automatically includes an SSH service.
 Our self-developed instance component waits until the SSH service is ready to satisfy requests.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{images/DaaS_DESIGN_Architecture_v11_english.pdf}
+\centering
+\includegraphics[width=.9\textwidth]{images/DaaS_DESIGN_Architecture_v11_english.pdf}
 \caption{Architecture of the novel DaaS DESIGN~\cite{OJCC_2023v8i1n01_Baun}.} \label{figure_architecture}
@@
 While developing the novel DaaS platform DESIGN, we faced multiple obstacles and challenges.
-Each of these issues belongs to one of the four categories: compatibility, performance, stability, and usability.
+Each of these issues belongs to one of these four categories: compatibility, performance, stability, and usability.
 This section presents the issues we were confronted with, the possible approaches we analyzed, and the solutions we chose.
@@
 In the best case, all applications of all popular operating systems can be integrated and used with such a novel platform.
-Several DaaS projects and solutions from the last two decades and similar concepts like web desktops or webtops lacked native support for popular everyday applications.
+Several DaaS projects and solutions from the last two decades and similar concepts like web desktops or webtops lacked native support for popular applications.
 
-Examples worth mentioning in this context are eyeOS and SilveOS.
-The project eyeOS~\cite{liu2012research,vidyabanu2011implementation} which simulated a desktop environment using HTML, the scripting language PHP, JavaScript, and the MySQL database management system.
+Examples worth mentioning in this context include eyeOS and SilveOS.
+The project eyeOS~\cite{liu2012research,vidyabanu2011implementation} which simulates a desktop environment using HTML, the scripting language PHP, JavaScript, and the MySQL database management system.
 The project SilveOS~\cite{garmpis2016design} mimics the look and feel of a Windows desktop in the browser using the discontinued application framework Microsoft Silverlight.
-Despite being open source, both solutions have never become widely used since it is impossible to import and use native Windows or Linux applications with such solutions.
+Despite being free software, both solutions have never become widely used since it is impossible to import and use native Windows or Linux applications there.
 All required applications have to be developed as web applications.
-This limitation especially caused most potential users to lose interest in web desktops because many popular native applications can hardly be replaced or re-developed, and many feature-rich and well-established Software-as-a-Service (SaaS) offerings like Microsoft Office 365 and Google Docs already exist.
+This limitation caused most potential users to lose interest in web desktops because many popular native applications can hardly be replaced or re-developed, and many feature-rich and well-established Software-as-a-Service (SaaS) offerings like Microsoft Office 365 and Google Docs already exist.
 
-The development of DaaS solutions in the past shows that a new DaaS will only be accepted by a broad user group if it enables the integration of native applications in an unchanged manner.
-It is also desirable that a DaaS supports applications for the Linux and Windows operating systems to enable flexible working and make the operation of a virtualization solution on the desktops unnecessary.
+A new DaaS is only accepted by a broad user group if it enables the integration of native applications in an unchanged manner.
+It is also desirable that a DaaS supports Linux and Windows applications both to enable flexible working and make the operation of a virtualization solution on the desktops unnecessary.
 Ideally, it should also be possible to run native Mac OS applications.
@@
 
-In order to describe the mission-critical performance aspects of our system,
-we initially give a short overview on potential challenges,
-inherent bottlenecks as well as other problem areas.
-We do that with respect to the underlying usecase model of DaaS-Systems
+To describe the mission-critical performance aspects of our system,
+we initially give a short overview of potential challenges,
+describe bottlenecks as well as other problem areas.
+We do that concerning the underlying use case model of DaaS-Systems
 and then follow with a more thorough description
-of our own specific design decissions and particular solutions.
+of our own specific design decisions and particular solutions.
 
-From our perspective, performance in general can be seen as the most crucial system property
+From our perspective, performance, in general can be seen as the most crucial system property
 which influences a majority of all other other relevant properties,
-such as user experience, system responsivness and overall efficiency.
+such as user experience, system responsiveness, and overall efficiency.
 As a result, we had to address multifaceted
 and potentially cross-concerning challenges
-on several distinct layers within our own architecture.
+on several distinct layers within our architecture.
 
-From the core infratructure to backend services
-and user-facing elements, each component inidvidually
-has to achieve certain minimum requirements in its own domain
-and in order to contribute to the overall performance goals of the whole system.
+From the core infrastructure to backend services
+and user-facing elements, each component individually
+has to achieve specific minimum requirements in its domain
+and to contribute to the overall performance goals of the whole system.
 Hence, we outline key limiting factors that impede performance,
@@
 aimed at mitigating these challenges.
-From our point of view the three most critial performance aspects
+From our point of view, the three most critical performance aspects
 for Daas-Systems are:
@@
 \end{itemize}
-As there is no generalized measurment system available,
+As there is no generalized measurement system available,
 we ordered them by their magnitude of impact
@@
 on the obtained experiences during our research project.
-Since each of them have their own characteristics,
+Since each of them has their own characteristics,
 specific behavior and requirements, implying a potential impact
@@
 we describe them individually and in more detail
-in the following seperate subsections.
+in the following separate subsections.
 
 \subsubsection{Infrastructure and Platform}
-In general can be said, that the chosen infrastructure, platform and utilized software
-may have the most influental impact on cpu and resource consumption
+In general, it can be said that the chosen infrastructure, platform, and utilized software
+may have the most influential impact on CPU and resource consumption
 which has a dramatic impact on the overall system performance.
-As a DaaS combines all aspects of IaaS, PaaS, SaaS into one system
+As a DaaS combines all aspects of IaaS, PaaS, and SaaS into one system
 each related performance characteristic has to be taken into account.
@@
 discussing relevant performance aspects of such systems,
-the obvious advanatge in most categories
-has a lightweight container solution compared to virtual machine
-if a identical program was tested in both environments
+the obvious advantage in most categories
+has a lightweight container solution compared to the virtual machine
+if an identical program was tested in both environments
 ~\cite{felter2015updated,potdar2020performance,seo2014performance}.
 
-On one hand, that implies for our solution,
+On the one hand, that implies for our solution,
 that a container-based approach is most likely preferable,
@@
 as not every app can be run in a container-based environment.
-In order to maximize coverage of possible apps being able to run on our system,
-we offer container-based and virtual machine based application hosting
-as a key concept of our solution.
+To maximize coverage of possible apps being able to run on our system,
+we offer container-based and virtual machine-based application hosting
+as a critical concept of our solution.
 
-As a second observation in the area of PaaS can be added
-that in most studies there is no dramatic difference
+As a second observation in the area of PaaS, it can be added
+that in most studies, there is no dramatic difference
 if derivates of operating systems are compared to each other.
@@
 ~\cite{huang2012performance}.  % win inside vm
-In contrast to that if operating system of different types are compared to each other,
+In contrast to that, if the operating systems of different types are compared to each other,
 there are more obvious results
@@
 
-As a result of our research we aim to spport
-any Windows or Linux based operating system in virtual machines
+As a result of our research, we aim to support
+any Windows or Linux-based operating system in virtual machines
 but can only support Linux and Wine in containers
-due to the fact that only terminal based Windows containers are supported by Microsoft.
+This is because only terminal-based Windows containers are supported by Microsoft.
 
 \subsubsection{Network Latency and Bandwidth}
-Another important aspect when providing web-based access to DaaS apps
-is that also network latency and bandwidth play a very important role
-which contribute to a large degree to the responsivenes of the system
+Another essential aspect when providing web-based access to DaaS apps
+is that also network latency and bandwidth play a significant role
+which contributes to a large degree to the responsiveness of the system
 directly experienced by the user.
@@
 we aim to support standardized frame-buffer protocols
-to enable user interaction and access to external devices such as printers or usb sticks.
+to enable user interaction and access to external devices such as printers or USB sticks.
 
@@
 especially when it comes to high-resolution video streams.
-In cases where audio, mouse movement or keyboard strokes are transmitted,
-the network latency plays a more important role.
-Therefore, the experienced responsivness and usability
+In cases where audio, mouse movement, or keyboard strokes are transmitted,
+the network latency plays a more critical role.
+Therefore, the experienced responsibility and usability
 crucially depends on the hosted application.
 
-In general, such application specific requirements can to a certain degree be controlled
+In general, such application-specific requirements can, to a certain degree, be controlled
 by protocol-inherent tuning parameters and utilized compression algorithms.
-Therefore, our solution aims to maximize coverage of potential usecases
+Therefore, our solution aims to maximize coverage of potential use cases
 by offering several solutions and a set of appropriate tuning options
 which intentionally forces the user to decide
-which solution actually fits best for him.
+which solution fits best for him.
 
-For that purpose we utilize the desktop gateway implementation Guacamole,
-which offers support for most utilized frame-buffer protocols VNC and RDP
+For that purpose, we utilize the desktop gateway implementation Guacamole,
+which offers support for the most utilized frame-buffer protocols, VNC and RDP
 as well as all relevant tuning options.
@@
 but also adds the ability to integrate audio support for VNC or other external devices
-as well as a javascript based webclient.
+as well as a JavaScript-based web client.
 
-Nevertheless, in order to integrate guacamole into the system,
-there are still things to consider which are not handled by guacamole
-but would be assumed from a full fledged daas solution.
+Nevertheless, to integrate guacamole into the system,
+there are still things to consider that are not handled by guacamole
+but would be assumed from a full-fledged DaaS solution.
 
-First and foremost must be considered how authentication is handled.
-On one hand, from a daas system a user would expect the system to work
+First and foremost, how authentication is handled must be considered.
+On one hand, from a DaaS system, a user would expect the system to work
 without forcing him to authenticate himself
-to more than one authentication system and also only once per session.
-On the other hand crucial information such as user passwords
-might in general better be kept secret to the system,
-such that users can not manipulate the provided system in an unintended way.
-A potential problem arises if taken into account
-that by default RDP requires a valid user session to the system.
+to multiple authentication systems and only once per session.
+On the other hand, crucial information such as user passwords
+might, in general, better be kept secret from the system
+so that users can not manipulate the provided system in an unintended way.
+A potential problem arises if it is considered
+that, by default, RDP requires a valid user session to the system.
 
-For that particular reason guacamole was integrated
-in a proxied approach by using proxied websockets.
-This enables to keep credentials of the generated system secret to the host system
+For that reason, guacamole was integrated with a proxied approach using proxied websockets.
+This enables keeping the credentials of the generated system secret to the host system
 while still providing authorized but passwordless access to the hosted application.
 
-Further the proxy mechanism enables to support dynamic desktop resizing
+Further, the proxy mechanism enables the support of dynamic desktop resizing
 which is a default guacamole feature for RDP,
 but which is not available in VNC through guacamole.
-By intercepting websocket opcodes and listening for size requests,
+By intercepting WebSocket opcodes and listening for size requests,
 the requested resolution can then be enforced
@@
 
-Both changes to the default guacamole usecase are mandatory
-in order to comply with the requirements for the underlying daas design
+Both changes to the default guacamole use case are mandatory
+to comply with the requirements for the underlying DaaS design
 but both also impact the overall system performance and responsiveness in a negative way.
-From a network perspective the integration as proxy adds an additonal network hop
-and leads to a noticable delay in the network communication.
-The protocol level evaluation of opcodes enhances that delay additonally
-and the perceived responsiveness might suffer
-if resize requests take more time then expected.
+From a network perspective, the integration as a proxy adds network hop
+and leads to a noticeable delay in network communication.
+The protocol-level evaluation of opcodes additionally enhances that delay, and perceived responsiveness might suffer if resize requests take more time than expected.
 Overall, the introduced overhead is still acceptable
-and enables platform-independent and seamless access for all defined daas usecases.
+and enables platform-independent and seamless access for all defined DaaS use cases.
 
 \subsubsection{Internal processing and Communication}
-Another important aspect directly related with performance and responsiveness
-is the approach in which data is internally processed, stored and communicated
+Another critical aspect directly related to performance and responsiveness
+is the approach in which data is internally processed, stored, and communicated
 to all involved components and instances.
@@
 
-In order to provide a satisfying user experience in a daas system
-certain components in a host system are required.
+To provide a satisfying user experience in a DaaS system
+specific components in a host system are required.
 This includes access to all backend services
@@
 a local or remote database, a local or remote authentication service,
-local and distributed filesystems as well as
+local and distributed filesystems, as well as
 virtual or physical network devices.
-Further a webserver is integrated to provide all relevant api endpoints
+Further, a web server is integrated to provide all relevant API endpoints
 for the frontend communication
-as well as all websocket endpoints for the viewer communication.
+as well as all WebSocket endpoints for viewer communication.
 
-Generally spoken all of these components might become a potential bottleneck
+Generally spoken, all of these components might become a potential bottleneck
 when being utilized with high throughputs
@@
 or being served from the same hardware.
-The total time needed for a request being fully processed
-can increase drastically if delays add up to each other.
+If delays add up, the total time needed for a request to be fully processed can increase drastically.
 
-On a architectural level all relevant components are therefore designed
+On an architectural level, all relevant components are, therefore designed
 for being distributed to different hosting systems
-if requirements of a certain user scenario suggests it.
+if the requirements of a particular user scenario suggest it.
 Nevertheless, it might still be advisable
-to keep certain components directly on the host system
-in order to ensure low latency component access.
+to keep specific components directly on the host system
+to ensure low latency component access.
 
-For all components with realtime and low latency usecases it is generally advisable
-to keep them as close to the core hosting system as possble.
+For all components with real-time and low-latency latency usecases, it is generally advisable
+to keep them as close to the core hosting system as possible.
 This includes all data and communication
-relevant to audio, video, mouse and keyboard transmission,
-but also the relevant backend-service to host containers and virtual machines.
+relevant to audio, video, mouse, and keyboard transmission,
+but also the backend service pertinent to host containers and virtual machines.
 That might be in the best case to host them natively on the host itself
 which enables virtual network and hardware access.
-In the worst case that might be at least connected to the same physical network
-in order to reduce network latencies as much as possbile.
+In the worst case, that might be at least connected to the same physical network
+to reduce network latencies as much as possible.
 
-For all other components of overarching nature
-such as database access, authentication provider and filesystems
-it is still advisable to keep a close distance to the host
-but it also highly depending on the total amounts of users
-and the specifics of their usecases.
-In small scales it is still advisable to keep such components on the same system
-but in larger scales it preferable to externalize such functionality to dedicated systems.
+For all other components of the overarching nature
+such as database access, authentication provider, and filesystems
+it is still advisable to keep a close distance from the host
+but it also highly depends on the total amounts of users
+and the specifics of their use cases.
+On small scales, it is still advisable to keep such components on the same system
+but on larger scales, it is preferable to externalize such functionality to dedicated systems.
 
@@
 
-Stability in Desktop as a Service (DaaS) applications is of significant importance
+Stability in DaaS applications is of significant importance
 to ensure that our solution can meet the demands of continuous
@@
 such that instances can be arbitrarily distributed across a pool of connected nodes.
-In the case of virtual machines, the responsible backend service proxmox implements
+In the case of virtual machines, the responsible backend service Proxmox implements
 such mechanisms as part of its 'Proxmox VE High Availability (HA)' cluster.

### 2024-04-28 paper.tex
Added 5 (Conclusion) and 6 (Outlook)

@@
 
-TODO: Zusammen am Schluss
+In this paper, we describe an architecture for a novel and feature-rich DaaS that is, by principle, superior to comparable products and approaches because it supports running unmodified Linux and Windows applications.
+Furthermore, we identified characteristic issues -- concerning compatibility, performance, stability, and usability -- during the development and implementation of our DaaS and present solutions to overcome these challenges.
 
@@
 
-TODO: Zusammen am Schluss
+For future work, we plan to do a performance analysis that includes an evaluation of different user behaviors for various popular applications and realistic deployment and utilization scenarios.
 

### 2024-04-29 paper.tex
Minor Language improvemens

@@
 \title{Challenges and Solutions of Developing and
-	Implementing a Revolutionary Novel Desktop-as-a-Service}
+	Implementing a Novel Desktop-as-a-Service}
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Challenges and Solutions of Developing and
+	Implementing a Novel DaaS}
 % If the paper title is too long for the running head, you can set
@@
 \begin{abstract}
-Desktop-as-a-Service (DaaS) is a cloud service category with great potential for improving the everyday work of running and using applications.
+Desktop-as-a-Service (DaaS) is a cloud service category with a tremendous potential for improving the everyday work of running and using applications.
 However, existing DaaS solutions have several drawbacks, which have limited the popularity of this service category in the past.
@@
 	\centering
-	\caption{Services used by our DaaS for Remote Access to individual Linux or Windows Applications}
+	\caption{Services used by our DaaS for Remote Access to individual Applications}
 	\begin{tabular}{|l|c|c|c|c|}

### 2024-04-29 paper.tex
Condense chapters performance and stability

@@
 
-DaaS has been a well-known service category since the emergence of cloud computing. Still, it got much less attention in research and literature
+DaaS has been a well-known service category since the emergence of cloud computing. Still, it got much less attention in research and literature.
 The challenges and obstacles when developing DaaS solutions, rather than infrastructure (IaaS) and platform services (PaaS), have seldom been discussed in the literature.
@@
 The paper's authors conclude that Guacamole, in combination with the Remote Desktop Protocol (RDP), is the best solution.
-The authors evaluated the lag time for remote audio playback in an Internet scenario.
-by measuring the time between the automated initiation of the live streaming using the media player software VLC and the arrival of the music data by analyzing network traffic via monitoring software Wireshark.
+The authors evaluated the lag time for remote audio playback in an Internet scenario by measuring the time between the automated initiation of the live streaming using the media player software VLC and the arrival of the music data by analyzing network traffic via monitoring software Wireshark.
 When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms.
@@
 
-To describe the mission-critical performance aspects of our system,
-we initially give a short overview of potential challenges,
-describe bottlenecks as well as other problem areas.
-We do that concerning the underlying use case model of DaaS-Systems
-and then follow with a more thorough description
-of our own specific design decisions and particular solutions.
+To describe the mission-critical performance aspects of the discussed system,
+we initially provide a brief overview of potential challenges
+by outlining bottlenecks and other relevant problem areas.
+This is done in the context of the underlying use case model of DaaS systems.
+Subsequently, we give a more comprehensive description
+of our specific design decisions and tailored solutions.
 
-From our perspective, performance, in general can be seen as the most crucial system property
-which influences a majority of all other other relevant properties,
+Performance, from our viewpoint, stands out as the most crucial system property,
+implying influence of other  relevant properties,
 such as user experience, system responsiveness, and overall efficiency.
-As a result, we had to address multifaceted
-and potentially cross-concerning challenges
-on several distinct layers within our architecture.
+Consequently, we encountered multifaceted challenges
+that potentially intersected across several layers within our architecture.
 
-From the core infrastructure to backend services
-and user-facing elements, each component individually
-has to achieve specific minimum requirements in its domain
-and to contribute to the overall performance goals of the whole system.
-Hence, we outline key limiting factors that impede performance,
-along with our experiences and tailored solutions
-aimed at mitigating these challenges.
-From our point of view, the three most critical performance aspects
-for Daas-Systems are:
-\begin{itemize}
-	\item Infrastructure and Platform
-	\item Network Latency and Bandwidth
-	\item Internal Processing and Communication
-\end{itemize}
-As there is no generalized measurement system available,
-we ordered them by their magnitude of impact
-and based them to the best of our knowledge
-on the obtained experiences during our research project.
-Since each of them has their own characteristics,
-specific behavior and requirements, implying a potential impact
-on the overall system performance and resource utilization,
-we describe them individually and in more detail
-in the following separate subsections.
+Hence, every component from the core infrastructure
+to all backend services and user-facing elements,
+must meet individual and domain-specific requirements
+to contribute to the overall performance goals of the entire system.
+Thus, we outline key limiting factors that hinder performance,
+along with our experiences and the chosen mitigation strategies.
+
+Given the absence of a generalized measurement system,
+we ranked these factors based on their magnitude of impact,
+drawing from our experiences obtained during the project.
+
+% From our point of view, the three most critical performance aspects
+% for Daas-Systems are:
+% \begin{itemize}
+% 	\item Infrastructure and Platform
+% 	\item Network Latency and Bandwidth
+% 	\item Internal Processing and Communication
+% \end{itemize}
+% Since each of them has their own characteristics,
+% specific behavior and requirements, implying a potential impact
+% on the overall system performance and resource utilization,
+% we describe them individually and in more detail
+% in the following separate subsections.
 
 \subsubsection{Infrastructure and Platform}
-In general, it can be said that the chosen infrastructure, platform, and utilized software
-may have the most influential impact on CPU and resource consumption
-which has a dramatic impact on the overall system performance.
-As a DaaS combines all aspects of IaaS, PaaS, and SaaS into one system
-each related performance characteristic has to be taken into account.
+In general can be said that the chosen infrastructure, platform,
+and utilized software combined may have the most impact on the overall resource consumption
+and are therefore of major importance.
+As a DaaS system combines all aspects of IaaS-, PaaS, and SaaS-based computation
+each individual performance characteristics has to be taken into account.
 
-For Infrastructure-as-a-Service can be said
-that our solution generally offers possibilities
-to use either virtual machines or docker containers.
-As can be seen in several scientific publications
-discussing relevant performance aspects of such systems,
-the obvious advantage in most categories
-has a lightweight container solution compared to the virtual machine
-if an identical program was tested in both environments
+Regarding Infrastructure-as-a-Service (IaaS),
+our solution generally offers the flexibility
+to use either virtual machines or containers.
+Scientific publications discussing relevant performance aspects of such systems
+highlight the apparent advantage of a lightweight container solution
+over virtual machines when identical programs are tested in both environments
 ~\cite{felter2015updated,potdar2020performance,seo2014performance}.
 
-On the one hand, that implies for our solution,
-that a container-based approach is most likely preferable,
-if that choice is available.
-On the other hand, that  choice is not always available,
-as not every app can be run in a container-based environment.
-To maximize coverage of possible apps being able to run on our system,
-we offer container-based and virtual machine-based application hosting
-as a critical concept of our solution.
+This suggests that a container-based approach is likely preferable
+if the choice is available.
+However, not every application can be run
+in a container-based environment,
+so we maximize coverage of potential usecases
+by offering both container-based and virtual machine-based application hosting.
 
-As a second observation in the area of PaaS, it can be added
-that in most studies, there is no dramatic difference
-if derivates of operating systems are compared to each other.
-Nevertheless, there are derivates
-which are more driven towards being lightweight and more performant
-~\cite{balen2020performance,boras2020performance}.  %nix vs. nix win vs. win
-% ~\cite{balen2020performance}. %win in vm
-In cases where a Windows virtual machine can be replaced with a natively operating wine platform,
-additional performance potentials can be leveraged
+Regarding Platform-as-a-Service (PaaS),
+studies show that there is no dramatic difference in performance
+when derivatives of operating systems are compared to each other
+~\cite{balen2020performance,boras2020performance}.
+Nevertheless, some derivatives are more lightweight and performant.
+In contrast, comparing operating systems of different types
+reveals more obvious results, particularly when considering
+low-level operating system functionality implementations
+such as interrupt handling, memory allocation, and boot time
+~\cite{sergeev2022docker,sulaiman2021comparison}.
+Additional performance potentials can be leveraged
+in cases where a Windows virtual machine can be replaced
+with a container-based solution integrating wine
 ~\cite{huang2012performance}.  % win inside vm
-In contrast to that, if the operating systems of different types are compared to each other,
-there are more obvious results
-especially when taking into account implementations of low-level operating system functionality,
-such as interrupt handling, memory allocation, driver-specific implementations
-or other more subtle requirements such as boot time
-or the time needed for maintenance or system updates
-~\cite{sergeev2022docker,sulaiman2021comparison}. % docker, win vs nix
-% ~\cite{sulaiman2021comparison}. % office apps win vs nix
 
-As a result of our research, we aim to support
-any Windows or Linux-based operating system in virtual machines
-but can only support Linux and Wine in containers
-This is because only terminal-based Windows containers are supported by Microsoft.
+For SaaS, performance largely depends on the underlying usage model
+and the actual software being virtualized.
+Depending on the use case, influential properties
+such as multi-tenancy, provision strategies or the virtualized data type
+might be utilized to varying degrees, resulting in overall
+better or worse performing systems ~\cite{zhong2010virtualization}.
+
+Despite the potential for usecase-specific optimization,
+we emphasize that for DaaS systems,
+coverage of a broad range of potential use cases is generally more favorable
+than solving a particular use case in the most efficient way.
+Therefore, we consider all user-facing desktop applications
+as plausible candidates for migration to our systems,
+rather than more complex distributed systems
+for which a tailored environments might be more suitable.
+
+As a result of these considerations, we aim to support
+any Windows or Linux-based application in virtual machines
+and support any Linux- and Wine-based applications for container environments.
 
 \subsubsection{Network Latency and Bandwidth}
-Another essential aspect when providing web-based access to DaaS apps
-is that also network latency and bandwidth play a significant role
-which contributes to a large degree to the responsiveness of the system
-directly experienced by the user.
-As the user utilizes its application through a web browser
-we aim to support standardized frame-buffer protocols
-to enable user interaction and access to external devices such as printers or USB sticks.
+Network Latency and Bandwidth are crucial factors
+in delivering web-based access to DaaS apps,
+directly impacting system responsiveness experienced by users.
+Supporting standardized frame-buffer protocols enables user interaction initially
+and provide additional access to external devices like printers or USB sticks.
+In cases involving for example video content transmission,
+network bandwidth often becomes the main bottleneck,
+especially with high-resolution video streams.
+In contrast to that,  when transmitting audio, mouse movements,
+or keyboard inputs, network latency becomes more important.
 
-Within such protocols and in cases where video content is transmitted,
-the provided network bandwidth might be the main bottleneck
-especially when it comes to high-resolution video streams.
-In cases where audio, mouse movement, or keyboard strokes are transmitted,
-the network latency plays a more critical role.
-Therefore, the experienced responsibility and usability
-crucially depends on the hosted application.
+Application-specific requirements can be managed to some extent
+through protocol-inherent tuning parameters and compression algorithms.
+Our solution therefore offers a range of tuning options
+in order to adapt to different use cases
+and empowering users to choose their own tailored solution.
 
-In general, such application-specific requirements can, to a certain degree, be controlled
-by protocol-inherent tuning parameters and utilized compression algorithms.
-Therefore, our solution aims to maximize coverage of potential use cases
-by offering several solutions and a set of appropriate tuning options
-which intentionally forces the user to decide
-which solution fits best for him.
+We utilize the Guacamole desktop gateway implementation,
+supporting widely-used frame-buffer protocols like VNC and RDP,
+along with relevant tuning options.
+Guacamole also enables audio support and integration with external devices
+through a JavaScript-based web client.
 
-For that purpose, we utilize the desktop gateway implementation Guacamole,
-which offers support for the most utilized frame-buffer protocols, VNC and RDP
-as well as all relevant tuning options.
-Guacamole not only implements out-of-the-box support for the protocols itself
-but also adds the ability to integrate audio support for VNC or other external devices
-as well as a JavaScript-based web client.
+However, integrating Guacamole into the system
+requires consideration of additional factors not handled by the software.
+Authentication mechanisms for example are particularly important in a DaaS context,
+as users expect seamless authentication across the whole system
+without multiple logins per session.
+Additionally, safeguarding sensitive information like user passwords is crucial
+for maintaining system security and user trust.
+To mitigate the risk of unintended system manipulation,
+Guacamole was integrated with a proxied approach and by using websockets.
+This ensures that credentials of a generated instance remain secret to the host system
+while still providing authorized, passwordless access to the hosted application.
 
-Nevertheless, to integrate guacamole into the system,
-there are still things to consider that are not handled by guacamole
-but would be assumed from a full-fledged DaaS solution.
+Moreover, the proxy mechanism enables support for other additional features
+such as dynamic desktop resizing.
+While Guacamole offers support for that by default in RDP,
+support is completely missing for the VNC protocol.
+Nevertheless, the proxied approach enables to intercept all exchanged websocket opcodes
+and by listening for size requests
+the requested feature could be added
+by directly interacting with the instance on another layer.
 
-First and foremost, how authentication is handled must be considered.
-On one hand, from a DaaS system, a user would expect the system to work
-without forcing him to authenticate himself
-to multiple authentication systems and only once per session.
-On the other hand, crucial information such as user passwords
-might, in general, better be kept secret from the system
-so that users can not manipulate the provided system in an unintended way.
-A potential problem arises if it is considered
-that, by default, RDP requires a valid user session to the system.
-
-For that reason, guacamole was integrated with a proxied approach using proxied websockets.
-This enables keeping the credentials of the generated system secret to the host system
-while still providing authorized but passwordless access to the hosted application.
-
-Further, the proxy mechanism enables the support of dynamic desktop resizing
-which is a default guacamole feature for RDP,
-but which is not available in VNC through guacamole.
-By intercepting WebSocket opcodes and listening for size requests,
-the requested resolution can then be enforced
-by directly interacting with the related container or virtual machine.
-
-Both changes to the default guacamole use case are mandatory
-to comply with the requirements for the underlying DaaS design
-but both also impact the overall system performance and responsiveness in a negative way.
-From a network perspective, the integration as a proxy adds network hop
-and leads to a noticeable delay in network communication.
-The protocol-level evaluation of opcodes additionally enhances that delay, and perceived responsiveness might suffer if resize requests take more time than expected.
-Overall, the introduced overhead is still acceptable
-and enables platform-independent and seamless access for all defined DaaS use cases.
+While these changes to the default Guacamole use case are necessary to align
+with the underlying DaaS design requirements,
+they also negatively impact overall system performance and responsiveness.
+The integration as a proxy introduces additional network hops,
+leading to noticeable delays in network communication.
+Additionally, the protocol-level evaluation of opcodes further increases this delay,
+potentially affecting perceived responsiveness
+if resize requests take longer than expected.
+However, despite the introduced overhead,
+the solution still enables platform-independent
+and seamless access for all defined DaaS use cases.
 
@@
 Another critical aspect directly related to performance and responsiveness
-is the approach in which data is internally processed, stored, and communicated
-to all involved components and instances.
-This includes the metadata being collected and processed by essential components
-but also implies raw physical data being directly exchanged with object instances.
+is how data is internally processed, stored, and communicated
+among all involved components and instances.
+This includes metadata collected and processed by essential components
+and involves direct exchange of raw data with object instances.
 
-To provide a satisfying user experience in a DaaS system
-specific components in a host system are required.
+Ensuring a satisfying user experience in a DaaS system
+implies specific components in a host system.
 This includes access to all backend services
-being used to create containers and virtual machines,
-a local or remote database, a local or remote authentication service,
-local and distributed filesystems, as well as
-virtual or physical network devices.
-Further, a web server is integrated to provide all relevant API endpoints
-for the frontend communication
-as well as all WebSocket endpoints for viewer communication.
+used to create containers and virtual machines,
+local or remote databases, authentication services, filesystems, and network devices.
+Additionally, a web server is integrated to provide relevant API endpoints
+for frontend communication and websocket endpoints for viewer communication.
 
-Generally spoken, all of these components might become a potential bottleneck
-when being utilized with high throughputs
-or large amounts of concurrent applications demanding low latencies.
-This is especially the case if being used in a sequential manner
-or being served from the same hardware.
-If delays add up, the total time needed for a request to be fully processed can increase drastically.
+Generally spoken, all of these components can potentially become bottlenecks
+when handling high throughputs or large amounts of concurrent applications
+requiring low latency access and especially if all necessary backend components
+are served from the same hardware.
+Architecturally, all relevant components
+must therefore be designed for distribution across different hosting systems.
+However, retaining certain components directly on the host system
+may still be required to ensure low-latency access.
 
-On an architectural level, all relevant components are, therefore designed
-for being distributed to different hosting systems
-if the requirements of a particular user scenario suggest it.
-Nevertheless, it might still be advisable
-to keep specific components directly on the host system
-to ensure low latency component access.
+From our observations, components requiring realtime and low-latency performance,
+should physically be kept as close to the core hosting system as possible.
+This includes all data and communication relevant to
+audio, video, mouse, and keyboard transmission,
+as well as all backend services for hosting containers and virtual machines.
+Ideally, these components should be hosted natively on the host system
+for virtual network and hardware access
+or at least be connected to the same physical network
+in order to minimize network latencies.
 
-For all components with real-time and low-latency latency usecases, it is generally advisable
-to keep them as close to the core hosting system as possible.
-This includes all data and communication
-relevant to audio, video, mouse, and keyboard transmission,
-but also the backend service pertinent to host containers and virtual machines.
-That might be in the best case to host them natively on the host itself
-which enables virtual network and hardware access.
-In the worst case, that might be at least connected to the same physical network
-to reduce network latencies as much as possible.
-
-For all other components of the overarching nature
-such as database access, authentication provider, and filesystems
-it is still advisable to keep a close distance from the host
-but it also highly depends on the total amounts of users
-and the specifics of their use cases.
-On small scales, it is still advisable to keep such components on the same system
-but on larger scales, it is preferable to externalize such functionality to dedicated systems.
+Regarding other components such as database access,
+authentication providers, and filesystems,
+maintaining a close distance from the host is advisable,
+but this depends largely on the total number of users
+as well as their specific use cases.
+On smaller scales, keeping these components
+in close physical proximity to the host remains prudent.
 
@@
 %	- Considerations:
-%		- Divide and conqueror
+%		- TODO: Mention that term: Divide and conquer
 %		- Requirement: Distributable components
@@
 
-Stability in DaaS applications is of significant importance
-to ensure that our solution can meet the demands of continuous
-and time-critical operations across varied user scenarios.
-From our point of view, any DaaS system must ensure the stability
-of all system components during runtime
-to maintain the ability to provide service continuity,
-handle growing user loads and transparently recover from failures,
-while still providing a seamless end-user experience.
-Such considerations are an integral part of our architectural concept
-as they directly impact the operational integrity
-of the system and its hosted instances,
-making them suitable for enterprise-level deployment and trust.
+Stability in DaaS applications is especially important
+for meeting the demands of continuous and time-critical operations
+across diverse user scenarios.
+Ensuring the stability of all system components during runtime is paramount
+to maintain service continuity, handle growing user loads,
+and transparently recover from failures
+while providing a seamless end-user experience.
+These considerations are integral to our architectural concept,
+directly impacting the operational integrity of the system
+and its suitability for enterprise-level deployment and trust.
 
-This definition of stability is underpinned by several architectural
-and operational strategies, including redundancy of components, scalability of nodes,
-and a robust backend service framework
-providing means of communication between components and nodes.
-These strategies, as well as their drawbacks and benefits
-will be further discussed on a component level
+This definition of stability is supported
+by various architectural and operational strategies,
+including component redundancy, node scalability,
+and a robust backend service framework facilitating communication
+between components and nodes.
+These strategies, along with their drawbacks and benefits,
+will be further explored at a component level
 and within the following subsections.
@@
 
-Scalability in general, but also specifically in DaaS environments
-can be approached in two fundamental ways:
-Horizontal and vertical scaling.
+Scalability in DaaS environments
+can be approached through two fundamental methods: horizontal and vertical scaling.
 Horizontal scaling, or scaling out, involves adding more machines or instances
-to a pool to manage the increased load.
-In contrast to that, vertical scaling or scaling up,
-refers to adding more power (CPU, RAM)
-to an existing machine~\cite{vaquero2011dynamically}.
-
-While horizontal scaling is generally more flexible
-and often preferred in cloud environments
-due to its ease of integration with existing architectures,
+to manage increased loads.
+In contrast, vertical scaling, or scaling up,
+adds more power (CPU, RAM) to existing machines
+~\cite{vaquero2011dynamically}.
+While horizontal scaling offers flexibility
+and is often preferred in cloud environments,
 vertical scaling remains important for applications
@@
 
-To comply with both, we defined means of configuration
-such that the system itself, as well as all of its backend services
-are designed to be split up into logical units
-which can be distributed to dedicated systems.
-Essentially, all backend services receive their dedicated configuration
-and can be accessed via tuples of hostname and port.
-Any component using such a backend service can obtain the relevant tuple
-and then use it to fetch requests from that service.
-By default and in a standalone configuration,
-all backend services reside on the same native host, and relevant tuples can be derived from the host configuration.
-In a distributed context, a selection of these services
-can be distributed to a dedicated and remote system
-to comply with horizontal scaling principles.
-The relevant tuple can then be derived from any node in the system
-by utilizing a common database containing the relevant meta information.
+To comply with both approaches,
+we designed the system and its backend services
+to be split into logical units that can be distributed across dedicated systems.
+Each backend service receives dedicated configuration
+and can be accessed via hostname and port tuples.
+By default, all backend services reside on the same host,
+with tuples derived from the host configuration.
+In a distributed context, services can be distributed to dedicated remote systems
+to adhere to horizontal scaling principles.
+Relevant tuples can then be derived from any node
+using a common database containing meta information.
 
-Vertical scaling of DaaS nodes is currently not planned
-but administrators are free to use custom 3rd party toolchains, e.g., for GPU virtualization.
-Although vertical scaling on a host basis
-is beyond the scope of our research project,
-it is more plausible to scale the instances
-being hosted within that system and for both dimensions.
+Vertical scaling of DaaS nodes is not currently planned,
+but administrators are free to use custom third-party toolchains,
+such as GPU virtualization, for vertical scaling on a per-host basis.
+Although vertical scaling within a host is beyond the scope of our research project,
+scaling the instances hosted within the system is feasible for both dimensions.
 
-For vertical scaling of instances, appropriate parameters and API endpoints are present
-to manipulate the configuration of their resource assignments,
-such that arbitrary configurations are possible.
-
-For horizontal scaling of instances
-the builtin mechanisms of each backend service are automatically configured and utilized
-such that instances can be arbitrarily distributed across a pool of connected nodes.
-In the case of virtual machines, the responsible backend service Proxmox implements
-such mechanisms as part of its 'Proxmox VE High Availability (HA)' cluster.
-In the case of containers, any backend service fulfilling
-similar concepts can be integrated, for example, with Kubernetes.
-Within such clusters, any DaaS instance is entirely agnostic to its hosting system
-and must only be reachable via a known hostname or IP.
+For vertical scaling of instances, we provide appropriate parameters
+and API endpoints to manipulate resource assignments,
+allowing for arbitrary configurations.
+For horizontal scaling of instances, the built-in mechanisms
+of each backend service are automatically configured
+and utilized to distribute instances across a pool of connected nodes.
+For virtual machines, Proxmox implements such mechanisms
+as part of its 'Proxmox VE High Availability (HA)' cluster.
+For containers, similar concepts can be integrated with any backend service,
+such as Kubernetes.
+Within these clusters,
+DaaS instances are entirely agnostic to their hosting system
+and only need to be reachable via a known hostname or IP as a single requirement.
 
@@
 
-Redundancy is critical for most distributed systems
-but especially for DaaS architectures,
-and serves as a safeguard against service interruptions and data loss
+Redundancy is crucial in distributed systems,
+particularly in DaaS architectures,
+to safeguard against service interruptions and data loss
 by duplicating essential components, functions, and data.
-This design principle ensures that systems remain highly available and reliable,
+This design principle ensures high availability and reliability,
 even during component failures.
-In the following, a list of components eligible for redundancy is given
-which briefly describes the most relevant strategies contained in our architecture.
+In the following, we provide a list of components eligible for redundancy,
+briefly describing the most relevant strategies in our architecture.
 
 \begin{itemize}
-\item \textbf{Frontend/Backend:} In our final version
-redundancy of the frontend and backend components will offer means of configuration
-such that multiple instances can be distributed to dedicated systems
-and can be the target of standardized load-balancing strategies.
-By doing so, the frontend or the backend may act as a gateway
-between the user and the backend service containing the relevant instances.
-In that manner, redundancy is achieved, and frontend or backend
-can forward requests and responses
-by identifying the correct communication endpoint
-based on URL's or specific parts of them,
-for example, by extracting hostnames
-or specific identifiers being unique within the DaaS domain.
-\item \textbf{Database:} The database is integrated as a backend service
-and as such, completely unaware of surrounding components
-and contexts in which it is used.
-The database can, therefore, be distributed to a dedicated system if needed.
-Generally, any modern database solution (e.g., MariaDB) implements rich features
-to maintain specific data consistency strategies
-such as data replication, clustering,
-dedicated failover mechanisms or also RAID control.
-Therefore, redundancies in our database system are currently not planned,
-but can easily be applied by users with such needs
-and by using database-specific features and a modified database adapter.
-\item \textbf{Authentication: } Authentication within our architecture
-is also implemented as a backend service
-and can, therefore, be distributed to any dedicated system,
-can be a target of standardized load-balancing mechanisms and strategies
-and is as well fully agnostic of surrounding components
-and calling contexts.
-The authentication system maintains
-its credential registry in a separate database,
-a customizable set of user and group privileges,
-and implements the well-defined OAuth 2.0 standard
-to allow remote authentication.
-A customizable amount of redundant systems might be configured in our final version.
-	      % TODO: Discuss: Is that true? Do we want to make Darios authenticator redundant?
-\item \textbf{Distributed Filesystem:} Within our architecture
-we use the distributed filesystem Ceph, which is a distributed storage system
-providing object, block, and file storage. In Ceph, redundancy is achieved
-through its self-healing and self-managing capabilities utilizing an algorithm
-called CRUSH (Controlled Replication Under Scalable Hashing).
-Data availability and fault tolerance are obtained
-by automatically managing replication
-of data pieces across different devices.
-Our designed architecture uses these capabilities
-to reliably provide
-file storage to host systems and all hosted DaaS instances
-and provides block storage to the infrastructural backend services.
-By hosting DaaS instances directly from the distributed block storage,
-arbitrary redundancy scenarios might be covered
-by using specific configuration parameters.
+	\item \textbf{Frontend/Backend:}
+	      In our final version,
+	      redundancy of the frontend and backend components
+	      will enable configuration for distributing multiple instances
+	      to dedicated systems and utilizing standardized load-balancing strategies.
+	      This setup allows the frontend or backend
+	      to serve as a gateway between the user and the backend service
+	      containing all relevant instances.
+	      Both components contribute to redundancy
+	      by distributing requests and responses to their respecitive recipients
+	      instead of processing them on their own.
+	\item \textbf{Database:}
+	      The database is integrated as a backend service
+	      and is therefore independent of surrounding components and contexts.
+	      It can be distributed to a dedicated system if necessary.
+	      Modern database solutions, like MariaDB,
+	      offer rich features for maintaining specific data consistency strategies
+	      such as data replication, clustering, failover mechanisms,
+	      and RAID control.
+	      Currently, redundancies in our database system are not planned
+	      but can be easily applied by users using such database-specific features
+	      and a modified database adapter.
+	\item \textbf{Authentication: }
+	      Authentication within our architecture is implemented as a backend service,
+	      allowing for distribution to any dedicated system.
+	      It can also be targeted by standardized load-balancing mechanisms
+	      and remains fully agnostic of surrounding components
+	      or calling contexts.
+	      The authentication system maintains its credential registry
+	      in a separate database, along with a customizable set
+	      of user and group privileges.
+	      Additionally, it implements the well-defined OAuth 2.0 standard
+	      to facilitate remote authentication.
+	      In our final version,
+	      a customizable amount of redundant systems may be configured.
+	\item \textbf{Distributed Filesystem:}
+	      Within our architecture, we utilize the distributed filesystem Ceph,
+	      which offers object, block, and file storage capabilities.
+	      Redundancy in Ceph is implicitly achieved
+	      through its self-healing and self-managing features, employing
+	      the CRUSH algorithm (Controlled Replication Under Scalable Hashing).
+	      Data availability and fault tolerance are ensured
+	      by automatically replicating data across different devices.
+	      Our architecture leverages these capabilities
+	      to reliably provide file storage to host systems
+	      and all hosted DaaS instances,
+	      while also offering block storage to the infrastructural backend services.
+	      Hosting DaaS instances directly from the distributed block storage
+	      enables migration and redundancy of instances or in other words, horizontal scaling.
 	\item \textbf{Instances:} Redundancy of instances
-It is not a necessary condition for DaaS systems in general, but applying such strategies under certain conditions might make sense.
+	      It is not a necessary condition for DaaS systems in general,
+	      but applying such strategies under certain conditions might be plausible.
 	      By using highly available clusters within our backend services
@@
 \subsubsection{Internal Communication}
-After discussing the principles being used to achieve scalability and redundancy
-it adheres to describe how communication flows are maintained
-in such a distributed environment.
+After discussing the principles of scalability and redundancy,
+we finally discuss how communication flows are managed in our distributed environment.
+Our proposed solution is rooted in a message-oriented architecture
+designed to fulfill DaaS properties for scalability and redundancy,
+as previously outlined.
 
-Our proposed solution is, at the core, designed
-as a message-oriented architecture
-which additionally aims to fulfill DaaS properties
-for scalability and redundancy as previously described.
+This implies that any action triggered, whether in the frontend on the client side,
+backend on the host side, or within hosted instances,
+must ensure that the necessary information for execution
+is readily available at the time of processing.
 
-That implies fundamentally that any action being triggered,
-may it either be in the frontend on the client side,
-in the backend on the host side
-or within any of the hosted instances
-must always fulfill the condition that the required information to execute that action
-must either be directly derivable from a request or a corresponding response
-or otherwise, must be immediately available to the component executing that action.
+We achieve this by storing runtime information in distributed block storage,
+relevant configuration files in distributed file storage,
+object metadata in a common database,
+and by utilizing a decentralized OAuth component for authentication.
 
-We achieve that by storing all runtime information of instances in the distributed block storage,
-by storing all relevant configuration files in our distributed file storage,
-by utilizing a common database to persist all object metadata
-and by using our decentralized OAuth component to handle authentication.
+This ensures that any change triggered by a node within the DaaS domain
+is immediately available to any other node relying in some way
+on that particular information change.
+Additionally, any extra information needed by a component must be fully contained
+within the related request or response.
+This enables arbitrarily distributed configurations as long as request
+and response messages reach their desired target eventually
+and as long relevant backend services remain available.
 
-By doing so, any change being triggered by a node within the DaaS network
-is immediately available to any other node in the system
-which relies on that information change.
-Apart from that, any additional information needed by a component
-is fully contained within the related request or response.
-
-Therefore, arbitrarily distributed configurations are possible
-as long as request and response messages reach their desired target, eventually
-and as long as the availability of all relevant backend services is guaranteed.
-
-If that is the case, each distributed DaaS node can typically
-just contain all time-critical components at a bare minimum
-to maintain an acceptable performance throughout the whole user-facing network stack.
-This would typically include the backend service hosting the relevant instances
-as well as the WebSocket proxy directly connected to the guacamole daemon
-and all other components may reside somewhere in the DaaS network.
+If that is the case, each distributed DaaS node might only contain essential and time-critical components,
+including the backend service hosting relevant instances as well as the described
+websocket proxy connected to the guacamole daemon and the user.
+Any other component may reside elsewhere within the DaaS network.
 

### 2024-04-29 paper.tex
Eliminate TODO's in usability section

@@
 \subsection{Usability}
+An essential feature of a DaaS platform is its strong user-friendliness,
+which eliminates the complexity of the different operating systems
+used inside virtual machines and containers,
+makes the underlying hardware transparent,
+and simplifies the necessary interactions as much as possible.
 
-% TODO: Christian (1)
-% TODO: Johannes (2)
+Since users already have an operating system on their clients,
+and the motivation for using a DaaS is the simple use of various applications,
+it makes no sense for most users to have additional desktops available in the browser.
+The focus should be on the applications.
+For this reason, when developing a new DaaS solution,
+the main goal should be to reduce the graphical output
+to the limits of the respective application.
+The concept uses a separate browser tab for each application
+whose content exclusively represents the application.
 
-An essential feature of a DaaS platform is its strong user-friendliness, which eliminates the complexity of the different operating systems used inside virtual machines and containers, makes the underlying hardware transparent, and simplifies the necessary interactions as much as possible.
+Developing the automatic adaptation of the output
+to the limits of the respective application window is challenging.
+Only a few RDP and VNC server implementations offer to export the graphical output
+of only individual windows or processes.
+For Windows operating systems, the proprietary built-in RDP server
+and the free software TightVNC offer the export of full desktops and single windows.
 
-Since users already have an operating system on their clients, and the motivation for using a DaaS is the simple use of various applications, it makes no sense for most users to have additional desktops available in the browser.
-The focus should be on the applications.
-For this reason, when developing a new DaaS solution, the main goal should be to reduce the graphical output to the limits of the respective application.
-The concept uses a separate browser tag for each application whose content exclusively represents the application.
+For Linux operating system deployments,
+the free software projects xrdp and x11vnc enable all desired features
+and single-app functionality can be easily obtained
+by using a custom window manager such as for example qtile
+~\footnote{\url{https://qtile.org/}}.
 
-Developing the automatic adaptation of the output to the limits of the respective application window is challenging.
-Only a few RDP and VNC server implementations offer to export the graphical output of only individual windows or processes.
-For Windows operating systems, the proprietary built-in RDP server and the free software TightVNC offer the export of full desktops and single windows.
-For Linux operating system deployments, the free software projects xrdp and x11vnc enable all desired features.
-Table ~\ref{tab:DaaS_Services_Overview} includes an overview of the mentioned service implementations for RDP and VNC protocols and their relevant characteristics.
-Further implementations like TigerVNC and protocols like NX NoMachine exist but do not seem useful for designing and implementing a novel DaaS because they do not meet all feature requirements.~\cite{OJCC_2023v8i1n01_Baun}
+Table~\ref{tab:DaaS_Services_Overview} includes an overview
+of the mentioned service implementations for RDP and VNC protocols
+and their relevant characteristics.
+Further implementations like TigerVNC and protocols like NX NoMachine exist
+but do not seem useful for designing and implementing a novel DaaS because
+they do not meet all feature requirements.~\cite{OJCC_2023v8i1n01_Baun}
 
@@
 		\hline
-		x11vnc                  & VNC               & supported              & Linux              & GPL2              \\
+		x11vnc                  & VNC               & not supported          & Linux              & GPL2              \\
 		\hline
@@
 
-The automatic adaptation of browser tabs content size, is done by.... TOTO: Johannes´...
+The automatic adaptation of viewer size and screen resolution,
+considering underlying protocols and systems,
+faced various limitations, briefly described below.
 
-TODO: Ideen und Konzepte zur Trennung von User und Admin-Ansicht
+One challenge was imposed by Guacamole's current lack of support
+for dynamic desktop resizing with VNC,
+leading to misaligned layouts and a poor user experience.
+To address this until such features are available in Guacamole
+\footnote{\url{https://issues.apache.org/jira/browse/GUACAMOLE-1196}},
+we implemented a separate mechanism using a websocket proxy
+and direct interaction with each instance.
 
-TODO: Herausforderungen, Entwicklung, Implementierung, Lessons-learned
+Additionally, protocol-specific behaviors may vary
+when hosted on different platforms or infrastructures.
+For instance, we observed connection terminations when resizing Linux-based containers,
+contrasting with the behavior observed in equivalent Linux-based virtual machines.
+To overcome such limitations, we created a separate layer around the Guacamole client,
+adapting to each infrastructure, platform, or protocol-specific configuration.
 
-TODO Johannes: Erlebnisse der letzten 3-4 Wochen aufschreiben
+Another challenge was managing RDP sessions which are typically tied to actual users,
+unlike VNC sessions which are typically authenticated against
+a server-maintained user registry.
+Utilizing the websocket proxy, the backend could pass credentials
+to the server without revealing them to the user.
+
+Essentially, Guacamole with its protocol implementations and additional features like audio and printing,
+provided the fundamental base for our core capabilities.
+
+Finally mentioned, certain instances might not be fully accessible during runtime
+due to hardware malfunctions, inappropriate system configurations for simply during system updates.
+Therefore, as a failover mechanism additional logic had to be added
+in order to provide reasonable debug views and recovery mechanisms,
+e.g.~to recreate instances from a specific snapshot or configuration state.
+
+A crucial lesson learned was therefore to realize that
+while individual problems can mostly be solved
+using established paradigms, services or libraries with only slight modifications,
+the real challenge for a full-fledged DaaS system lies in it's abilities
+to combining such isolated solutions appropriately
+and for each specific infrastructure and platform configuration.
 

### 2024-05-01 paper.tex
Add figures to chapters performance and stability

@@
 \usepackage{graphicx}
+\usepackage{tikz}
+\usetikzlibrary{positioning, shapes.geometric}
+\usepackage{adjustbox}
 % Used for displaying a sample figure. If possible, figure files should
@@
 
-As a result of these considerations, we aim to support
-any Windows or Linux-based application in virtual machines
+Figure~\ref{hosting_scenarios} visualizes the available hosting startegies for our platform.
+We aim to support any Windows or Linux-based application in virtual machines
 and support any Linux- and Wine-based applications for container environments.
+
+\begin{figure}
+	\centering
+	\begin{adjustbox}{width=\textwidth}
+		\begin{tikzpicture}[every node/.style={minimum height=1cm, align=center}]
+			%Type Column
+			\node[minimum width=2cm] at (-4,-2) {IaaS};
+			\node[minimum width=2cm] at (-4,-1) {PaaS};
+			\node[minimum width=2cm] at (-4,0) {SaaS};
+			\node[minimum width=2cm] at (-4,1) {DaaS};
+			% DaaS
+			\node[draw, minimum width=12cm, fill=gray!10] at (3,1) {DESIGN Frontend};
+			%Linux
+			\node[draw, minimum width=6cm, fill=gray!50] at (0,0) {Linux App};
+			\node[draw, minimum width=6cm, fill=gray!80] at (0,-1) {Linux OS};
+			\node[draw, minimum width=3cm, fill=gray!100] at (-1.5,-2) {Virtual Machine};
+			\node[draw, minimum width=3cm, fill=gray!100] at (1.5,-2) {Container};
+			%Windows
+			\node[draw, minimum width=6cm, fill=gray!50] at (6,0) {Windows App};
+			\node[draw, minimum width=3cm, fill=gray!80] at (4.5,-1) {Windows OS};
+			\node[draw, minimum width=3cm, fill=gray!80] at (7.5,-1) {Linux OS, Wine};
+			\node[draw, minimum width=3cm, fill=gray!100] at (4.5,-2) {Virtual Machine};
+			\node[draw, minimum width=3cm, fill=gray!100] at (7.5,-2) {Container};
+		\end{tikzpicture}
+	\end{adjustbox}
+	\caption{Application hosting strategies.}\label{hosting_scenarios}
+\end{figure}
 
@@
 Guacamole also enables audio support and integration with external devices
-through a JavaScript-based web client.
+through a JavaScript-based web client (guacamole.js).
+As depicted in figure~\ref{websocket_proxy},
+the proxy component recives HTTP requests from the guacamole.js webclient
+and forwarded by the backend by using regular WebSocket opcodes (Guacamole protocol).
+Such communication is forwarded to the Guacamole daemon (guacd)
+which in turn communicates to each instance
+by using the frame-buffer protocol of choice (VNC or RDP).
 
 However, integrating Guacamole into the system
-requires consideration of additional factors not handled by the software.
+requires consideration of additional factors besides the frame-buffer protocols itself.
 Authentication mechanisms for example are particularly important in a DaaS context,
@@
 for maintaining system security and user trust.
-To mitigate the risk of unintended system manipulation,
-Guacamole was integrated with a proxied approach and by using websockets.
+To mitigate the risk of unintended system manipulation, Guacamole was integrated
+in a proxied approach and by using websockets.
 This ensures that credentials of a generated instance remain secret to the host system
@@
 and seamless access for all defined DaaS use cases.
+
+\begin{figure}
+	\centering
+	\begin{adjustbox}{width=\textwidth}
+		\begin{tikzpicture}[every node/.style={minimum height=1cm, align=center}]
+
+			% Nodes center row
+			\node[draw, minimum width=3cm, fill=gray!10] (browser) at (0,0) {guacamole.js };
+			\node[draw, minimum width=3cm, fill=gray!50] (back) at (4,0) {DESIGN Backend};
+			\node[draw, minimum width=3cm, fill=gray!50] (proxy) at (8,0) {Proxy};
+			\node[draw, minimum width=3cm, fill=gray!50] (guacd) at (12,0) {guacd};
+			\node[draw, minimum width=3cm, fill=gray!100] (inst) at (16,0) {Instances};
+			% Arrows center row
+			\draw[<->, thick, black] (browser.east) -- (back.west);
+			\draw[<->, thick, black] (back.east) -- (proxy.west);
+			\draw[<->, thick, black] (proxy.east) -- (guacd.west);
+			\draw[<->, thick, black] (guacd.east) -- (inst.west);
+			% Upper row
+			\node[draw, minimum width=3cm, fill=gray!10] (client-backend) at (2,2) {HTTP/WebSocket};
+			\node[draw, minimum width=3cm, fill=gray!50] (backend-proxy) at (6,2) {HTTP/WebSocket};
+			\node[draw, minimum width=3cm, fill=gray!50] (proxy-guacd) at (10,2) {Guacamole protocol};
+			\node[draw, minimum width=3cm, fill=gray!100] (guacd-inst) at (14,2) {Framebuffer protocols/\\VNC,RDP};
+			% Dashed lines upper row
+			\draw[-, dashed, thick, black] (2,0) -- (2,1.5);
+			\draw[-, dashed, thick, black] (6,0) -- (6,1.5);
+			\draw[-, dashed, thick, black] (10,0) -- (10,1.5);
+			\draw[-, dashed, thick, black] (14,0) -- (14,1.5);
+			% Arrow bottom row
+			\draw[<->, thick, black] (0,-1) -- (3.8,-1) node[midway, below, font=\footnotesize] {Client Communication};
+			\draw[<->, thick, black] (4,-1) -- (11.8,-1)node[midway, below, font=\footnotesize] {Backend Communication};
+			\draw[<->, thick, black] (12,-1) -- (16,-1)node[midway, below, font=\footnotesize] {Instance Communication};
+		\end{tikzpicture}
+	\end{adjustbox}
+	\caption{WebSocket Proxy Components.}\label{websocket_proxy}
+\end{figure}
 
@@
 From our observations, components requiring realtime and low-latency performance,
-should physically be kept as close to the core hosting system as possible.
+should physically be kept as close to the core hosting system as possible
+as shown in figure~\ref{perf_internal_communication}.
 This includes all data and communication relevant to
@@
 in close physical proximity to the host remains prudent.
+
+\begin{figure}
+	\centering
+	\begin{adjustbox}{width=\textwidth}
+		\begin{tikzpicture}[every node/.style={minimum height=1cm, align=center}]
+			% Standalone Realtime
+			\node[draw, minimum width=8cm, fill=gray!10] (singlefront) at (0,-0.5) {DESIGN Frontend};
+			\node[draw, minimum width=8cm, fill=gray!30] (singleback) at (+0,-2) {DESIGN Backend};
+			\node[draw, minimum width=4cm, fill=gray!100] at (-2,-3) {Infrastructure\\(Proxmox/Docker) };
+			\node[draw, minimum width=4cm, fill=gray!100] at (-2,-4) {Proxy};
+			\node[draw, minimum width=4cm, fill=gray!100] at (-2,-5) {Backend API};
+			% Standalone Services
+			\node[draw, minimum width=4cm, fill=gray!50] at (+2,-3) {Database};
+			\node[draw, minimum width=4cm, fill=gray!50] at (+2,-4) {Filesystem};
+			\node[draw, minimum width=4cm, fill=gray!50] at (+2,-5) {Authentication};
+			%Standalone Caption
+			\node[minimum width=8cm] (browser) at (0,-6) {Standalone};
+			% Distributed Realtime
+			\node[draw, minimum width=10cm, fill=gray!10] (distfront) at (10,-0.5) {DESIGN Frontend};
+			\node[draw, minimum width=10cm, fill=gray!30] (distback) at (10,-2) {DESIGN Backend};
+			\node[draw, minimum width=4cm, fill=gray!100] (inf) at (7,-3) {Infrastructure\\(Proxmox/Docker) };
+			\node[draw, minimum width=4cm, fill=gray!100] (proxy) at (7,-4) {Proxy};
+			\node[draw, minimum width=4cm, fill=gray!100] (api) at (7,-5) {Backend API};
+			% Distributed Services
+			\node[draw, minimum width=4cm, fill=gray!50] (db) at (13,-3) {Database};
+			\node[draw, minimum width=4cm, fill=gray!50] (fs) at (13,-4) {Filesystem};
+			\node[draw, minimum width=4cm, fill=gray!50] (auth) at (13,-5) {Authentication};
+			% Network
+			\node[draw, minimum width=1cm, minimum height=3cm, fill=white] (browser) at (10,-4) {N\\E\\T\\W\\O\\R\\K};
+			% Distributed Caption
+			\node[minimum width=10cm] (browser) at (10,-6) {Distributed};
+			% Arrows back<->front
+			\draw[<->, thick, black] (singleback.north) -- (singlefront.south);
+			\draw[<->, thick, black] (distback.north) -- (distfront.south);
+			% Arrows realtime
+			\draw[<->, thick, black] (inf.east) -- ++(0.5,0);
+			\draw[<->, thick, black] (proxy.east) -- ++(0.5,0);
+			\draw[<->, thick, black] (api.east) -- ++(0.5,0);
+			% Arrows services
+			\draw[<->, thick, black] (db.west) ++(-0.5,0) -- ++(0.5,0);
+			\draw[<->, thick, black] (fs.west) ++(-0.5,0) -- ++(0.5,0);
+			\draw[<->, thick, black] (auth.west) ++(-0.5,0) -- ++(0.5,0);
+		\end{tikzpicture}
+	\end{adjustbox}
+	\caption{Internal Communication Flow}\label{perf_internal_communication}
+\end{figure}
 
@@
 such as GPU virtualization, for vertical scaling on a per-host basis.
-Although vertical scaling within a host is beyond the scope of our research project,
+Although vertical scaling within a host
+is beyond the scope of our research project,
 scaling the instances hosted within the system is feasible for both dimensions.
-
 For vertical scaling of instances, we provide appropriate parameters
@@
 allowing for arbitrary configurations.
+
 For horizontal scaling of instances, the built-in mechanisms
@@
 For virtual machines, Proxmox implements such mechanisms
-as part of its 'Proxmox VE High Availability (HA)' cluster.
+as part of its `Proxmox VE High Availability (HA)' cluster.
 For containers, similar concepts can be integrated with any backend service,
@@
 DaaS instances are entirely agnostic to their hosting system
-and only need to be reachable via a known hostname or IP as a single requirement.
+and only need to be reachable
+via a known hostname or IP as a single requirement.
+As shown in figure~\ref{scale_objects_horizontal},
+we provide a globally maintained state of data, authentication and storage
+which is accessible either by horizontally scaled backends and frontends
+or otherwise,
+by horizontally and vertically scaled instances in our DaaS object domain.
 
-\subsubsection{Redundancy of Components}
+\begin{figure}
+	\centering
+	\begin{adjustbox}{width=\textwidth}
+		\begin{tikzpicture}[every node/.style={minimum width=4cm, minimum height=1cm, align=center}]
+
+			% User
+			\node[draw, minimum width=4cm] (user) at (5,1.5) {User};
+			% Clusters
+			\node[] at (-5,-0.5) {Frontend Cluster\\(Horizontal)};
+			\node[] at (-5,-2.5) {Backend Cluster\\(Horizontal)};
+			\node[] at (-5,-4.5) {Object Domain\\(Horizontal and Vertical)};
+
+			% Node1
+			\node[draw, fill=gray!10]  (front1) at (0,-0.5) {DESIGN Frontend 1};
+			\node[draw, fill=gray!30]  (back1)  at (0,-2.0) {DESIGN Backend};
+			\node[draw, fill=gray!50]  (real1)   at (0,-3.0) {Realtime Components};
+			\node[draw, fill=gray!100] (inst1)  at (0,-4.5) {DaaS Instance (ID=1234)};
+			% Node2
+			\node[draw, fill=gray!10]  (front2) at (5,-0.5) {DESIGN Frontend};
+			\node[draw, fill=gray!30]  (back2) at (5,-2.0) {DESIGN Backend};
+			\node[draw, fill=gray!50]  (real2) at (5,-3.0) {Realtime Components};
+			\node[draw, fill=gray!100] (inst2)  at (5,-4.5) {DaaS Instance (ID=1234)};
+			% Node3
+			\node[draw, fill=gray!10]  (front3) at (10,-0.5) {DESIGN Frontend};
+			\node[draw, fill=gray!30]  (back3) at (10,-2.0) {DESIGN Backend};
+			\node[draw, fill=gray!50]  (real3) at (10,-3.0) {Realtime Components};
+			\node[draw, fill=gray!100] (inst3)  at (10,-4.5) {DaaS Instance (ID=1234)};
+			% Domain
+			\node[draw, dashed, minimum width=14.5cm, minimum height=1.25cm]  (domfront) at (5,-0.5){};
+			\node[draw, dashed, minimum width=14.5cm, minimum height=2.25cm]  (domback) at (5,-2.5){};
+			\node[draw, dashed, minimum width=14.5cm, minimum height=1.25cm]  (dominst) at (5,-4.5){};
+			% Backend Services
+			\node[draw, minimum width=2cm, minimum height=5cm] (backends) at (14,-2.5){Backend\\Services};
+			% Arrows back<->Inst
+			\draw[<->, thick, black] (inst1.north) -- (real1.south);
+			\draw[<->, thick, black] (inst2.north) -- (real2.south);
+			\draw[<->, thick, black] (inst3.north) -- (real3.south);
+			% % Arrows back<->front
+			\draw[<->, thick, black] (front1.south) -- (back1.north);
+			\draw[<->, thick, black] (front2.south) -- (back2.north);
+			\draw[<->, thick, black] (front3.south) -- (back3.north);
+			% % Arrows Insts
+			\draw[<->, thick, black] (inst1.east) -- (inst2.west);
+			\draw[<->, thick, black] (inst2.east) -- (inst3.west);
+			% % Arrows User
+			\draw[->, thick, black] (user.south) -- (front1.north);
+			\draw[->, thick, black] (user.south) -- (front2.north);
+			\draw[->, thick, black] (user.south) -- (front3.north);
+			% % Arrows Backend
+			\draw[<->, thick, black] (domfront.east) -- ++(0.75,0);
+			\draw[<->, thick, black] (domback.east) --  ++(0.75,0);
+			\draw[<->, thick, black] (dominst.east) --  ++(0.75,0);
+		\end{tikzpicture}
+	\end{adjustbox}
+	\caption{DESIGN Component Scaling}\label{scale_objects_horizontal}
+\end{figure}
+
+\subsubsection{Redundancy of Backend Services}
 
@@
 even during component failures.
-In the following, we provide a list of components eligible for redundancy,
-briefly describing the most relevant strategies in our architecture.
-
+Figure~\ref{service_redundancy} depicts all major DESIGN components eligible for redundancy
+and which will be briefly described following:
 \begin{itemize}
@@
 	      instead of processing them on their own.
+	\item \textbf{Instances:} Redundancy of instances
+	      It is not a necessary condition for DaaS systems in general,
+	      but applying such strategies under certain conditions might be plausible.
+	      By using highly available clusters within our backend services
+	      redundancy can be achieved by using the corresponding API endpoints
+	      previously described and by parameterizing them appropriately.
 	\item \textbf{Database:}
@@
 	      enables migration and redundancy of instances or in other words, horizontal scaling.
-	\item \textbf{Instances:} Redundancy of instances
-	      It is not a necessary condition for DaaS systems in general,
-	      but applying such strategies under certain conditions might be plausible.
-	      By using highly available clusters within our backend services
-	      redundancy can be achieved by using the corresponding API endpoints
-	      previously described and by parameterizing them appropriately.
 \end{itemize}
+\begin{figure}
+	\centering
+	\begin{adjustbox}{width=\textwidth}
+		\begin{tikzpicture}[every node/.style={minimum height=1.5cm, align=center}]
+			% Node
+			\node[draw, minimum width=4cm, fill=gray!10] (front0) at (-4,-3.0) {DESIGN Frontend};
+			\node[draw, minimum width=4cm, fill=gray!30] (back0) at (-4,-4.5) {DESIGN Backend};
+			\node[draw, minimum width=4cm, fill=gray!50] (inf0) at (-4,-6.0) {DESIGN Infrastructure};
+			\node[draw, minimum width=4cm, fill=gray!10] (db0) at (0,-3.0) {Database};
+			\node[draw, minimum width=4cm, fill=gray!30] (auth0) at (0,-4.5) {Authentication};
+			\node[draw, minimum width=4cm, fill=gray!50] (fs0) at (0,-6.0) {Storage};
+			% Frontends
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (front1) at (-9,-2.5) {Frontend 1};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (front2) at (-9,-3.0) {Frontend 2};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (front3) at (-9,-3.5) {Frontend 3};
+			\draw[->, dashed, thick, black] (front0.west) -- (front1.east);
+			\draw[->, dashed, thick, black] (front0.west) -- (front2.east);
+			\draw[->, dashed, thick, black] (front0.west) -- (front3.east);
+			% Backends
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (back1) at (-9,-4.0) {Backend 1};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (back2) at (-9,-4.5) {Backend 2};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (back3) at (-9,-5.0) {Backend 3};
+			\draw[->, dashed, thick, black] (back0.west) -- (back1.east);
+			\draw[->, dashed, thick, black] (back0.west) -- (back2.east);
+			\draw[->, dashed, thick, black] (back0.west) -- (back3.east);
+			% Instances
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (inst1) at (-9,-5.5) {Instances 1};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (inst2) at (-9,-6.0) {Instances 2};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (inst3) at (-9,-6.5) {Instances 3};
+			\draw[->, dashed, thick, black] (inf0.west) -- (inst1.east);
+			\draw[->, dashed, thick, black] (inf0.west) -- (inst2.east);
+			\draw[->, dashed, thick, black] (inf0.west) -- (inst3.east);
+			% DB
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (db1) at (5,-2.5) {Database 1};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (db2) at (5,-3.0) {Database 2};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!10] (db3) at (5,-3.5) {Database 3};
+			\draw[->, dashed, thick, black] (db0.east) -- (db1.west);
+			\draw[->, dashed, thick, black] (db0.east) -- (db2.west);
+			\draw[->, dashed, thick, black] (db0.east) -- (db3.west);
+			% Auth
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (auth1) at (5,-4.0) {Authentication 1};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (auth2) at (5,-4.5) {Authentication 2};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!30] (auth3) at (5,-5.0) {Authentication 3};
+			\draw[->, dashed, thick, black] (auth0.east) -- (auth1.west);
+			\draw[->, dashed, thick, black] (auth0.east) -- (auth2.west);
+			\draw[->, dashed, thick, black] (auth0.east) -- (auth3.west);
+			% Fs
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (fs1) at (5,-5.5) {Storage 1};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (fs2) at (5,-6.0) {Storage 2};
+			\node[draw, minimum width=4cm,minimum height=0.5cm, fill=gray!50] (fs3) at (5,-6.5) {Storage 3};
+			\draw[->, dashed, thick, black] (fs0.east) -- (fs1.west);
+			\draw[->, dashed, thick, black] (fs0.east) -- (fs2.west);
+			\draw[->, dashed, thick, black] (fs0.east) -- (fs3.west);
+			% Arrows Bottom
+			\draw[<->, thick, black] (-11,-7.5) -- (-2.1,-7.5)node[midway, below, font=\footnotesize] {User-Facing Components};
+			\draw[<->, thick, black] (-1.9,-7.5) -- (7,-7.5)node[midway, below, font=\footnotesize] {Service Components};
+		\end{tikzpicture}
+	\end{adjustbox}
+	\caption{DESIGN Component Redundancy}\label{service_redundancy}
+\end{figure}
 
@@
 within the related request or response.
-This enables arbitrarily distributed configurations as long as request
+As shown in figure~\ref{message_routing}, this enables arbitrarily distributed configurations as long as request
 and response messages reach their desired target eventually
@@
 Any other component may reside elsewhere within the DaaS network.
+
+\begin{figure}
+	\centering
+	\begin{adjustbox}{width=\textwidth}
+		\begin{tikzpicture}[every node/.style={minimum width=4cm, minimum height=1cm, align=center}]
+			% Frontends
+			\node[draw,minimum height=0.5cm, fill=gray!10] (front1) at (-4,0.3) {Frontend 1};
+			\node[draw,minimum height=0.5cm, fill=gray!10] (front2) at (0,0.3) {Frontend 2};
+			\node[draw,minimum height=0.5cm, fill=gray!10] (front3) at (+4,0.3) {Frontend 3};
+			% Backends
+			\node[draw,minimum height=0.5cm, fill=gray!30] (back1) at (-4,1.1) {Backend 1};
+			\node[draw,minimum height=0.5cm, fill=gray!30] (back2) at (+0,1.1) {Backend 2};
+			\node[draw,minimum height=0.5cm, fill=gray!30] (back3) at (+4,1.1) {Backend 3};
+			% Instances
+			\node[draw,minimum width=12cm, minimum height=1cm, fill=gray!80] (instdomain) at (+0,2.1) {};
+			\node[draw,minimum height=0.5cm, fill=gray!80] (inst) at (+0,2) {Instance};
+			% Message
+			\node[draw,minimum height=0.5cm, fill=gray!50] (message) at (0,-0.75) {Message};
+			% Arrows instance
+			\draw[->, thick, black] (inst.west) -- ++(-3,0) node[midway, above=-5pt, font=\scriptsize] {Switch to host 1};
+			\draw[->, thick, black] (inst.east) -- ++(3,0) node[midway, above=-5pt, font=\scriptsize] {Switch to host 3};
+			% Arrows backend
+			\draw[->, thick, black] (back1.north) -- ++(0,0.25);
+			\draw[->, thick, black] (back2.north) -- ++(0,0.25);
+			\draw[->, thick, black] (back3.north) -- ++(0,0.25);
+			% Arrows frontend
+			\draw[->, thick, black] (front1.north) -- (back1.south);
+			% \draw[->, thick, black] (front1.north) -- (back2.south);
+			% \draw[->, thick, black] (front1.north) -- (back3.south);
+			% \draw[->, thick, black] (front2.north) -- (back1.south);
+			\draw[->, thick, black] (front2.north) -- (back2.south);
+			% \draw[->, thick, black] (front2.north) -- (back3.south);
+			% \draw[->, thick, black] (front3.north) -- (back1.south);
+			% \draw[->, thick, black] (front3.north) -- (back2.south);
+			\draw[->, thick, black] (front3.north) -- (back3.south);
+			% Arrows Message
+			\draw[->, thick, black] (message.north) -- (front1.south);
+			\draw[->, thick, black] (message.north) -- (front2.south);
+			\draw[->, thick, black] (message.north) -- (front3.south);
+		\end{tikzpicture}
+	\end{adjustbox}
+	\caption{Message Oriented Routing.}\label{message_routing}
+\end{figure}
 

### 2024-05-02 paper.tex
Improved Image 1 and merged section 5 and 6

@@
 Section~\ref{sec:AnalysisPossibleComponents} presents the most relevant challenges we faced, analyses our options for handling these, and describes the solution we found.
-Finally, section~\ref{sec:Conclusions} discusses conclusions, and section~\ref{sec:NextSteps} includes directions for future work.
+Finally, section~\ref{sec:Conclusions} discusses conclusions and includes directions for future work.
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=.75\linewidth]{images/DaaS_DESIGN_Architecture_v12_english.pdf}
+	\caption{Architecture of the novel DaaS DESIGN~\cite{OJCC_2023v8i1n01_Baun}.} 
+	\label{figure_architecture}
+\end{figure}
+
 The open-source server virtualization platform Proxmox VE runs bare-metal and offers virtual machines with the Kernel-based Virtual Machine (KVM) and container-based virtualization via LXC Linux container.
@@
 % TODO: Die letzten 3 Abschnitte noch überarbeiten und präzisieren
-
-\begin{figure}
-	\centering
-	\includegraphics[width=.9\textwidth]{images/DaaS_DESIGN_Architecture_v11_english.pdf}
-	\caption{Architecture of the novel DaaS DESIGN~\cite{OJCC_2023v8i1n01_Baun}.} \label{figure_architecture}
-\end{figure}
 
@@
 
-\section{Conclusion}
+\section{Conclusion and Outlook}
 \label{sec:Conclusions}
@@
 
-\section{Outlook}
-\label{sec:NextSteps}
+% \section{Outlook}
+% \label{sec:NextSteps}
 

### 2024-05-03 paper.tex
Some language improvements

@@
 \begin{abstract}
-	Desktop-as-a-Service (DaaS) is a cloud service category with a tremendous potential for improving the everyday work of running and using applications.
+Desktop-as-a-Service (DaaS) is a cloud service category with tremendous potential for improving the everyday work of running and using applications.
 However, existing DaaS solutions have several drawbacks, which have limited the popularity of this service category in the past.
-Our novel DaaS solution called DESIGN not only overcomes the limitations of existing solutions but also offers significant benefits.
-It has the potential to revolutionize the way we use applications, supporting non-modified Linux and Windows applications with just a browser on the client side.
+Our novel DaaS solution, DESIGN, overcomes existing solutions' limitations and offers significant benefits.
+It can potentially revolutionize how we use applications, supporting non-modified Linux and Windows applications with just a browser on the client side.
 This paper describes some of the knowledge we gained during the development because of the challenges we faced and the solutions we found.
@@
 This paper discusses challenges during the development and implementation of a novel Desktop-as-a-Service (DaaS) solution called DESIGN that enables the deployment and usage of unmodified Linux and Windows applications in the same way as web applications and can run inside public resources but can also be deployed in a private context.
-Since all interaction with DESIGN and the applications that reside inside is done via a user's web browser, the users can use any client, no matter what hardware or host operating system it is compromised of.
+Since all interaction with DESIGN and the applications inside is done via a user's web browser, the users can use any client, no matter what hardware or host operating system it is compromised of.
 A browser is the only software component required to use the novel DaaS service.
 
-Undoubtedly, the development of a feature-rich DaaS system is a complex task, riddled with numerous obstacles and challenges.
-This paper meticulously addresses these challenges, categorizing them under compatibility, performance, stability, and usability.
-It not only brings these obstacles to light but also presents the solutions we devised to overcome them, along with the valuable lessons we learned in the process.
+Undoubtedly, developing a feature-rich DaaS system is a complex task, riddled with numerous obstacles and challenges.
+This paper meticulously categorizes these challenges into compatibility, performance, stability, and usability.
+It brings these obstacles to light and presents the solutions we devised to overcome them, along with the valuable lessons we learned in the process.
 

### 2024-05-03 paper.tex
Some language improvements

@@
 This paper meticulously categorizes these challenges into compatibility, performance, stability, and usability.
-It brings these obstacles to light and presents the solutions we devised to overcome them, along with the valuable lessons we learned in the process.
+It reveals these obstacles and presents the solutions we devised to overcome them, along with the valuable lessons we learned in the process.
 
@@
 The paper's authors conclude that Guacamole, in combination with the Remote Desktop Protocol (RDP), is the best solution.
-The authors evaluated the lag time for remote audio playback in an Internet scenario by measuring the time between the automated initiation of the live streaming using the media player software VLC and the arrival of the music data by analyzing network traffic via monitoring software Wireshark.
+The authors evaluated the lag time for remote audio playback in an Internet scenario by measuring the time between the automated initiation of the live streaming using the media player software VLC and the arrival of the music data by analyzing network traffic via the monitoring software Wireshark.
 When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms.
@@
 One emphasis of the paper is comparing the transfer rate of the different protocols for three scenarios.
-In the Office Software Suite scenario, the transfer rate is measured and compared when performing several different tasks like opening and editing a document (typing), loading an image, and saving a document.
+In the Office Software Suite scenario, the transfer rate is measured and compared when performing several tasks like opening and editing a document (typing), loading an image, and saving a document.
 The web browsing scenario covers requesting three different web pages, and for the video streaming scenario, the same YouTube video file was viewed at different resolutions.
@@
 It is important to understand that the protocols RDP and VNC are implemented in open-source projects and proprietary solutions for many different operating systems.
-Depending on the different quality levels of the many existing implementations, it is difficult to generalize about the performance of the protocols, as individual implementations may have better or worse performance characteristics than expected.
+Depending on the various quality levels of the many existing implementations, it is difficult to generalize about the performance of the protocols, as individual implementations may have better or worse performance characteristics than expected.
 
@@
 
-Exporting the graphical user interface of Linux and Windows applications, is possible using the a protocols VNC or RDP.
+Exporting the graphical user interface of Linux and Windows applications is possible using VNC or RDP protocols.
 Free server implementations exist for both operating system families.
-Since the focus of our DaaS solution is to export only the application's graphical user interface and not full desktops -- a feature not all server implementations include -- is the number of available projects and solutions limited.
+Since our DaaS solution focuses on exporting only the application's graphical user interface and not full desktops -- a feature not all server implementations include -- the number of available projects and solutions is limited.
 For Linux containers, the free software projects x11vnc and Xrdp can be used in the DESIGN DaaS.
@@
 
-One of the main goals of DESIGN is to allow all interactions to be carried out solely by using a browser.
+One of the main goals of DESIGN is to allow all interactions to be carried out solely using a browser.
 Thus, a broker or proxy software that mediates between the VNC or RDP implementation and the browser is required.
-The most advanced solution available is the free software project Apache Guacamole which supports VNC and RDP both and includes further relevant features regarding sound and printing.
+The most advanced solution available is the free software project Apache Guacamole, which supports VNC and RDP and includes further relevant features regarding sound and printing.
 
@@
 
-When starting a virtual machine or container instance, it automatically includes an SSH service.
+Starting a virtual machine or container instance automatically includes an SSH service.
 Our self-developed instance component waits until the SSH service is ready to satisfy requests.
-In case applications are to be started or modifications are to be carried out, the instance component sends these requests to the virtual machine or container.
+If applications are to be started or modifications are to be carried out, the instance component sends these requests to the virtual machine or container.
 Modification requests include resolution changes and fetching host status information.
 
-In general, the self-developed instance component allows users to carry out essential administration and usage tasks that the operating system does not offer via a generalized API.
+The self-developed instance component generally allows users to carry out essential administration and usage tasks that the operating system does not offer via a generalized API.
 
@@
 While developing the novel DaaS platform DESIGN, we faced multiple obstacles and challenges.
-Each of these issues belongs to one of these four categories: compatibility, performance, stability, and usability.
-This section presents the issues we were confronted with, the possible approaches we analyzed, and the solutions we chose.
+Each issue belongs to one of four categories: compatibility, performance, stability, and usability.
+This section presents the issues we faced, the possible approaches we analyzed, and the solutions we chose.
 
@@
 Examples worth mentioning in this context include eyeOS and SilveOS.
-The project eyeOS~\cite{liu2012research,vidyabanu2011implementation} which simulates a desktop environment using HTML, the scripting language PHP, JavaScript, and the MySQL database management system.
+The project eyeOS~\cite{liu2012research,vidyabanu2011implementation} simulates a desktop environment using HTML, the scripting language PHP, JavaScript, and the MySQL database management system.
 The project SilveOS~\cite{garmpis2016design} mimics the look and feel of a Windows desktop in the browser using the discontinued application framework Microsoft Silverlight.
@@
 A new DaaS is only accepted by a broad user group if it enables the integration of native applications in an unchanged manner.
-It is also desirable that a DaaS supports Linux and Windows applications both to enable flexible working and make the operation of a virtualization solution on the desktops unnecessary.
+It is also desirable that a DaaS supports Linux and Windows applications both to allow flexible working and make the operation of a virtualization solution on the desktops unnecessary.
 Ideally, it should also be possible to run native Mac OS applications.
@@
 However, Guacamole supports sound using PulseAudio\footnote{On Linux operating systems, sound and microphone devices are accessed via sound servers like PulseAudio and PipeWire, which interact with the sound interface present in the kernel, e.g., Advanced Linux Sound Architecture (ALSA) or Open Sound System (OSS)} as a workaround solution.
-VNC also does not offer file transfer and access to external storage media, but as a workaround solution, Guacamole supports file transfer file transfer via the SSH File Transfer Protocol (SFTP) which is an extension of the Secure Shell protocol (SSH).
+VNC also does not offer file transfer and access to external storage media.
+Still, as a workaround solution, Guacamole supports file transfer file transfer via the SSH File Transfer Protocol (SFTP), which is an extension of the Secure Shell protocol (SSH).
 
@@
 Performance, from our viewpoint, stands out as the most crucial system property,
-implying influence of other  relevant properties,
+implying the influence of other  relevant properties,
 such as user experience, system responsiveness, and overall efficiency.
@@
 \subsubsection{Infrastructure and Platform}
-In general can be said that the chosen infrastructure, platform,
+In general, it can be said that the chosen infrastructure, platform,
 and utilized software combined may have the most impact on the overall resource consumption
 and are therefore of major importance.
-As a DaaS system combines all aspects of IaaS-, PaaS, and SaaS-based computation
-each individual performance characteristics has to be taken into account.
+As a DaaS system combines all aspects of IaaS-, PaaS, and SaaS-based computation, each performance characteristic must be considered.
 
@@
 Depending on the use case, influential properties
-such as multi-tenancy, provision strategies or the virtualized data type
+such as multi-tenancy, provision strategies, or the virtualized data type
 might be utilized to varying degrees, resulting in overall
@@
 coverage of a broad range of potential use cases is generally more favorable
-than solving a particular use case in the most efficient way.
+than solving a particular use case most efficiently.
 Therefore, we consider all user-facing desktop applications
@@
 rather than more complex distributed systems
-for which a tailored environments might be more suitable.
-
-Figure~\ref{hosting_scenarios} visualizes the available hosting startegies for our platform.
+for which tailored environments might be more suitable.
+
+Figure~\ref{hosting_scenarios} visualizes the available hosting strategies for our platform.
 We aim to support any Windows or Linux-based application in virtual machines
-and support any Linux- and Wine-based applications for container environments.
+and any Linux- and Wine-based applications in container environments.
 

### 2024-05-03 paper.tex
Some language improvements

@@
 and provide additional access to external devices like printers or USB sticks.
-In cases involving for example video content transmission,
+In cases involving, for example, video content transmission,
 network bandwidth often becomes the main bottleneck,
@@
 through protocol-inherent tuning parameters and compression algorithms.
-Our solution therefore offers a range of tuning options
-in order to adapt to different use cases
-and empowering users to choose their own tailored solution.
+Our solution, therefore, offers a range of tuning options
+to adapt to different use cases
+and empowering users to choose their own tailored solutions.
 
 We utilize the Guacamole desktop gateway implementation,
-supporting widely-used frame-buffer protocols like VNC and RDP,
+supporting widely used frame-buffer protocols like VNC and RDP,
 along with relevant tuning options.
@@
 As depicted in figure~\ref{websocket_proxy},
-the proxy component recives HTTP requests from the guacamole.js webclient
-and forwarded by the backend by using regular WebSocket opcodes (Guacamole protocol).
+the proxy component receives HTTP requests from the guacamole.js web client
+and is forwarded by the backend by using regular WebSocket opcodes (Guacamole protocol).
 Such communication is forwarded to the Guacamole daemon (guacd)
-which in turn communicates to each instance
-by using the frame-buffer protocol of choice (VNC or RDP).
+which in turn communicates to each instance using the frame-buffer protocol of choice (VNC or RDP).
 
@@
 requires consideration of additional factors besides the frame-buffer protocols itself.
-Authentication mechanisms for example are particularly important in a DaaS context,
+Authentication mechanisms, for example, are essential in a DaaS context,
 as users expect seamless authentication across the whole system
@@
 To mitigate the risk of unintended system manipulation, Guacamole was integrated
-in a proxied approach and by using websockets.
-This ensures that credentials of a generated instance remain secret to the host system
+using a proxied approach and WebSockets.
+This ensures that the credentials of a generated instance remain secret to the host system
 while still providing authorized, passwordless access to the hosted application.
@@
 such as dynamic desktop resizing.
-While Guacamole offers support for that by default in RDP,
-support is completely missing for the VNC protocol.
-Nevertheless, the proxied approach enables to intercept all exchanged websocket opcodes
+While Guacamole offers support for that by default in RDP, support for the VNC protocol is completely missing.
+Nevertheless, the proxied approach enables intercepting all exchanged WebSocket opcodes
 and by listening for size requests
@@
 This includes metadata collected and processed by essential components
-and involves direct exchange of raw data with object instances.
+and involves the direct exchange of raw data with object instances.
 
@@
 Additionally, a web server is integrated to provide relevant API endpoints
-for frontend communication and websocket endpoints for viewer communication.
+for frontend communication and WebSocket endpoints for viewer communication.
 
@@
 when handling high throughputs or large amounts of concurrent applications
-requiring low latency access and especially if all necessary backend components
+requiring low latency access, especially if all necessary backend components
 are served from the same hardware.
 Architecturally, all relevant components
-must therefore be designed for distribution across different hosting systems.
-However, retaining certain components directly on the host system
+must be designed for distribution across different hosting systems.
+However, retaining specific components directly on the host system
 may still be required to ensure low-latency access.
 
-From our observations, components requiring realtime and low-latency performance,
+From our observations, components requiring real-time and low-latency performance,
 should physically be kept as close to the core hosting system as possible
@@
 or at least be connected to the same physical network
-in order to minimize network latencies.
-
-Regarding other components such as database access,
+to minimize network latencies.
+
+Regarding other components, such as database access,
 authentication providers, and filesystems,
 maintaining a close distance from the host is advisable,
-but this depends largely on the total number of users
+but this depends mainly on the total number of users
 as well as their specific use cases.
@@
 to be split into logical units that can be distributed across dedicated systems.
-Each backend service receives dedicated configuration
+Each backend service receives a dedicated configuration
 and can be accessed via hostname and port tuples.
@@
 but administrators are free to use custom third-party toolchains,
-such as GPU virtualization, for vertical scaling on a per-host basis.
+such as GPU virtualization for vertical scaling on a per-host basis.
 Although vertical scaling within a host
@@
 and utilized to distribute instances across a pool of connected nodes.
-For virtual machines, Proxmox implements such mechanisms
-as part of its `Proxmox VE High Availability (HA)' cluster.
-For containers, similar concepts can be integrated with any backend service,
+Proxmox implements such mechanisms
+for virtual machines as part of its `Proxmox VE High Availability (HA)' cluster.
+Similar concepts can be integrated for containers with any backend service,
 such as Kubernetes.
@@
 As shown in figure~\ref{scale_objects_horizontal},
-we provide a globally maintained state of data, authentication and storage
+we provide a globally maintained state of data, authentication, and storage
 which is accessible either by horizontally scaled backends and frontends
@@
 	      Both components contribute to redundancy
-	      by distributing requests and responses to their respecitive recipients
+	      by distributing requests and responses to their respective recipients
 	      instead of processing them on their own.
@@
 	      The database is integrated as a backend service
-	      and is therefore independent of surrounding components and contexts.
+	      and is, therefore, independent of surrounding components and contexts.
 	      It can be distributed to a dedicated system if necessary.
@@
 	      It can also be targeted by standardized load-balancing mechanisms
-	      and remains fully agnostic of surrounding components
+	      and remains entirely agnostic of surrounding components
 	      or calling contexts.
@@
 	      In our final version,
-	      a customizable amount of redundant systems may be configured.
+	      a customizable number of redundant systems may be configured.
 	\item \textbf{Distributed Filesystem:}
@@
 	      Hosting DaaS instances directly from the distributed block storage
-	      enables migration and redundancy of instances or in other words, horizontal scaling.
+	      enables migration and redundancy of instances or, in other words, horizontal scaling.
 \end{itemize}
@@
 including the backend service hosting relevant instances as well as the described
-websocket proxy connected to the guacamole daemon and the user.
+WebSocket proxy connected to the guacamole daemon and the user.
 Any other component may reside elsewhere within the DaaS network.
@@
 and single-app functionality can be easily obtained
-by using a custom window manager such as for example qtile
+by using a custom window manager such, as for example, qtile
 ~\footnote{\url{https://qtile.org/}}.
@@
 \footnote{\url{https://issues.apache.org/jira/browse/GUACAMOLE-1196}},
-we implemented a separate mechanism using a websocket proxy
+we implemented a separate mechanism using a WebSocket proxy
 and direct interaction with each instance.
@@
 
-Another challenge was managing RDP sessions which are typically tied to actual users,
-unlike VNC sessions which are typically authenticated against
+Another challenge was managing RDP sessions, which are typically tied to actual users,
+unlike VNC sessions, which are typically authenticated against
 a server-maintained user registry.
-Utilizing the websocket proxy, the backend could pass credentials
+Utilizing the WebSocket proxy, the backend could pass credentials
 to the server without revealing them to the user.
 
-Essentially, Guacamole with its protocol implementations and additional features like audio and printing,
+Essentially, Guacamole, with its protocol implementations and additional features like audio and printing,
 provided the fundamental base for our core capabilities.
 
-Finally mentioned, certain instances might not be fully accessible during runtime
-due to hardware malfunctions, inappropriate system configurations for simply during system updates.
-Therefore, as a failover mechanism additional logic had to be added
-in order to provide reasonable debug views and recovery mechanisms,
-e.g.~to recreate instances from a specific snapshot or configuration state.
-
-A crucial lesson learned was therefore to realize that
+Finally, certain instances might not be fully accessible during runtime
+due to hardware malfunctions and inappropriate system configurations during system updates.
+Therefore, as a failover mechanism, additional logic had to be added
+to provide reasonable debug views and recovery mechanisms,
+e.g., recreate instances from a specific snapshot or configuration state.
+
+A crucial lesson learned was, therefore, to realize that
 while individual problems can mostly be solved
-using established paradigms, services or libraries with only slight modifications,
-the real challenge for a full-fledged DaaS system lies in it's abilities
-to combining such isolated solutions appropriately
+using established paradigms, services, or libraries with only slight modifications,
+the real challenge for a full-fledged DaaS system lies in its abilities
+to combine such isolated solutions appropriately
 and for each specific infrastructure and platform configuration.
@@
 
-This work was funded by the Federal Ministry for Economic Affairs and Climate Action
+The Federal Ministry for Economic Affairs and Climate Action funded this work
 ('\textsl{Bundesministerium f\"ur Wirtschaft und Klimaschutz}')

### 2024-05-04 paper.tex
Minor changes (missing spaces, remove comments, add TODO's)

@@
 
-% TODO: Christian (1)
-% TODO: Johannes (2)
-
 \section{Related Work}
 \label{sec:relatedworkAbschnitt}
-% TODO:  Introduce: Infratstructure-as-a-Service  and others (IaaS, PaaS, SaaS, DaaS)
-% TODO: Christian (1)
-% TODO: Johannes (2)
 
@@
 \end{itemize}
-
+% TODO: Mention hardware-accelerated framebuffer protocols like \cite{hanimplementation} \cite{darcy2022latency}
+% TODO: More scientific papers: \cite{di2021network} \cite{carrascosa2022cloud}
 For every deployment scenario or cloud service category possible, users expect a latency value that is not annoying when interacting with the service and its applications.
@@
 % 
-% TODO: Christian (1)
-% TODO: Johannes (2)
-
 Figure~\ref{figure_architecture} shows the architecture of our novel DaaS solution DESIGN that can run unmodified Linux and Windows applications in Linux containers (Linux and Windows applications compatible with Wine) and virtual machines (Windows applications incompatible with Wine).
-
 \begin{figure}
@@
 
-% TODO: Christian (1)
-% TODO: Johannes (2)
-
 While developing the novel DaaS platform DESIGN, we faced multiple obstacles and challenges.
@@
 \subsection{Compatibility}
-% 
-% TODO: Christian (1)
-% TODO: Johannes (2)
 
@@
 \subsection{Performance}
-% Initial brainstorm:
-% + Welche Performance-Paramter / Kriterien sind hier relevant für die Anwendung / Nutzung
-% + Fokus: Bottlenecks, Möglichkeiten der Performance-Steigerung, Grenzen/Limits, die man akzeptieren muss
-% + Latenz hängt auch immer von Entfernung und Übertragungsmedium ab.
-% + Skalierbarkeit: Stand und Möglichkeite
-% + Ceph-Dateisystem
-% + Datenbank SQlite aktuell. Perspektivisch MariaDB
-% + Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host.
-%
-% Structure:
-% performance of:
-% - our serice itself
-% - backend services
-% - Viewer performance
-% Limiting Factors
-% - Hosting
-% - Network
-% - Internal data and communication
-% Chosen techniques/solutions/protocols:
-% - Filesystem (Ceph)
-% - Database (sqlite, mariadb)
-% - Guacamole (Proxy approach)
-% Similar_Examples:
-% - Cloud Gaming efforts 
-%    - technical solutions, 
-%    - latencies, 
-%    - hardware support and integration)
-%
-% TODO: Johannes (1)
-% TODO: Christian (2)
-% TODO: Christian: Latenz bei Cloud-Gaming Erfahrungen
-% TODO: 
-%       - Include overall takeaway messages somewhere before:
-%           - IaaS, PaaS and SaaS combined is defined as DaaS
-%           - Define terms: Backend-service, Host-System, Instance
-%           - We want a stable and usable system first, performance comes second
-%           - Low effort to use and integrate is the most crucial aspect!!!
-%           - No "Gamestation", but all modern applications shall be suported
-%               - "Regular" office apps (Spreadsheets, Editor, Email, Filemanager)
-%               - 3D modelling (Blender, Mayo, etc.)
-%               - Audio/Video Editing? (kdenlive, Adobe Premiere, OBS?)
-%               - Services (Webserver, Emailserver, Fileserver, etc.)
-%           - If several solutions are plausible, implement all and let the user decide
-%           - Give definitions for all specified daas-usecase
-%       - More references
-%       - Illustrate each subsection
-%       - Typos, grammar, etc.
-
 To describe the mission-critical performance aspects of the discussed system,
@@
 drawing from our experiences obtained during the project.
-
-% From our point of view, the three most critical performance aspects
-% for Daas-Systems are:
-% \begin{itemize}
-% 	\item Infrastructure and Platform
-% 	\item Network Latency and Bandwidth
-% 	\item Internal Processing and Communication
-% \end{itemize}
-% Since each of them has their own characteristics,
-% specific behavior and requirements, implying a potential impact
-% on the overall system performance and resource utilization,
-% we describe them individually and in more detail
-% in the following separate subsections.
 
@@
 might be utilized to varying degrees, resulting in overall
-better or worse performing systems ~\cite{zhong2010virtualization}.
+better or worse performing systems~\cite{zhong2010virtualization}.
 
@@
 \subsection{Stability}
-% TODO: Johannes (1)
-% TODO: Christian (2)
-% TODO: Johannes: Beschreibung zur Position der Komponenten, Möglichkeiten der Parallelisierung und Einschränkungen
-%
-% Was ist hier relevant?
-% Verfügbarkeit des Service und der Benutzerdaten
-% Redundante Komponenten (Hardware und Software)
-% Welche Komponenten kann man sinnvoll redundant vorhalten?
-% Was sind die Vor- und Nachteile?
-% Auch Kosten?
-% Wo sehen wir aktuell in unserer Architektur Risiken für Dienstausfall und Datenverlust?
-% Features von Proxmox
-% Ceph-Dateisystem
-% Datenbank SQlite aktuell. Perspektivisch MariaDB
-% Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host. Proxmox kann sich darum kümmern?!
-% Docker haben/nicht haben / Vor- und Nachteile
-% Hier auch auf die Skalierbarkeit beziehen
-% Structure:
-%	- Stability
-%		- Redundancy
-%		- Scalability
-%	- Backend-services
-%		- Database
-%		- Authentication
-%		- Distributed Filesystem
-%		- Container/VM (Proxmox, Kubernetes?) 
-%		- Instance error => Recreate
-%	- Considerations:
-%		- TODO: Mention that term: Divide and conquer
-%		- Requirement: Distributable components
-%	- Risks:
-%		- Filesystem error
-%		- Database error
-%		- Authentication error
-%		- Infrastructure error (Container/VM)
-%	- Benefits/Drawbacks:
-%		- Distribution of components is distribution of load => Performance
-%		- Eliminates bottlenecks
-%		- Errors affect only errornous component
-
 Stability in DaaS applications is especially important
@@
 and single-app functionality can be easily obtained
-by using a custom window manager such, as for example, qtile
-~\footnote{\url{https://qtile.org/}}.
+by using a custom window manager such,
+as for example, qtile~\footnote{\url{https://qtile.org/}}.
 
@@
 but do not seem useful for designing and implementing a novel DaaS because
-they do not meet all feature requirements.~\cite{OJCC_2023v8i1n01_Baun}
+they do not meet all feature requirements~\cite{OJCC_2023v8i1n01_Baun}.
 
@@
 leading to misaligned layouts and a poor user experience.
-To address this until such features are available in Guacamole
-\footnote{\url{https://issues.apache.org/jira/browse/GUACAMOLE-1196}},
+To address this until such features are available in
+Guacamole~\footnote{\url{https://issues.apache.org/jira/browse/GUACAMOLE-1196}},
 we implemented a separate mechanism using a WebSocket proxy
@@
 
-% \section{Outlook}
-% \label{sec:NextSteps}
-
 For future work, we plan to do a performance analysis that includes an evaluation of different user behaviors for various popular applications and realistic deployment and utilization scenarios.
@@
 % ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
 \bibliographystyle{splncs04}
 \bibliography{biblio}
-%
-
-% \bibliography{\jobname_no_dashes_at_repeated_names,references}{}
 

### 2024-05-07 paper.tex
More proof read, shortening and references

@@
 
-Several works in literature have analyzed the latency of Cloud gaming services that have, in some aspects, similar requirements to DaaS offerings.
-According to Lampe et al.~\cite{lampe2014assessing}, the latency of locally executed games is compromised of the following components:
-
+Several works in literature have analyzed
+the latency of Cloud gaming services that have, in some aspects,
+similar requirements to DaaS offerings.
+According to Lampe et al.~\cite{lampe2014assessing},
+the latency of locally executed games is compromised of the following components:
 \begin{itemize}
@@
 
-In cloud gaming scenarios, the following components increase latency further:
-
+In cloud gaming scenarios the following components increase latency further:
 \begin{itemize}
@@
 \end{itemize}
-% TODO: Mention hardware-accelerated framebuffer protocols like \cite{hanimplementation} \cite{darcy2022latency}
-% TODO: More scientific papers: \cite{di2021network} \cite{carrascosa2022cloud}
-For every deployment scenario or cloud service category possible, users expect a latency value that is not annoying when interacting with the service and its applications.
-Choy et al.~\cite{ChoyWongSimonRosenberg2012}, Claypool et al.~\cite{claypool2010latency}, and Jarschel et al.~\cite{jarschel2011evaluation} consider a maximum tolerable latency, based on subjective tests, being around 100\,ms.
-This relatively low value sets the demand for a fast network interaction between client and service and limits their maximum distance.
-As mentioned in Clincy et al.~\cite{clincy2013subjective}, The former cloud gaming service provider OnLive specified in the year 2010 that the distance between the service provider and the consumer should not be greater than 1000 miles (approx.
-1600\,km) because otherwise, the round-trip communications delay time between will be too long for video games.
 
-On first thought, the latency requirements of a modern computer game are hardly the same as those of a desktop application.
-However, just as with computer games, the lagging operation of desktop programs has a negative impact on the user experience, and with video conferencing or other graphics-intensive applications, the utilization of resources and latency requirements are very similar to computer games.
+For every deployment scenario or cloud service category possible,
+users expect a latency value that is not annoying
+when interacting with the service and its applications.
+Choy et al.~\cite{ChoyWongSimonRosenberg2012},
+Claypool et al.~\cite{claypool2010latency},
+and Jarschel et al.~\cite{jarschel2011evaluation}
+consider a maximum tolerable latency, based on subjective tests,
+being around 100\,ms. This relatively low value sets the demand for a fast network interaction
+between client and service and limits their maximum distance.
+As mentioned in Clincy et al.~\cite{clincy2013subjective},
+The former cloud gaming service provider OnLive specified in the year 2010
+that the distance between the service provider and the consumer
+should not be greater than 1000 miles (approx. 1600\,km) because otherwise,
+the round-trip communications delay time between will be too long for video games.
+A more in-depth comparison of
+commercially available cloud gaming services is given
+by Di Domenico et al.~\cite{di2021network}
+or by Carrascosa et al.~\cite{carrascosa2022cloud}.
+Additionally, implied delays and round-trips can be further reduced
+by using more sophisticated transmission protocols
+directly interacting with GPU drivers,
+such as Moonlight~\cite{hanimplementation}~\cite{darcy2022latency}.
+
+On first thought, the latency requirements of a modern computer game
+are hardly the same as those of a desktop application.
+However, just as with computer games,
+the lagging operation of desktop programs has a negative impact on the user experience,
+and with video conferencing or other graphics-intensive applications,
+the utilization of resources and latency requirements are very similar to computer games.
 
@@
 \subsection{Performance}
-To describe the mission-critical performance aspects of the discussed system,
-we initially provide a brief overview of potential challenges
-by outlining bottlenecks and other relevant problem areas.
-This is done in the context of the underlying use case model of DaaS systems.
-Subsequently, we give a more comprehensive description
-of our specific design decisions and tailored solutions.
-
 Performance, from our viewpoint, stands out as the most crucial system property,
@@
 Consequently, we encountered multifaceted challenges
-that potentially intersected across several layers within our architecture.
+that potentially intersected across several layers within our architecture
+and which will be described in the following subsections.
 
-Hence, every component from the core infrastructure
-to all backend services and user-facing elements,
-must meet individual and domain-specific requirements
-to contribute to the overall performance goals of the entire system.
-Thus, we outline key limiting factors that hinder performance,
-along with our experiences and the chosen mitigation strategies.
-
-Given the absence of a generalized measurement system,
-we ranked these factors based on their magnitude of impact,
-drawing from our experiences obtained during the project.
+% To describe the mission-critical performance aspects of the discussed system,
+% we initially provide a brief overview of potential challenges
+% by outlining bottlenecks and other relevant problem areas.
+% This is done in the context of the underlying use case model of DaaS systems.
+% Subsequently, we give a more comprehensive description
+% of our specific design decisions and tailored solutions.
+% Hence, every component from the core infrastructure
+% to all backend services and user-facing elements,
+% must meet individual and domain-specific requirements
+% to contribute to the overall performance goals of the entire system.
+% Thus, we outline key limiting factors that hinder performance,
+% along with our experiences and the chosen mitigation strategies.
+% Given the absence of a generalized measurement system,
+% we ranked these factors based on their magnitude of impact,
+% drawing from our experiences obtained during the project.
 
@@
 over virtual machines when identical programs are tested in both environments
-~\cite{felter2015updated,potdar2020performance,seo2014performance}.
+~\cite{felter2015updated,potdar2020performance}.
 
@@
 such as interrupt handling, memory allocation, and boot time
-~\cite{sergeev2022docker,sulaiman2021comparison}.
+~\cite{sulaiman2021comparison}.
 Additional performance potentials can be leveraged
@@
 with a container-based solution integrating wine
-~\cite{huang2012performance}.  % win inside vm
+~\cite{huang2012performance}.
 
@@
 Our solution, therefore, offers a range of tuning options
-to adapt to different use cases
-and empowering users to choose their own tailored solutions.
+to adapt to different use cases on a per-user basis.
 
@@
 As depicted in figure~\ref{websocket_proxy},
-the proxy component receives HTTP requests from the guacamole.js web client
-and is forwarded by the backend by using regular WebSocket opcodes (Guacamole protocol).
-Such communication is forwarded to the Guacamole daemon (guacd)
-which in turn communicates to each instance using the frame-buffer protocol of choice (VNC or RDP).
+the proxy component receives HTTP requests from guacamole.js
+and is forwarded by the backend through WebSocket opcodes (Guacamole protocol).
+These opcodes are then transmitted to the Guacamole daemon (guacd)
+which in turn communicates to each instance
+by using regular frame-buffer protocols (VNC or RDP).
 
@@
 as users expect seamless authentication across the whole system
-without multiple logins per session.
+without multiple logins.
 Additionally, safeguarding sensitive information like user passwords is crucial
@@
 Another critical aspect directly related to performance and responsiveness
-is how data is internally processed, stored, and communicated
-among all involved components and instances.
-This includes metadata collected and processed by essential components
-and involves the direct exchange of raw data with object instances.
+is how data is internally processed, stored, and communicated.
+This includes metadata collected and processed
+and involves data exchange with all relevant backend services or instances.
 
 Ensuring a satisfying user experience in a DaaS system
-implies specific components in a host system.
-This includes access to all backend services
-used to create containers and virtual machines,
-local or remote databases, authentication services, filesystems, and network devices.
+implies specific components to be available.
+This includes access to all infrastructural backend services,
+a databases, authentication services, filesystems, and network devices.
 Additionally, a web server is integrated to provide relevant API endpoints
-for frontend communication and WebSocket endpoints for viewer communication.
+for frontend communication and WebSocket endpoints for the viewer.
 
@@
 when handling high throughputs or large amounts of concurrent applications
-requiring low latency access, especially if all necessary backend components
-are served from the same hardware.
+and is especially true if all components are served from the same hardware.
 Architecturally, all relevant components
-must be designed for distribution across different hosting systems.
-However, retaining specific components directly on the host system
+must therefore be designed for distribution across different hosts.
+However, keeping specific components directly on the host system
 may still be required to ensure low-latency access.
@@
 to minimize network latencies.
+All other components might be distributed to dedicated systems
+if large user amounts and specific use cases have to be handled
+but keeping them in close physical proximity
+to the host remains prudent in smaller scales.
 
-Regarding other components, such as database access,
-authentication providers, and filesystems,
-maintaining a close distance from the host is advisable,
-but this depends mainly on the total number of users
-as well as their specific use cases.
-On smaller scales, keeping these components
-in close physical proximity to the host remains prudent.
+Allthough we can not yet provide a full-fledged performance analysis,
+we already measured promising results in our local test environment.
+Depending on the exact setup,
+the solution provides
+reasonable round-trip times and delays for each communication direction
+visualized in figure~\ref{perf_internal_communication}.
+In a minimal setup, we measured delays of around 10ms for client communication,
+below 1ms for internal backend communication
+but up to 1000ms for instance communication.
+But these numbers are only rough estimates
+as certain optimization concepts remain to be implemented.
 
@@
 to maintain service continuity, handle growing user loads,
-and transparently recover from failures
-while providing a seamless end-user experience.
+and transparent failure recovery while providing a seamless end-user experience.
 These considerations are integral to our architectural concept,
@@
 
-This definition of stability is supported
-by various architectural and operational strategies,
-including component redundancy, node scalability,
-and a robust backend service framework facilitating communication
-between components and nodes.
-These strategies, along with their drawbacks and benefits,
-will be further explored at a component level
-and within the following subsections.
+Therefore, key strategies, such as component redundancy, node scalability
+as well as the internal communication concept
+will be described in the following subsections.
 
 \subsubsection{Scalability of Components}
-
 Scalability in DaaS environments
 can be approached through two fundamental methods: horizontal and vertical scaling.
-Horizontal scaling, or scaling out, involves adding more machines or instances
+Horizontal scaling (scale out), involves adding more machines or instances
 to manage increased loads.
-In contrast, vertical scaling, or scaling up,
+In contrast, vertical scaling (scale up),
 adds more power (CPU, RAM) to existing machines
@@
 and can be accessed via hostname and port tuples.
-By default, all backend services reside on the same host,
-with tuples derived from the host configuration.
+By default, all backend services reside on the same host.
 In a distributed context, services can be distributed to dedicated remote systems
 to adhere to horizontal scaling principles.
-Relevant tuples can then be derived from any node
-using a common database containing meta information.
 
@@
 	      The database is integrated as a backend service
-	      and is, therefore, independent of surrounding components and contexts.
-	      It can be distributed to a dedicated system if necessary.
+	      and is, therefore, independent of surrounding components and contexts
+	      and can be externalized to a dedicated systems.
 	      Modern database solutions, like MariaDB,
 	      offer rich features for maintaining specific data consistency strategies
-	      such as data replication, clustering, failover mechanisms,
-	      and RAID control.
+	      such as data replication, clustering, failover mechanisms or RAID control.
 	      Currently, redundancies in our database system are not planned
-	      but can be easily applied by users using such database-specific features
+	      but can be easily applied by using such database-specific features
 	      and a modified database adapter.
@@
 	      Authentication within our architecture is implemented as a backend service,
-	      allowing for distribution to any dedicated system.
-	      It can also be targeted by standardized load-balancing mechanisms
-	      and remains entirely agnostic of surrounding components
-	      or calling contexts.
-	      The authentication system maintains its credential registry
-	      in a separate database, along with a customizable set
-	      of user and group privileges.
-	      Additionally, it implements the well-defined OAuth 2.0 standard
-	      to facilitate remote authentication.
-	      In our final version,
-	      a customizable number of redundant systems may be configured.
+	      also allowing for distribution to dedicated systems.
+	      It can be targeted by standardized load-balancing mechanisms
+	      and remains entirely agnostic of calling contexts.
+	      The system maintains its credential registry in a separate database,
+	      along with customizable user and group privileges and implements
+	      the well-defined OAuth 2.0 standard to facilitate remote authentication.
 	\item \textbf{Distributed Filesystem:}
-	      Within our architecture, we utilize the distributed filesystem Ceph,
-	      which offers object, block, and file storage capabilities.
-	      Redundancy in Ceph is implicitly achieved
+	      In our architecture, we utilize the distributed filesystem Ceph,
+	      which offers means of object, block, and file storage.
+	      Redundancy is implicitly achieved
 	      through its self-healing and self-managing features, employing
@@
 	      Data availability and fault tolerance are ensured
-	      by automatically replicating data across different devices.
-	      Our architecture leverages these capabilities
-	      to reliably provide file storage to host systems
-	      and all hosted DaaS instances,
-	      while also offering block storage to the infrastructural backend services.
-	      Hosting DaaS instances directly from the distributed block storage
-	      enables migration and redundancy of instances or, in other words, horizontal scaling.
+	      by automatically replicating data across multiple systems.
+	      Such benefits are leveraged by providing reliable file storage to all instances,
+	      and by providing block storage to the infrastructural backend services.
+	      By doing that, migration and redundancy of instances is achieved
+	      or, in other words, horizontal scaling.
 \end{itemize}
@@
 			% Arrows Bottom
-			\draw[<->, thick, black] (-11,-7.5) -- (-2.1,-7.5)node[midway, below, font=\footnotesize] {User-Facing Components};
-			\draw[<->, thick, black] (-1.9,-7.5) -- (7,-7.5)node[midway, below, font=\footnotesize] {Service Components};
+			\draw[<->, thick, black] (-11,-7.5) -- (-2.1,-7.5) node[midway, above, yshift=-0.5cm, font=\footnotesize] {User-Facing Components};
+			\draw[<->, thick, black] (-1.9,-7.5) -- (7,-7.5) node[midway, above, yshift=-0.5cm, font=\footnotesize] {Service Components};
 		\end{tikzpicture}
@@
 After discussing the principles of scalability and redundancy,
-we finally discuss how communication flows are managed in our distributed environment.
-Our proposed solution is rooted in a message-oriented architecture
+we finally describe how communication flows are managed within our system.
+Our solution is rooted in a message-oriented architecture
 designed to fulfill DaaS properties for scalability and redundancy,
@@
 This implies that any action triggered, whether in the frontend on the client side,
-backend on the host side, or within hosted instances,
-must ensure that the necessary information for execution
-is readily available at the time of processing.
+in the backend on the host side, or within hosted instances,
+must ensure that all necessary information is readily available at the time of processing.
 
@@
 object metadata in a common database,
-and by utilizing a decentralized OAuth component for authentication.
+and by utilizing a decentralized authentication component.
 
@@
 Additionally, any extra information needed by a component must be fully contained
-within the related request or response.
+within the related request.
 As shown in figure~\ref{message_routing}, this enables arbitrarily distributed configurations as long as request
-and response messages reach their desired target eventually
-and as long relevant backend services remain available.
+and response messages reach their desired target eventually.
 
-If that is the case, each distributed DaaS node might only contain essential and time-critical components,
-including the backend service hosting relevant instances as well as the described
-WebSocket proxy connected to the guacamole daemon and the user.
-Any other component may reside elsewhere within the DaaS network.
+If that is the case, each distributed host system might only contain it's essential components,
+including the backend service with all hosted instances
+as well as the described proxy component for direct user interaction.
+Any other component may reside somewhere in the DaaS network.
 
@@
 
-In this paper, we describe an architecture for a novel and feature-rich DaaS that is, by principle, superior to comparable products and approaches because it supports running unmodified Linux and Windows applications.
-Furthermore, we identified characteristic issues -- concerning compatibility, performance, stability, and usability -- during the development and implementation of our DaaS and present solutions to overcome these challenges.
+In this paper, we describe an architecture for a novel
+and feature-rich DaaS that is, by principle,
+superior to comparable products and approaches
+because it supports running unmodified Linux and Windows applications.
+Furthermore, we identified characteristic issues
+-- concerning compatibility, performance, stability, and usability --
+during the development and implementation of our DaaS
+and present solutions to overcome these challenges.
 
-For future work, we plan to do a performance analysis that includes an evaluation of different user behaviors for various popular applications and realistic deployment and utilization scenarios.
+For future work,
+we plan to replace our temporary solution for internal communication
+with a more efficient approach as our preliminary performance tests revealed
+a potential bottlenack in that particular area.
+We also plan to apply certain optimization strategies
+with respect to backend and instance communication.
+We further plan to enhance our yet limited performance testing
+and will provide a full-fledged performance analysis
+of realsitic deployment scnarios and utilization patterns
+which include an evaluation of different user behaviors
+within various popular applications.
 

### 2024-05-07 paper.tex
Add orcid for both authors
Remove another typo

@@
 	Johannes Bouché}
-% \author{Christian Baun\orcidID{0009-0004-9955-3752} \and
-% 	Johannes Bouché\orcidID{TODO}}
-%
+\author{Christian Baun\orcidID{0009-0004-9955-3752} \and
+	Johannes Bouché\orcidID{0009-0009-8003-4019}
+}
+
 \authorrunning{C. Baun, J. Bouché}
@@
 VNC also does not offer file transfer and access to external storage media.
-Still, as a workaround solution, Guacamole supports file transfer file transfer via the SSH File Transfer Protocol (SFTP), which is an extension of the Secure Shell protocol (SSH).
+Still, as a workaround solution, Guacamole supports file transfer via the SSH File Transfer Protocol (SFTP), which is an extension of the Secure Shell protocol (SSH).
 
@@
 and which will be described in the following subsections.
-
-% To describe the mission-critical performance aspects of the discussed system,
-% we initially provide a brief overview of potential challenges
-% by outlining bottlenecks and other relevant problem areas.
-% This is done in the context of the underlying use case model of DaaS systems.
-% Subsequently, we give a more comprehensive description
-% of our specific design decisions and tailored solutions.
-% Hence, every component from the core infrastructure
-% to all backend services and user-facing elements,
-% must meet individual and domain-specific requirements
-% to contribute to the overall performance goals of the entire system.
-% Thus, we outline key limiting factors that hinder performance,
-% along with our experiences and the chosen mitigation strategies.
-% Given the absence of a generalized measurement system,
-% we ranked these factors based on their magnitude of impact,
-% drawing from our experiences obtained during the project.
 

### 2024-05-10 paper.tex
Minor spellings in abstarct and introduction

@@
 \begin{abstract}
-	Desktop-as-a-Service (DaaS) is a cloud service category with tremendous potential for improving the everyday work of running and using applications.
-However, existing DaaS solutions have several drawbacks, which have limited the popularity of this service category in the past.
-Our novel DaaS solution, DESIGN, overcomes existing solutions' limitations and offers significant benefits.
-It can potentially revolutionize how we use applications, supporting non-modified Linux and Windows applications with just a browser on the client side.
-This paper describes some of the knowledge we gained during the development because of the challenges we faced and the solutions we found.
+	Desktop-as-a-Service (DaaS) is a cloud service category
+	with tremendous potential for improving the everyday work
+	of running and using applications.
+	However, existing DaaS solutions have several drawbacks,
+	which have limited the popularity of this service category in the past.
+	Our novel DaaS solution, DESIGN, overcomes existing solutions limitations
+	and offers significant benefits.
+	It can potentially revolutionize how we use applications,
+	supporting non-modified Linux and Windows applications
+	with just a browser on the client side.
+	This paper describes some of the knowledge we gained
+	during the development because of the challenges
+	we faced and the solutions we found.
 
@@
 
-This paper discusses challenges during the development and implementation of a novel Desktop-as-a-Service (DaaS) solution called DESIGN that enables the deployment and usage of unmodified Linux and Windows applications in the same way as web applications and can run inside public resources but can also be deployed in a private context.
-Since all interaction with DESIGN and the applications inside is done via a user's web browser, the users can use any client, no matter what hardware or host operating system it is compromised of.
+This paper discusses challenges during the development
+and implementation of a novel Desktop-as-a-Service (DaaS) solution
+called DESIGN that enables the deployment
+and usage of unmodified Linux and Windows applications
+in the same way as web applications
+and can run inside public resources
+but can also be deployed in a private context.
+Since all interaction with DESIGN and the hosted applications
+is done via a user's web browser,
+the users can use any client, no matter what hardware or host operating system
+it is compromised of.
 A browser is the only software component required to use the novel DaaS service.
@@
 
-This document is structured as follows.
-Section~\ref{sec:relatedworkAbschnitt} discusses related work on DaaS solutions from an academic perspective, whereas section~\ref{sec:RecommendDaaSarchitecture} describes the state-of-the-art DaaS architecture DESIGN we designed and implemented.
-Section~\ref{sec:AnalysisPossibleComponents} presents the most relevant challenges we faced, analyses our options for handling these, and describes the solution we found.
-Finally, section~\ref{sec:Conclusions} discusses conclusions and includes directions for future work.
+Therefore, this document discusses in Section~\ref{sec:relatedworkAbschnitt}
+the related work on DaaS solutions from an academic perspective,
+whereas Section~\ref{sec:RecommendDaaSarchitecture}
+describes the state-of-the-art DaaS architecture DESIGN we designed and implemented.
+Section~\ref{sec:AnalysisPossibleComponents} presents the most relevant challenges we faced,
+analyses our options for handling these,
+and describes the solution we found.
+Finally, Section~\ref{sec:Conclusions} discusses conclusions
+and includes directions for future work.
 

### 2024-05-10 paper.tex
Minor spellings in related work

@@
 
-Magana et al.~\cite{magana2019remote} compared in 2019 the most popular remote desktop protocols and software implementations.
-The authors analyzed five protocols: PCoIP used in the Amazon WorkSpaces, RDP, TeamViewer, VNC, and Citrix Independent Computing Architecture (ICA).
-In the paper, the network transfer rate and its relation to the quality experienced by the DaaS user are evaluated by assuming three scenarios: using an office software suite, web browsing, and video streaming.
-One emphasis of the paper is comparing the transfer rate of the different protocols for three scenarios.
-In the Office Software Suite scenario, the transfer rate is measured and compared when performing several tasks like opening and editing a document (typing), loading an image, and saving a document.
-The web browsing scenario covers requesting three different web pages, and for the video streaming scenario, the same YouTube video file was viewed at different resolutions.
-The authors conclude that RDP requires fewer downstream and upstream network resources for all three scenarios.
-It is important to understand that the protocols RDP and VNC are implemented in open-source projects and proprietary solutions for many different operating systems.
-Depending on the various quality levels of the many existing implementations, it is difficult to generalize about the performance of the protocols, as individual implementations may have better or worse performance characteristics than expected.
+Magana et al.~\cite{magana2019remote} compared in 2019
+the most popular remote desktop protocols and software implementations.
+The authors analyzed five protocols:
+PCoIP used in the Amazon WorkSpaces, RDP, TeamViewer, VNC,
+and Citrix Independent Computing Architecture (ICA).
+In the paper, the network transfer rate and its relation to the quality
+experienced by the DaaS user are evaluated
+by assuming three scenarios:
+using an office software suite, web browsing, and video streaming.
+One emphasis of the paper is comparing the transfer rate
+of the different protocols for three scenarios.
+In the Office Software Suite scenario,
+the transfer rate is measured and compared
+when performing several tasks like opening
+and editing a document (typing), loading an image, and saving a document.
+The web browsing scenario covers requesting three different web pages,
+and for the video streaming scenario,
+the same YouTube video file was viewed at different resolutions.
+The authors conclude that RDP requires fewer downstream
+and upstream network resources for all three scenarios.
+Generally can be said, that protocols RDP and VNC
+are among the most popular frame-buffer protocols
+which are utilized in countless open-source and proprietary projects
+for different operating systems.
+Depending on the various quality levels of the many existing implementations,
+it is difficult to generalize about the performance of the protocols,
+as individual implementations may have better
+or worse performance characteristics than expected.
 
@@
 consider a maximum tolerable latency, based on subjective tests,
-being around 100\,ms. This relatively low value sets the demand for a fast network interaction
+being around 100\,ms.
+This relatively low value sets the demand for a fast network interaction
 between client and service and limits their maximum distance.
@@
 However, just as with computer games,
-the lagging operation of desktop programs has a negative impact on the user experience,
+the lagging operation of desktop programs
+has a negative impact on the user experience,
 and with video conferencing or other graphics-intensive applications,
-the utilization of resources and latency requirements are very similar to computer games.
+the utilization of resources and latency requirements
+are very similar to computer games.
 

### 2024-05-10 paper.tex
Minor spellings in chapter architecture

@@
 	\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
-%
 \maketitle              % typeset the header of the contribution
-%
+
 \begin{abstract}
@@
 \end{abstract}
-%
-%
-%
+
 \section{Introduction}
-
 This paper discusses challenges during the development
@@
 
-\section{Related Work}
-\label{sec:relatedworkAbschnitt}
-
-DaaS has been a well-known service category since the emergence of cloud computing. Still, it got much less attention in research and literature.
-The challenges and obstacles when developing DaaS solutions, rather than infrastructure (IaaS) and platform services (PaaS), have seldom been discussed in the literature.
-
-Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using the IaaS solution OpenStack~\cite{sefraoui2012openstack} and analyzed the characteristics and performance aspects of using noVNC, which is a client for the protocol Virtual Network Computing (VNC) that is implemented in the Hypertext Markup Language (HTML) and JavaScript, the protocol SPICE (Simple Protocol for Independent Computing Environments), and the remote desktop gateway Apache Guacamole as solutions for providing access to the desktop via a browser.
-One focus of the paper is the redirection of the sound interface of the virtual desktop.
-The paper's authors conclude that Guacamole, in combination with the Remote Desktop Protocol (RDP), is the best solution.
-The authors evaluated the lag time for remote audio playback in an Internet scenario by measuring the time between the automated initiation of the live streaming using the media player software VLC and the arrival of the music data by analyzing network traffic via the monitoring software Wireshark.
-When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms.
-When using Guacamole combined with the protocol VNC, the average lag was around 1750\,ms.
-The paper also includes a short evaluation of the lag for video frames represented by background color changes of a Java application.
-Again, the configuration Guacamole and RDP offered the best performance video update with an average lag of around 300\,ms.
+\section{Related Work}\label{sec:relatedworkAbschnitt}
+
+DaaS has been a well-known service category since the emergence of cloud computing.
+Still, it got much less attention in research and literature.
+The challenges and obstacles when developing DaaS solutions,
+rather than infrastructure (IaaS) and platform services (PaaS),
+have seldom been discussed in the literature.
+
+Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS
+using the IaaS solution OpenStack~\cite{sefraoui2012openstack}
+and analyzed the characteristics and performance aspects of using noVNC,
+which is a client for the protocol Virtual Network Computing (VNC)
+that is implemented in the Hypertext Markup Language (HTML) and JavaScript,
+the protocol SPICE (Simple Protocol for Independent Computing Environments),
+and the remote desktop gateway Apache Guacamole as solutions
+for providing access to the desktop via a browser.
+One focus of the paper is the redirection
+of the sound interface of the virtual desktop.
+The paper's authors conclude that Guacamole,
+in combination with the Remote Desktop Protocol (RDP), is the best solution.
+The authors evaluated the lag time for remote audio playback
+in an Internet scenario by measuring the time between the automated initiation
+of the live streaming using the media player software VLC
+and the arrival of the music data by analyzing network traffic
+via the monitoring software Wireshark.
+When using Guacamole combined with the protocol RDP,
+the average lag was around 750\,ms.
+When using Guacamole combined with the protocol VNC,
+the average lag was around 1750\,ms.
+The paper also includes a short evaluation of the lag for video frames
+represented by background color changes of a Java application.
+Again, Guacamole and RDP offered
+the best video performance with an average lag of around 300\,ms.
 The configuration of Guacamole and VNC caused an average lag of around 500\,ms.
@@
 
-\section{Architecture}
-\label{sec:RecommendDaaSarchitecture}
+\section{Architecture}\label{sec:RecommendDaaSarchitecture}
 % 
-Figure~\ref{figure_architecture} shows the architecture of our novel DaaS solution DESIGN that can run unmodified Linux and Windows applications in Linux containers (Linux and Windows applications compatible with Wine) and virtual machines (Windows applications incompatible with Wine).
+Figure~\ref{figure_architecture} shows the architecture
+of our novel DaaS solution DESIGN
+that can run unmodified Linux and Windows applications in Linux containers
+(Native Linux applications and Native Windows applications compatible with Wine)
+and virtual machines (Native Windows applications incompatible with Wine).
 \begin{figure}
@@
 
-Exporting the graphical user interface of Linux and Windows applications is possible using VNC or RDP protocols.
+Exporting the graphical user interface of Linux and Windows applications
+is possible using VNC or RDP protocols.
 Free server implementations exist for both operating system families.
-Since our DaaS solution focuses on exporting only the application's graphical user interface and not full desktops -- a feature not all server implementations include -- the number of available projects and solutions is limited.
-For Linux containers, the free software projects x11vnc and Xrdp can be used in the DESIGN DaaS.
-For Windows VMs, the free software TightVNC and the Windows-internal RDP server can be used.
-
-One of the main goals of DESIGN is to allow all interactions to be carried out solely using a browser.
-Thus, a broker or proxy software that mediates between the VNC or RDP implementation and the browser is required.
-The most advanced solution available is the free software project Apache Guacamole, which supports VNC and RDP and includes further relevant features regarding sound and printing.
-
-The core components of DESIGN are our self-developed Web UI and the Control Application, which carry out the communication between all service components.
+Since our DaaS solution focuses on exporting only the application's graphical user interface
+and not full desktops -- a feature not all server implementations include --
+the number of available projects and solutions is limited.
+For Linux containers, the free software projects
+x11vnc and Xrdp can be used in the DESIGN DaaS.
+For Windows VMs, the free software TightVNC
+and the Windows-internal RDP server can be used.
+
+One of the main goals of DESIGN is to allow all interactions to be carried out
+solely using a browser.
+Thus, a broker or proxy software that mediates between the VNC or RDP
+implementation and the browser is required.
+The most advanced solution available
+is the free software project Apache Guacamole, which supports VNC and RDP
+and includes further relevant features regarding sound and printer support.
+
+The core components of DESIGN are our self-developed Web UI
+and the Control Application,
+which carry out the communication between all service components.
 
 Starting a virtual machine or container instance automatically includes an SSH service.
-Our self-developed instance component waits until the SSH service is ready to satisfy requests.
-If applications are to be started or modifications are to be carried out, the instance component sends these requests to the virtual machine or container.
-Modification requests include resolution changes and fetching host status information.
-
-The self-developed instance component generally allows users to carry out essential administration and usage tasks that the operating system does not offer via a generalized API.
-
-% TODO: Die letzten 3 Abschnitte noch überarbeiten und präzisieren
-
-\section{Challenges and Solutions}
-\label{sec:AnalysisPossibleComponents}
+Our self-developed instance component waits
+until the SSH service is ready to satisfy requests.
+If applications are to be started or modifications are to be carried out,
+the instance component sends these requests to the virtual machine or container.
+Modification requests include resolution changes
+and fetching host status information.
+
+The self-developed instance component generally allows users
+to carry out essential administration
+and usage tasks that the operating system does not offer via a generalized API.
+
+\section{Challenges and Solutions}\label{sec:AnalysisPossibleComponents}
 

### 2024-05-10 paper.tex
Change delay in chapter 'Internal Processing and Communication[1~#

@@
 
-While developing the novel DaaS platform DESIGN, we faced multiple obstacles and challenges.
-Each issue belongs to one of four categories: compatibility, performance, stability, and usability.
-This section presents the issues we faced, the possible approaches we analyzed, and the solutions we chose.
+While developing the novel DaaS platform DESIGN,
+we faced multiple obstacles and challenges.
+Each issue belongs to one of four categories:
+compatibility, performance, stability, and usability.
+This section presents the issues we faced,
+the possible approaches we analyzed, and the solutions we chose.
 
@@
 
-The first relevant characteristic of a DaaS that a user is confronted with is the list of compatible applications.
-In the best case, all applications of all popular operating systems can be integrated and used with such a novel platform.
-Several DaaS projects and solutions from the last two decades and similar concepts like web desktops or webtops lacked native support for popular applications.
+The first relevant characteristic of a DaaS
+that a user is confronted with
+is the list of compatible applications.
+In the best case,
+all applications of all popular operating systems
+can be integrated and used with such a novel platform.
+Several DaaS projects and solutions from the last two decades
+and similar concepts like web desktops
+or webtops lacked native support for popular applications.
 
 Examples worth mentioning in this context include eyeOS and SilveOS.
-The project eyeOS~\cite{liu2012research,vidyabanu2011implementation} simulates a desktop environment using HTML, the scripting language PHP, JavaScript, and the MySQL database management system.
-The project SilveOS~\cite{garmpis2016design} mimics the look and feel of a Windows desktop in the browser using the discontinued application framework Microsoft Silverlight.
-Despite being free software, both solutions have never become widely used since it is impossible to import and use native Windows or Linux applications there.
+The project eyeOS~\cite{liu2012research,vidyabanu2011implementation}
+simulates a desktop environment using HTML,
+the scripting language PHP, JavaScript,
+and the MySQL database management system.
+The project SilveOS~\cite{garmpis2016design}
+mimics the look and feel of a Windows desktop in the browser
+using the discontinued application framework Microsoft Silverlight.
+Despite being free software,
+both solutions have never become widely
+used since it is impossible to import
+and use native Windows or Linux applications there.
 All required applications have to be developed as web applications.
-This limitation caused most potential users to lose interest in web desktops because many popular native applications can hardly be replaced or re-developed, and many feature-rich and well-established Software-as-a-Service (SaaS) offerings like Microsoft Office 365 and Google Docs already exist.
-
-A new DaaS is only accepted by a broad user group if it enables the integration of native applications in an unchanged manner.
-It is also desirable that a DaaS supports Linux and Windows applications both to allow flexible working and make the operation of a virtualization solution on the desktops unnecessary.
+This limitation caused most potential users to lose interest in web desktops
+because many popular native applications can hardly be replaced or re-developed,
+and many feature-rich and well-established Software-as-a-Service (SaaS)
+offerings like Microsoft Office 365 and Google Docs already exist.
+
+A new DaaS is only accepted by a broad user group if it
+enables the integration of native applications in an unchanged manner.
+It is also desirable that a DaaS supports Linux and Windows applications
+both to allow flexible working and make the operation
+of a virtualization solution on the desktops unnecessary.
 Ideally, it should also be possible to run native Mac OS applications.
-However, this goal is difficult to achieve under legal conditions due to the Mac OS operating system's licensing restrictions.
-
-Further essential features for most users are printing, sound, and compatibility with external storage media.
-
-Guacamole supports redirecting audio, printing, and disk access via the protocol RDP.
-The printing feature does not allow access to local or remote printers; instead, it allows users to print directly to PDF files, which are received by the user's web browser.
-As a prerequisite, the GhostScript interpreter must be deployed on the Guacamole server.
-File transfer over RDP supports Guacamole by emulating a virtual disk drive that persistently stores data on the Guacamole server.~\cite{ApacheGuacamole2024}
-
-Compared to RDP, the protocol VNC offers fewer features in Guacamole since it does not support sound on its own.
-However, Guacamole supports sound using PulseAudio\footnote{On Linux operating systems, sound and microphone devices are accessed via sound servers like PulseAudio and PipeWire, which interact with the sound interface present in the kernel, e.g., Advanced Linux Sound Architecture (ALSA) or Open Sound System (OSS)} as a workaround solution.
+However, this goal is difficult to achieve under legal conditions
+due to the Mac OS operating system's licensing restrictions.
+
+Further essential features for most users are printing, sound,
+and compatibility with external storage media.
+
+Guacamole supports redirecting audio, printing,
+and disk access via the protocol RDP.
+The printing feature does not allow access to local or remote printers;
+instead, it allows users to print directly to PDF files,
+which are received by the user's web browser.
+As a prerequisite, the GhostScript interpreter
+must be deployed on the Guacamole server.
+File transfer over RDP supports Guacamole
+by emulating a virtual disk drive that persistently stores data
+on the Guacamole server.~\cite{ApacheGuacamole2024}
+
+Compared to RDP, the protocol VNC offers fewer features in Guacamole
+since it does not support sound on its own.
+However, Guacamole supports sound using
+PulseAudio\footnote{On Linux operating systems,
+	sound and microphone devices are accessed via sound servers like PulseAudio and PipeWire,
+	which interact with the sound interface present in the kernel,
+	e.g., Advanced Linux Sound Architecture (ALSA) or Open Sound System (OSS)}
+as a workaround solution.
 VNC also does not offer file transfer and access to external storage media.
-Still, as a workaround solution, Guacamole supports file transfer via the SSH File Transfer Protocol (SFTP), which is an extension of the Secure Shell protocol (SSH).
+Still, as a workaround solution, Guacamole supports file transfer
+via the SSH File Transfer Protocol (SFTP),
+which is an extension of the Secure Shell protocol (SSH).
 
@@
 In a minimal setup, we measured delays of around 10ms for client communication,
-below 1ms for internal backend communication
+around 100ms for internal backend communication
 but up to 1000ms for instance communication.
@@
 as certain optimization concepts remain to be implemented.
+% Explanation delays: Half-way of a Roundtrip.
+% We state 1000ms for instance-communication (500ms per connection, we make two, so 1000ms)
+% We state 100ms for internal comm, which is a roundtrip of 200ms for auth divided by two, so 100ms
+% 10ms for client delay we can take a clientside timestamp when sending request and a server side timestamp when the request is intially received.
+% So we directly measure a deley, no rtt. The delay varias between 9 and 14ms, so we state roughly 10ms.
+% In my newest tests this number goes up to 50ms, but thats ok for what we receive in advance.
 

### 2024-05-10 paper.tex
Typo in heading 'Internal Processing and Communication'

@@
 
-\subsubsection{Internal processing and Communication}
+\subsubsection{Internal Processing and Communication}
 Another critical aspect directly related to performance and responsiveness

### 2024-05-10 paper.tex
Typo in Redundancy of Backend Services
Typos in Conclusion

@@
 	\item \textbf{Instances:} Redundancy of instances
-	      It is not a necessary condition for DaaS systems in general,
+	      is not a necessary condition for DaaS systems in general,
 	      but applying such strategies under certain conditions might be plausible.
@@
 
-\section{Conclusion and Outlook}
-\label{sec:Conclusions}
+\section{Conclusion and Outlook}\label{sec:Conclusions}
 
@@
 and will provide a full-fledged performance analysis
-of realsitic deployment scnarios and utilization patterns
+of realistic deployment scenarios and utilization patterns
 which include an evaluation of different user behaviors

### 2024-06-01 paper.tex
Final typo cheker commit
Cleaned up some sources and fixed some horizontal spacings between footnotes and text

@@
 directly interacting with GPU drivers,
-such as Moonlight~\cite{hanimplementation}~\cite{darcy2022latency}.
+such as Moonlight~\cite{darcy2022latency,hanimplementation}.
 
@@
 by emulating a virtual disk drive that persistently stores data
-on the Guacamole server.~\cite{ApacheGuacamole2024}
+on the Guacamole server.\footnote{\url{https://guacamole.apache.org/doc/1.5.5/gug/configuring-guacamole.html}}
 
@@
 highlight the apparent advantage of a lightweight container solution
-over virtual machines when identical programs are tested in both environments
-~\cite{felter2015updated,potdar2020performance}.
+over virtual machines when identical programs are tested in both
+environments~\cite{felter2015updated,potdar2020performance}.
 
@@
 studies show that there is no dramatic difference in performance
-when derivatives of operating systems are compared to each other
-~\cite{balen2020performance,boras2020performance}.
+when derivatives of operating systems are compared to each
+other~\cite{balen2020performance,boras2020performance}.
 Nevertheless, some derivatives are more lightweight and performant.
@@
 low-level operating system functionality implementations
-such as interrupt handling, memory allocation, and boot time
-~\cite{sulaiman2021comparison}.
+such as interrupt handling, memory allocation, and boot
+time~\cite{sulaiman2021comparison}.
 Additional performance potentials can be leveraged
 in cases where a Windows virtual machine can be replaced
-with a container-based solution integrating wine
-~\cite{huang2012performance}.
+with a container-based solution integrating 
+wine~\cite{huang2012performance}.
 
@@
 
-Allthough we can not yet provide a full-fledged performance analysis,
+Although we can not yet provide a full-fledged performance analysis,
 we already measured promising results in our local test environment.
@@
 but up to 1000ms for instance communication.
-But these numbers are only rough estimates
+However, these numbers are only rough estimates
 as certain optimization concepts remain to be implemented.
@@
 \subsection{Stability}
-Stability in DaaS applications is especially important
+Stability in DaaS applications is essential
 for meeting the demands of continuous and time-critical operations
@@
 In contrast, vertical scaling (scale up),
-adds more power (CPU, RAM) to existing machines
-~\cite{vaquero2011dynamically}.
+adds more power (CPU, RAM) to existing 
+machines~\cite{vaquero2011dynamically}.
 While horizontal scaling offers flexibility
@@
 	      and is, therefore, independent of surrounding components and contexts
-	      and can be externalized to a dedicated systems.
+	      and can be externalized to a dedicated system.
 	      Modern database solutions, like MariaDB,
 	      offer rich features for maintaining specific data consistency strategies
-	      such as data replication, clustering, failover mechanisms or RAID control.
+	      such as data replication, clustering, failover mechanisms, or RAID control.
 	      Currently, redundancies in our database system are not planned
@@
 This implies that any action triggered, whether in the frontend on the client side,
-in the backend on the host side, or within hosted instances,
+in the backend on the host side or within hosted instances,
 must ensure that all necessary information is readily available at the time of processing.
@@
 
-If that is the case, each distributed host system might only contain it's essential components,
+If that is the case, each distributed host system might only contain its essential components,
 including the backend service with all hosted instances
@@
 by using a custom window manager such,
-as for example, qtile~\footnote{\url{https://qtile.org/}}.
+as for example, qtile\footnote{\url{https://qtile.org}}.
 
@@
 To address this until such features are available in
-Guacamole~\footnote{\url{https://issues.apache.org/jira/browse/GUACAMOLE-1196}},
+Guacamole\footnote{\url{https://issues.apache.org/jira/browse/GUACAMOLE-1196}},
 we implemented a separate mechanism using a WebSocket proxy
@@
 with a more efficient approach as our preliminary performance tests revealed
-a potential bottlenack in that particular area.
+a potential bottleneck in that particular area.
 We also plan to apply certain optimization strategies
-with respect to backend and instance communication.
+concerning backend and instance communication.
 We further plan to enhance our yet limited performance testing
